### 2020-05-21 protocol.Rmd

## General notes

- Record everything in the lab notebook.
- Take note of any mistakes or errors.
- Ideally start flux measurements around 8:00 or 9:00 o'clock on Monday
  morning to allow time for three measurements on that day.

## Initial setup
1. Sterilize potting mix, DI water, and mason jar disks in the autoclave.
2. Sterilize jars with 70% ethanol.
3. Add 45 g of steril potting mix, 5 g of live soil, and 3.5 mL of sterile DI water to each jar and cap.
4. Spike each jar with 4.3 mL of 99% CH~4~.

Allow these to incubate for at least two weeks before taking a flux
measurements. Also, uncap them in the hood for ~30 minutes and respike with 4.3
mL of 99% CH~4~ every 3-4 days to maintain aerobic conditions and maintain
elevated CH~4~. So, for example, if the jars are spiked on Thursday, flush them
on Monday and Thursday of the following two weeks and then take the flux
measurements on the third Monday.

## Flux sampling

### Monday

1. Uncap jars in hood with blower on to return them to atmospheric gas
concentrations and to avoid cross-contamination.
2. Measure standard curve on GC with 5-12 points for CH~4~. Record the
   concentration of the standard in the bottle (ppm), the volume of the
   injected standard (mL), the
concentration of the injected standard (ppm), and the CH~4~ curve area. Also record the
slope and y-intercept for the resulting equation as well as the R^2^. Continue running standard curves until the R^2^ is 99% and ideally 99.9%.
3. Cap jars in the hood.
4. Spike jars to ~1000 ppm CH~4~ using 0.43 mL of 99% CH~4~ and record time for
   each jar to the minute (t~0~).
5. Measure t~1~ immediately after spiking all jars resulting in a t~1~ that is
20 to 30 minutes after t~0~. Measure jars in the order that they were spiked.
Record the time, CH~4~ curve area, and CO~2~ curve area.
6. Record t~2~ after 3-5 hours.
7. Record t~3~ after 5-8 hours.

### Tuesday

8. Record t~4~ after 24 hours.

### Wednesday

9. Record t~5~ after 48 hours.
10. Enter the standard curve and flux data into R. Visually inspect the graphs
    to determine how many points to include in the linear portion of the curve
    (the first 3 to 5 points) and calculate the linear decrease in CH~4~
    concentration (Vmax). 
11. Determine the top 2 (16.7%) or 3 (25%) of the ~12 replicate jars for the
    positive selection treatment (p). Homogenize and reserve the soil from
    those jars in the hood. Randomly select an equal number of jars from the
    neutral selection treatment (n). Homogenize and reserve the soil from those
    jars in the biosafety cabinet. Clearly label each of these containers.
12. Empty the rest of the soil from the jars into an autoclave container, cover
    it with aluminum foil, and label it "disposal". Clean the jars and lids
    with soap and rinse and dry thoroughly.
13. Add at least `r 45 * 24` grams of fresh potting mix to an autoclave tray
    and cover with aluminum foil. Wrap the disks in a foil packet. Set aside
    the rings, they do not need to be autoclaved.
14. Autoclave the potting soil, lids, and disposal soil on the gravity cycle for 40
    minutes. 
15. Remove from the autoclave, put the jars and lids into the hood, and allow
    to cool overnight.

### Thursday

16. Sterilize the inside of each jar with 70% ethanol.
17. Add 45 grams of sterile potting mix to each jar.
18. Add 5 grams of the selected soil homogenate to each 'p' jar and 5 grams of
    the random soil to each 'n' jar. 
19. Add 3.5 mL of sterile DI water to each jar. 
20. Cap each jar and spike with 4.3 mL CH~4~.

## Preferred data format

Dates should be recorded as `yyyy-mm-dd` and times should be recorded as
`yyyy-mm-dd hh:mm` (seconds are optional). Use standard column headers with no
caps and no spaces. Instead, simply use "variable_units" format.

Column Headers | Meaning
-------------- | ----------------------------------------------------------
`flux_date`    | The date flux measurements began (i.e. Monday)
`standard_ppm` | The concentration of the standard in the bottle. 
`injection_ml` | The volume of standard injected into the GC. 
`injectin_ppm` | The concentration of standard injected into the gc (standard_ppm multiplied by injection_ml) 
`ch4_area`     | The area-under-the-curve for the methane peak output from the GC.