### 2020-10-20 Presentation/presentation.Rmd
start presentation for lab meeting

---
title: "Artificial Ecosystem Selection Update"
author: "Andrew Morris"
date: "10/21/2020"
output: 
  revealjs::revealjs_presentation:
    transition: fade
    reveal_options:
      slideNumber: true
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(tidyverse)
library(metafor)
library(morris)
library(broman)
```

# Outline

## What I'm hoping to get out of this

- Show you what I've been working on
- Hear your questions to see if that guides me down any path
- ???

# The Question

## 
One of the fundamental questions in ecology is, "How does the biodiversity
  that we observe influence ecosystem functions?"

## 
Historically, attempts to address this question have focused on alpha
  diversity as it relates to the *rate* or *stability* of an ecosystem function

##
For certain functions, this works pretty well. For example, the stability of
  primary productivity in temperate plant communities in the face of drought is
  correlated with species richness of plants.

## 
However, microbial community data exhibit certain limitations that cause this
  approach to fail:

  1. Without a consistent species concept, it is not obvious at what level of
     biological organization to evaluate this relationship. (the ASV? the
     genus? the family?)
  2. It is not always clear which members of a community are members of a
     particular functional group. (This can be determiend morpholigically for
     plants, but must be determined genetically in natural microbial
     communities.)
  3. Microbial communities are immensely diverse and without solutions to the
     previous two problems, sifting through that diversity is challenging.

## When we apply this BEF approach to microbial communities, there are limited
  gains in terms of our knowlegde of the controls on ecosystem function

# A new approach to linking microbial communities to ecosystem function

## An alternative approach

- One way to think about microbial communities and ecosystem functions is as assemblages of genes within an ecosystem and a trait at the ecosystem level.
- In fact, this often the only way we can measure microbial communities in natural ecosystems.

## Experiment

I collected a single soil core from near UO campus and used sterile potting mix as the subtrate. Ecosystems were created in 500 mL mason jars with 45 g of sterile potting mix and 5 g of living soil and were brought to 60% of field capacity (the optimal level for aerobic processes such as methane oxidation). I created 24 jars in this way and randomly assigned them to a Positive or Neutral selection line.

# Is there a response to selection at the ecosystem level on soil methane oxidation rate?

##

```{r, fig.width=4, fig.height=3, out.width=6e2, out.height=4.5e2}

fluxes <- read_tsv('../Output/fluxes.tsv')
ggplot(fluxes, aes(x = passage, y = estimate, color = treat)) + 
  geom_jitter() +
  scale_color_manual(name = 'Treatment', labels = c('Neutral', 'Positive'),
                     values = c('gray60', 'darkorange2')) +
  labs(x = "Passage", y = 'Flux (-k)') +
  theme_bw()

```

##

```{r, fig.width=4, fig.height=3, out.width=6e2, out.height=4.5e2}

fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate),
            se = se(estimate)) %>% 
ggplot(mapping = aes(x = passage, y = mean, ymin = mean - se, ymax = mean + se,
                     color = treat)) + 
  geom_pointrange() +
  scale_color_manual(name = 'Treatment', values = c('gray60', 'darkorange2')) +
  theme_bw()
```

##

There was a response to selection on soil methane oxidation rate in the positive treatment relative to the neutral treatment.

```{r, fig.width=4, fig.height=3, out.width=6e2, out.height=4.5e2}

deviance <- read_tsv('../Output/deviance.tsv')
dev_fit <- readRDS('../Output/dev_fit.rds')
dev_pred <- predict(dev_fit)
passages <- length(deviance$passage)

ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_ribbon(aes(ymin = dev_pred$ci.lb, ymax = dev_pred$ci.ub), alpha = 0.2) +
  geom_pointrange(color = 'darkorange2') + 
      geom_segment(aes(x = 1, xend = passages, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*passages), color = 'darkorange2', size = 1) +
  labs(x = "Passage Number", y = "Deviation from Control (P - N)")+
  theme_bw()

```

```{r}
init <- 
  fluxes %>% 
  group_by(passage) %>% 
  summarize(mean = mean(estimate)) %>% 
  filter(passage == 1) %>% 
  pull()

init_se <- 
  fluxes %>% 
  group_by(passage) %>% 
  summarize(se = se(estimate)) %>% 
  filter(passage == 1) %>% 
  pull()
```

z = `r myround(dev_fit$zval[2], 2)`, p = `r myround(dev_fit$pval[2], 3)`

slope = `r myround(dev_fit$beta[2], 3)` $\pm$ `r myround(dev_fit$se[2], 3)`

(initial mean flux  = `r myround(init, 3)` $\pm$ `r myround(init_se, 3)`)