### 2021-06-17 R/02-process-herit.Rmd

---
title: "Process Heritability Data"
author: "Andrew Morris"
date: "`r Sys.Date()`"
output: html_document
---

This script takes in the flux data from script 01 and the data on which jars
were selected in each passage. It then aligns these data so that for each
passage there is a mean flux for all parents, the selected parents, and all
offspring.

```{r initialize_packages, message=FALSE}
library(tidyverse)
```

```{r initialize_variables}
source('functions.R')
```

```{r import_data, message=FALSE}
selected_jars <- read_csv(paste0(raw_dir, 'selected.csv'))
fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))
```

## Organize parent/offspring values

```{r process_herit_data}
# Pull out the selected parental jars
selected <- 
  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
  select(treat, estimate, passage)

# Pull out all parental jars
parental <-
  fluxes %>% 
  select(passage, treat, estimate)

# Pull out all offspring jars
# The `passage - 1` function sets the value of generation `n + 1` 
# as the "offspring" value for generation `n`. 
offspring <- 
  fluxes %>%
  select(treat, estimate, passage) %>% 
  filter(passage != 1) %>% 
  mutate(passage = passage - 1)

# Calculate means within each generation for all parental jars,
# all offspring jars, and all selected parental jars
parental_mean <-
  parental %>% 
  group_by(passage, treat) %>% 
  summarize(parental = mean(estimate), .groups = "drop")

offspring_mean <-
  offspring %>% 
  group_by(passage, treat) %>% 
  summarize(offspring = mean(estimate), .groups = "drop")

selected_mean <-
  selected %>% 
  group_by(passage, treat) %>% 
  summarize(selected = mean(estimate), .groups = "drop")

# Combine all parent/offspring data into one data frame
heritability <- 
  parental_mean %>% 
  left_join(offspring_mean, by = c("passage", "treat")) %>% 
  left_join(selected_mean, by = c("passage", "treat")) %>% 
  #filter(treat == "p") %>% 
  filter(passage != 5) %>% # Remove passage 5, which has no offspring,
  # because it's the last generation
  ungroup()

write_tsv(heritability, paste0(der_dir, 'heritability.tsv'))
```