### 2021-06-17 R/03-response.Rmd

---
title: "Estimate Response to Selection"
author: "Andrew Morris"
date: "`r Sys.Date()`"
output: html_document
---

This script intakes the flux data from script 01. It then calculates deviance
as the difference between the positive line and the neutral line in each
passage. The errors can either be calculated as the sum of the errors for
each line or the standard error of difference, that is, the square root of
the sum of the squared errors, which is slightly smaller than the sum. In
this case, I have chosen the SED.
```{r initialize_packages, message=FALSE}
library(tidyverse)
library(metafor)
library(morris)
library(broom)
```

```{r initialize_variables}
source('functions.R')
```

```{r import_data, message=FALSE}
fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))
```

Standard error of the difference function

```{r functions}
# Standard error of the difference
sed <- function(p, n, na.rm = FALSE) {
  sqrt(se(p, na.rm = na.rm) ^ 2 + se(n, na.rm = na.rm)^ 2)
}
```

## Plot the raw data

```{r}
ratio.values <- (max(fluxes$passage)-min(fluxes$passage))/(max(fluxes$estimate)-min(fluxes$estimate))

plot_flux <- function(fluxes, passage, estimate) {
ggplot(fluxes, aes(x = {{passage}}, y = {{estimate}}, color = treat)) +
  theme_classic() +
  geom_jitter(aes(alpha = selected), width = 0.1) +
  stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2')) +
  scale_alpha_manual(name = "Selected", labels = c('Not', 'Selected'), values = c(0.3, 1))
}

plot_log_flux <- function(fluxes, passage, estimate_log10) {
ggplot(fluxes, aes(x = {{passage}}, y = {{estimate_log10}}, color = treat)) +
  theme_classic() +
  geom_jitter(aes(alpha = selected), width = 0.1) +
  stat_smooth(aes(group = treat), method = 'lm', se = FALSE, formula = 'y ~ x') +
  #  coord_fixed(ratio.values) +
  labs(x = "Passage Number", y = expression(log[10] * "(Methane Consumption Rate (-k))")) +
  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2')) +
  scale_alpha_manual(name = "Selected", labels = c('Not', 'Selected'), values = c(0.3, 1))
}

plot_flux(fluxes, passage, estimate)

#ggsave('fluxes_raw.pdf', width = 5, height = 4)

```

## 3. Plot the log-transformed data

```{r}
# ratio.values <- (max(fluxes$passage)-min(fluxes$passage))/(max(fluxes$estimate_log10)-min(fluxes$estimate_log10))
# 
# selected$selected <- TRUE
# 
# selected_data <- 
# fluxes %>% 
#   left_join(selected, by = c('passage', 'jar')) %>% 
#   filter(selected == TRUE)

plot_log_flux(fluxes, passage, estimate_log10)

ggplot(fluxes, aes(x = passage, y = estimate_log10, color = treat)) +
  theme_classic() +
  geom_jitter(aes(alpha = selected), width = 0.2) +
  stat_smooth(aes(group = treat), method = 'lm', se = FALSE, formula = 'y ~ x') +
#  coord_fixed(ratio.values) +
  labs(x = "Passage Number", y = "log10(Methane Oxidation Rate (-k))") +
  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2')) +
  scale_alpha_manual(name = "Selected", labels = c('Not', 'Selected'), values = c(0.5, 1))

#ggsave('fluxes_log.pdf', width = 5, height = 4)
```

```{r}
ggplot(fluxes, aes(x = passage, y = estimate_log10, color = treat)) +
  theme_classic() +
  geom_jitter(width = 0.2) +
  stat_smooth(aes(group = treat), method = 'lm', se = FALSE, formula = 'y ~ x') +
  #  coord_fixed(ratio.values) +
  labs(x = "Passage Number", y = expression(log[10] * "(Methane Consumption Rate (-k))")) +
  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2'))

#ggsave('fluxes_log.png', width = 5, height = 4)
```

# 3. Test a difference of slopes

```{r}
fit <- lm(estimate_log10 ~ 0 + passage * treat, data = fluxes)
summary(fit)
fit
10^fit$coef["passage:treatp"]

fit1 <- lm(estimate_log10 ~ passage + treat, data = fluxes)
summary(fit1)
anova(
      lm(estimate_log10 ~ passage * treat, data = fluxes),
      lm(estimate_log10 ~ passage + treat, data = fluxes)
)
```

4. Plot the back-transformed slope

```{r}
summary(fit)
fit$coef[[3]]
fit$coef[[4]]

ggplot(fluxes, aes(x = passage, y = estimate, color = treat)) +
  theme_classic() +
  geom_jitter() +
  geom_abline(intercept = fit$coef[[3]], slope = fit$coef[[4]]^10) +
  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2'))
```

```{r}
####################################################################
# End the correct way to do it
################################################################

# Calculate deviance as p - n
deviance <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate), .groups = "drop") %>% 
  pivot_wider(names_from = treat, values_from = mean) %>% 
  mutate(deviance = p - n)

# Calculate variance of deviance as SED
deviance$se <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  mutate(x = 1:nrow(fluxes)) %>% 
  pivot_wider(names_from = treat, values_from = estimate) %>% 
  group_by(passage) %>% 
  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>%
  pull(sed)

# Fit deviance model
fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)

plot(fit)

# Plot deviance
ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange(color = 'darkorange2') + 
  geom_abline(intercept = fit$beta[1], slope = fit$beta[2])

# Take deviance model parameters and export into a data.frame
rma_output <- data.frame(
Estimate = fit$beta,
se = fit$se,
z.value = fit$zval,
p.value = fit$pval,
upper.ci = fit$ci.lb,
lower.ci = fit$ci.ub
)

# Fit deviance model without passage 2
fit_no2 <- 
filter(deviance, passage != 2) %>% 
rma(deviance ~ passage, se^2, method="FE", data = .)

# Take deviance model withou passage 2 parameters and export into a data.frame

rma_output_no2 <- data.frame(
Estimate = fit_no2$beta,
se = fit_no2$se,
z.value = fit_no2$zval,
p.value = fit_no2$pval,
upper.ci = fit_no2$ci.lb,
lower.ci = fit_no2$ci.ub
)

# R (response to selection) for the model w/o sample 2
R_no2 <- rma_output_no2[[2, 1]]
R <- rma_output[[2, 1]]

# Plot the no2 model
filter(deviance, passage != 2) %>% 
ggplot(., aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange(color = 'darkorange2') + 
  geom_abline(intercept = fit_no2$beta[1], slope = fit_no2$beta[2])

#write_tsv(deviance, '../Output/deviance.tsv')
#saveRDS(rma_output, '../Output/response.rds')
#saveRDS(fit, '../Output/dev_fit.rds')

# Calculate deviance as percentage of p - n
deviance <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate), .groups = "drop") %>% 
  pivot_wider(names_from = treat, values_from = mean) %>% 
  mutate(deviance = (p - n) / n)

# Calculate variance of deviance as SED
deviance$se <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  mutate(x = 1:nrow(fluxes)) %>% 
  pivot_wider(names_from = treat, values_from = estimate) %>% 
  group_by(passage) %>% 
  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>%
  pull(sed)

# Fit deviance model
fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)

plot(fit)

# Plot deviance
ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange(color = 'darkorange2') + 
  geom_abline(intercept = fit$beta[1], slope = fit$beta[2])

# Take deviance model parameters and export into a data.frame
rma_output <- data.frame(
Estimate = fit$beta,
se = fit$se,
z.value = fit$zval,
p.value = fit$pval,
upper.ci = fit$ci.lb,
lower.ci = fit$ci.ub
)

# Fit separte slopes for Positive and Neutral with robust M-estimator regression

library(MASS)
anova(
fit <- rlm(estimate ~ passage * treat, data = fluxes),
fit <- rlm(estimate ~ passage + treat, data = fluxes))

summary(fit)
anova(fit)
new_data <- cbind(fluxes, predict(fit, interval = 'confidence'))
colnames(new_data) 
ggplot(new_data, aes(x = passage, y = estimate, color = treat)) +
  theme_bw() +
  geom_jitter() +
  geom_line(aes(passage, fit)) +
  geom_ribbon(aes(ymin=lwr,ymax=upr), alpha=0.3)

ggplot(new_data, aes(x = passage, y = estimate)) +
  theme_bw() +
  geom_jitter(aes(color = treat)) +
  geom_line(aes(passage, fit, color = treat)) +
  geom_ribbon(aes(ymin=lwr,ymax=upr, group = treat), alpha=0.3) +
  ylim(-0.01, 0.08)
ggplot(fluxes, aes(x = passage, y = estimate, color = treat)) + 
  geom_point() + 
  stat_smooth(method='lm')
#ggsave('raw_fluxes.pdf')
```