### 2021-06-24 R/03-response.Rmd
convert response to selection script to rmd

---
title: "Estimate Response to Selection"
author: "Andrew Morris"
date: "`r Sys.Date()`"
output: html_document
---

This script intakes the flux data from script 01. It then fits a linear model to the flux data over time per treatment and tests a difference of slope between the positive and the neutral treatment. Finally, it 
```{r initialize_packages, message=FALSE}
library(tidyverse)
library(morris)
library(broom)
library(knitr)
#library(broman)
```

```{r initialize_variables}
source('functions.R')
```

```{r import_data, message=FALSE}
fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))

fluxes <-
  fluxes %>% 
  mutate(passage = passage - 1)
```

## Plot the raw data

```{r plot_raw}
ratio <- calc_plot_ratios(fluxes$passage, fluxes$ch4)

plot_flux(fluxes, passage, ch4, ratio)
#ggsave('fluxes_raw.pdf', width = 5, height = 4)
```

## Plot the log-transformed data

```{r plot_log10}
ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)

plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)

#ggsave('fluxes_log.pdf', width = 5, height = 4)
```

## Test a difference of slopes

```{r response_model}
# Fit full interaction model
inter_model <- lm(log_ch4 ~ passage * treat, data = fluxes)

# No treatment model
no_treat_model <- lm(log_ch4 ~ passage, data = fluxes)

# Back calculate the slope of the positive treatment
pos_slope <- 10^inter_model$coef["passage:treatp"]

```

Here I fit two models with flux as the dependent variable and passage and treatment as the independent variables. The first is the interaction model, which includes an interaction between passage and treatment. The second is an additive model that lacks this interaction. Here is the model comparison using `anova`, which indicates the full model is somewhat better:

```{r compare_models}
# Compare interaction to no interaction
compare_mods <- anova(inter_model, no_treat_model)
kable(tidy(compare_mods))
```

Here is the full interaction model:

```{r inter_model}
kable(tidy(inter_model))

```

The intercept for the Neutral treatment was significantly different from zero at `r round(tidy(inter_model)[[1, 2]], 2)` (SE = `r round(tidy(inter_model)[[1, 3]], 2)`, t = `r round(tidy(inter_model)[[1, 4]], 2)`, p = `r round(tidy(inter_model)[[1, 5]], 2)`). The Positive treatment started at a significantly lower methane oxidation rate than the Neutral treatment with a y-intercept of `r round(tidy(inter_model)[[3, 2]], 2)` relative to Neutral (SE = `r round(tidy(inter_model)[[3, 3]], 2)`, t = `r round(tidy(inter_model)[[3, 4]], 2)`, p = `r round(tidy(inter_model)[[3, 5]], 2)`). There was no change in methane oxidation rate of the Neutral treatment over the five passages (slope = `r round(tidy(inter_model)[[2, 2]], 2)`, SE = `r round(tidy(inter_model)[[2, 3]], 2)`, t = `r round(tidy(inter_model)[[2, 4]], 2)`, p = `r round(tidy(inter_model)[[2, 5]], 2)`). By contrast, there was a significant increase in the Positive treatment of `r round(tidy(inter_model)[[4, 2]], 2)` per passage which is approximately a `r round((pos_slope - 1) * 100, 0)`% increase in methane oxidation rate per passage (SE = `r round(tidy(inter_model)[[4, 3]], 2)`, t = `r round(tidy(inter_model)[[4, 4]], 2)`, p = `r round(tidy(inter_model)[[4, 5]], 2)`).

## Previous Method

Previously, I was using the `metafor` package to test the slope of the difference between the positive and neutral line. For this model, I calculate the difference between the Positive and Neutral treatment and fit that difference as the response. I also calculate the standard error of the difference and included that variance in the model. I then fit the model `difference ~ passage` to test the slope of the difference over time.

```{r eval=FALSE}

library(metafor)
# Standard error of the difference
sed <- function(p, n, na.rm = FALSE) {
  sqrt(se(p, na.rm = na.rm) ^ 2 + se(n, na.rm = na.rm)^ 2)
}
# Calculate deviance as p - n
deviance <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate), .groups = "drop") %>% 
  pivot_wider(names_from = treat, values_from = mean) %>% 
  mutate(deviance = p - n)

# Calculate variance of deviance as SED
deviance$se <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  mutate(x = 1:nrow(fluxes)) %>% 
  pivot_wider(names_from = treat, values_from = estimate) %>% 
  group_by(passage) %>% 
  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>%
  pull(sed)

# Fit deviance model
fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)

plot(fit)

# Plot deviance
ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange(color = 'darkorange2') + 
  geom_abline(intercept = fit$beta[1], slope = fit$beta[2])

# Take deviance model parameters and export into a data.frame
rma_output <- data.frame(
Estimate = fit$beta,
se = fit$se,
z.value = fit$zval,
p.value = fit$pval,
upper.ci = fit$ci.lb,
lower.ci = fit$ci.ub
)

# Fit deviance model without passage 2
fit_no2 <- 
filter(deviance, passage != 2) %>% 
rma(deviance ~ passage, se^2, method="FE", data = .)

# Take deviance model withou passage 2 parameters and export into a data.frame

rma_output_no2 <- data.frame(
Estimate = fit_no2$beta,
se = fit_no2$se,
z.value = fit_no2$zval,
p.value = fit_no2$pval,
upper.ci = fit_no2$ci.lb,
lower.ci = fit_no2$ci.ub
)

# R (response to selection) for the model w/o sample 2
R_no2 <- rma_output_no2[[2, 1]]
R <- rma_output[[2, 1]]

# Plot the no2 model
filter(deviance, passage != 2) %>% 
ggplot(., aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange(color = 'darkorange2') + 
  geom_abline(intercept = fit_no2$beta[1], slope = fit_no2$beta[2])

#write_tsv(deviance, '../Output/deviance.tsv')
#saveRDS(rma_output, '../Output/response.rds')
#saveRDS(fit, '../Output/dev_fit.rds')

# Calculate deviance as percentage of p - n
deviance <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate), .groups = "drop") %>% 
  pivot_wider(names_from = treat, values_from = mean) %>% 
  mutate(deviance = (p - n) / n)

# Calculate variance of deviance as SED
deviance$se <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  mutate(x = 1:nrow(fluxes)) %>% 
  pivot_wider(names_from = treat, values_from = estimate) %>% 
  group_by(passage) %>% 
  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>%
  pull(sed)

# Fit deviance model
fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)

plot(fit)

# Plot deviance
ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange(color = 'darkorange2') + 
  geom_abline(intercept = fit$beta[1], slope = fit$beta[2])

# Take deviance model parameters and export into a data.frame
rma_output <- data.frame(
Estimate = fit$beta,
se = fit$se,
z.value = fit$zval,
p.value = fit$pval,
upper.ci = fit$ci.lb,
lower.ci = fit$ci.ub
)

# Fit separte slopes for Positive and Neutral with robust M-estimator regression

library(MASS)
anova(
fit <- rlm(estimate ~ passage * treat, data = fluxes),
fit <- rlm(estimate ~ passage + treat, data = fluxes))

summary(fit)
anova(fit)
new_data <- cbind(fluxes, predict(fit, interval = 'confidence'))
colnames(new_data) 
ggplot(new_data, aes(x = passage, y = estimate, color = treat)) +
  theme_bw() +
  geom_jitter() +
  geom_line(aes(passage, fit)) +
  geom_ribbon(aes(ymin=lwr,ymax=upr), alpha=0.3)

ggplot(new_data, aes(x = passage, y = estimate)) +
  theme_bw() +
  geom_jitter(aes(color = treat)) +
  geom_line(aes(passage, fit, color = treat)) +
  geom_ribbon(aes(ymin=lwr,ymax=upr, group = treat), alpha=0.3) +
  ylim(-0.01, 0.08)
ggplot(fluxes, aes(x = passage, y = estimate, color = treat)) + 
  geom_point() + 
  stat_smooth(method='lm')
#ggsave('raw_fluxes.pdf')
```