### 2021-08-17 Manuscript/manuscript.Rmd
update manuscript and analyses and figures plus add citations

---
title: Artificial ecosystem selection to deduce the mapping between microbial community structure and ecosystem function
author: "Andrew H. Morris and Brendan J. M. Bohannan"
output:
  word_document:
    reference_docx: reference.docx
bibliography: bibliography.bib
editor_options:
  chunk_output_type: inline
---

```{r options, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r initialize_packages}
library(tidyverse)
library(captioner)
library(broom)
library(broman)
library(morris)
```

```{r load_data}
heritability <- read_tsv('../Output/heritability.tsv')
fluxes <- read_tsv('../Output/fluxes.tsv')

load(file = '../Output/response_model.Rdata')
load(file = '../Output/herit_model.Rdata')

source('../R/functions.R')

```

```{r figure_captions, include=FALSE}

figures <- captioner()

figures("response", "Response to selection on soil CH~4~ oxidation rate fit by ordinary least squares regression. The y-axis is the log~10~ of the additive inverse of the first-order exponential decay constant `k` (i.e., `-k`) with units day^-1^. The x-axis is the passage number. The orange line and points are the Positive treatment and the gray line and points are the Neutral treatment. The dark points are the selected jars in each passage and the light points are the non-selected jars. Regression lines aIn passage 2, the jars with the small `-k` were selected due to an error in the flux calculations.")

figures("herit", "Ordinary least squares regression of mid-offspring CH~4~ flux on mid-parent CH~4~ flux. Mid-parent is the mean of the jars selected to inoculate the next passage. Mid-offspring are the mean of all twelve jars produced in one passage. The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment.")

```

# Abstract

# Introduction

Soil microbial communities are central to the functioning of ecosystems
[@crowther2019]. And yet, microbes have historically been excluded from
ecosystems models in favor of abiotic variables [@schimel1995]. This is in part
because attempts to establish quantitative relationships between microbial
marker genes and ecosystem functions have not been fruitful [@graham2016,
@rocca2015]. Here we use a novel approach to estimate how much of the variation
in ecosystem function can be attributed to variation in the microbial community.
We then identify microbial markers of ecosystem function. These experiments will
expand our ability to model microbial ecosystem-function relationships and allow
us to manage ecosystems for particular ecosystem functions [@fuhrman2009,
@schimel1998].

In macroorganismal communities, there appears to be a relationship between the
diversity of a community and the resilience of ecosystem functions to
environmental perturbations [@hooper2005]. Typically, diversity is represented
as the richness of a particular functional group, such as photosynthetic plants.
Similar studies in microbial communities rarely find a relationship between
diversity and function [@graham2016]. However, these studies define
diversity much more broadly by looking at the richness of the entire bacterial
and archaeal community using 16S rRNA gene variants. Studies that focus more
narrowly on a particular functional group still rarely find a significant
correlation between the abundance of a functional group and the rate of the
corresponding process [@rocca2015]. For certain functions in certain
ecosystems, these relationships can be informative. For example, the abundance
of a single methanogenic species (Candidatus ‘*Methanoflorens
stordalenmirensis*’) is correlated with the CH~4~:CO~2~ ratio in a permafrost
ecosystem [@mccalley2014]. In addition, bacterial and archaeal community composition is
correlated with methane emissions in forests and pastures of the Brazilian
Amazon [@meyer2020]. Therefore, there appears to be some indication that
microbial community structure drives variation in methane cycling in soils.

Why have we been unsuccessful at identifying relationships between
microbial community structure and ecosystem functions? Microbial communities are
extremely diverse, particularly in soil environments [@whitman1998].
Characterizing these communities using broad phylogenetic markers, such as 16S,
reveals thousands of mostly uninformative taxa which makes it difficult to
identify markers of ecosystem function. Some studies constrain this diversity by
characterizing individual protein-coding genes. However, this approach assumes
investigators already know what limits the rate of a process in a particular
ecosystem. Typically, this is assumed to be the gene that encodes the final step
in a pathway--for example, the *mcrA* gene, which encodes a subunit of the
enzyme that performs the final step in methanogenesis. However, regulation may
happen further upstream (or downstream) within the pathway and therefore the
abundance of the final gene would not correlate with the rate of the process. In
the case of methane emissions, the limitation may be in generating the
substrates for methanogenesis or may be in the consumption of methane by
methanotrophs.

To further investigate the relationship between microbial community structure
and ecosystem function, we want to determine how much of the variation in soil
methane cycling is attributable to variation in the microbial community. This
ecological question is analogous to the problem of estimating heritability in
quantitative genetics [@morris2020]. Variation in an organismal trait is
determined by the sum of genetic variation, environmental variation, and the interaction between the two [@falconer1996]. One goal of quantitative genetics is to determine the
proportion of the variation attributable to genetics. This is commonly
estimated as the narrow-sense heritability (h^2^) defined as the additive
genetic variance. This goal can be achieved through artificial selection
experiments with selection on the character of interest [@falconer1996]. We
applied artificial ecosystem selection to whole microbial communities by
selecting on the ecosystem-scale methane oxidation rate in order to estimate the
additive genetic variance of soil methane oxidation [@swenson2000].

To identify which organisms are important for regulating methane cycling, we
compared the composition of the artificially selected microbial
communities and a control set of communities without selection. Artificial
ecosystem selection has a similar effect to enrichment culturing by amplifying
the population of interest - in this case methanotrophs [@swenson2000]. This will reduce the
diversity of soil microbes and allow for greater power in detecting significant
markers of methane oxidation rate. To do this, we extracted DNA from several of
the ecosystems within the selection experiment and sequenced the 16S rRNA gene
in order to perform a differential abundance test on methane oxidation rate.

# Materials & Methods

## Selection experiment
We performed an artificial ecosystem selection experiment [*sensu* @swenson2000]
by passaging replicate soil microbial communities. The trait we selected on was
CH~4~ oxidation rate. Our experiment had two selection lines with twelve jars
each for a total of twenty-four jars per passage. One line was a positive
selection line where the two or three jars with the highest CH~4~ oxidation rate
were chosen to inoculate the next set of positive jars. The other line was a
neutral selection line where an equal number of jars as the positive line were
chosen at random to inoculate the next set of neutral jars. The number of jars
chosen was based on the distribution of fluxes among the positive jars, i.e., we
chose the top three jars unless only two jars had considerably greater oxidation
rates based on visual inspection of histograms. The experiment was carried out
over five passages.

The initial soil microbial community was sampled from the top 10 cm of an upland
mineral soil under a deciduous forest ecosystem near the University of Oregon
campus in Eugene, OR, USA. Incubations were performed in 500 mL mason jars with
rubber septa installed in the lids. Each jar was sterilized with 70% ethanol to
which was added 45 g of autoclaved potting mix, 5 g of living soil, and 3.5 mL
of sterile deionized water to bring them to 60% of field capacity. The jars were
then homogenized, capped, and injected with 4.3 mL of 99% CH~4~ to bring the
headspace concentration to approximately 1000 ppm CH~4~. To create the two
treatment groups, twenty-four jars were created in an identical manner and then
randomly assigned to either the positive or neutral selection line. Jars were
flushed and respiked twice per week to maintain aerobic conditions and elevated
CH~4~ concentrations and were incubated at ambient temperature for approximately
three weeks. Methane oxidation rates were determined at the end of the
incubation period. These values were then used to determine the jars selected to
inoculate the next set of jars.

## Methane oxidation rate measurements

Methane oxidation rates were determined after flushing and spiking jars to
approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from each
jar immediately after spiking and then at time points 3, 6, 24, and 48 hours for
a 5-point curve. Samples were immediately injected into a SRI model 8610C gas
chromatograph (Torrance, CA, USA) to determine the headspace CH~4~
concentration. Fluxes were calculated from a first-order exponential decay
function as decay constant `k` with units day^-1^. Oxidation rates are presented
as the additive inverse of `k` (i.e., `-k`) so that a more positive value
represents a greater oxidation rate. The jars selected for the positive
treatment in passage 2 had the lowest methane oxidation rate of the twelve jars
due to a calculation error in the methane oxidation rate. All other passages
correctly used the top three jars.

## Soil DNA extraction and sequencing

A subsample of soil from the starting inoculum and from every jar in passages 2
and 5 was collected and stored at -80$^\circ$C. Soil DNA was extracted from 0.25
g soil. Negative controls were extracted from autoclaved potting mix and
DNase-free water. Extractions were performed using the DNeasy PowerSoil kit
(QIAGEN, Germany) and quantified using Qubit dsDNA HS Assay Kit (Thermo Fisher
Scientific, USA). In order to estimate the diversity and relative abundance of
the bacterial and archaeal taxa in our soil ecosystems, we sequenced the V4
region of the 16S rRNA gene using the using the 515F − 806R primer combination
[@caporaso2011]. PCR mixtures were: 10 $\mu$l NEBNext Q5 Hot Start HiFi PCR
master mix, 9.2 $\mu$l primer mixture (1.09 $\mu$M concentration), and 0.8
$\mu$l of DNA template. Reaction conditions were: 98$^\circ$C for 30 s
(initialization), 35 cycles of 98$^\circ$C for 10 s (denaturation), 61$^\circ$C
for 20 s (annealing), and 72$^\circ$C for 20 s (extension), and 72$^\circ$C for
2 m (final extension). Amplicons were purified twice using 0.8x ratio Mag-Bind
RxnPure Plus isolation beads (Omega Bio-Tek, USA). Sequencing libraries were
prepared using a dual-indexing approach (Fadrosh et al., 2014; Kozich et al.,
2013). Amplicon concentrations were quantified using Qubit (Thermo Fisher
Scientific Technologies, USA) multiplexed at equimolar concentration. Sequencing
was performed at the University of Oregon Genomics Core Facility on the Illumina
NovaSeq 6000 with paired-end 150 bp reads.

## Bioinformatics

Bioinformatics processing was performed in `R` (R Core Team 2021). Demultiplexed
sequencing reads were denoised using `DADA2` [@callahan2016]. Taxonomic
assignment for the 16S reads was performed using the RDP naive Bayesian
classifier [@wang2007].

## Statistical Analysis

### Response to selection

To test whether there was a significant change in CH~4~ oxidation rate (k) as a
response to selection, we tested a difference in slopes between the positive
and neutral selection lines. The CH~4~ oxidation rates were strongly right
skewed with most values close to zero and few large, positive values. This
resulted in residuals that did not meet the assumptions of constant variance
and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
transformed to better meet the assumptions of a linear model and to make
figures easier to interpret. First, we tested the effect of treatment by
fitting two nested models with and without treatment using `lm` in `R`. The
model formulation for the full model was `flux ~ passage * treatment` and the
model formulation for the reduced model was `flux ~ passage`. We compared these
models using the likelihood ratio test with the `anova` function in `R`. We
then fit the full model to estimate the slope of the positive line, which
represents the change in CH~4~ oxidation rate per passage as a response to
selection.

### Ecosystem heritability

To estimate the proportion of variance in ecosystem function due to variation in
the microbial community, we estimated narrow-sense heritability ($h^2$) as the
regression of offspring on mid-parent [@falconer1996]. The mid-parent was the
mean for all three selected jars and the offspring was the mean for all jars
produced by those parents. First, we tested if there was an effect of treatment
on the heritability estiamte. The model formulation for the full model was
`offspring ~ parent * treatment` and the model formulation for the reduced model
was `offspring ~ parent`. We compared these models using the likelihood ratio
test with the `anova` function `R`. We then fit the full model to estimate the
heritability of the Positive and Neutral treatments.

### Community response to selection

Diversity estimates and tests were performed using `DivNet` [@willis2020]. We
tested a difference in Shannon diversity between the Positive and Neutral
treatment and between the beginning and the end of the experiment with model
formulation `diversity ~ treatment * passage`. We also tested a difference in
beta-diversity as both Bray-Curtis distance and Jaccard distance between the
Positive and Neutral treatment and the beginning and end of the experiment with
model formulation `diversity ~ treatment * passage`.

To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
differential abundance using a beta-binomial model [@martin2020]. We compared
taxa in passage 5 between the positive and neutral treatment to identify taxa
that were enriched or depleted as a response to selection. We also tested
differential abundance with differences in CH~4~ oxidation rate to identify taxa
correlated with CH~4~ flux.

Possible tests:

- Random forest or gradient boosting to predict Positive vs. Neutral.

Test if random forest or gradient boosting can predict whether this is the positive line or the neutral line based solely on the microbial community?

# Results

## Methane oxidation

## Response to selection

We observed a response to selection on soil methane oxidation rate at the whole
ecosystem level (`r figures("response", display="cite")`). In addition, the
response to selection varied with treatment 
`r print_lrt(compare_response_mods, 2, 'LRT of nested models with and without treatment')`. At the start of the experiment, the Positive
treatment had a mean methane oxidation rate that was
`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`% lower than the
Neutral treatment `r print_lm(response_model, 3, 'difference of y-intercepts')`. 
There was no change in methane oxidation rate of the Neutral
treatment over the five passages `r print_lm(response_model, 2)`. By contrast,
the Positive treatment had a `r round(per_slope)`% increase in methane
oxidation rate per passage `r print_lm(response_model, 4)`.

```{r response_figure, fig.width=4, fig.height=3, fig.cap=figures("response")}
ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)

plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)

```

## Ecosystem heritability

We estimated heritability as the regression of mid-offspring on mid-parent.
Offspring CH~4~ oxidation rates were correlated with parental CH~4~ oxidation
rates in both the Positive treatment `r print_lm(herit_model, 4)` and the
Neutral treatment `r print_lm(herit_model, 2)`. Notably, the sign of the effect
was reversed with a positive heritability for the Positive treatment and a negative heritability for the Neutral treatment.

```{r herit_fig, fig.width=5, fig.height=3, fig.cap=figures("herit")}
plot_herit(heritability)
```

## Community Response

Alpha diversity
- treatment
- flux

Beta diversity
- treatment
- flux

Differential abundance
- treatment
- flux

# Discussion

We observed a response to selection on soil CH~4~ oxidation rate of the whole soil microbial community. This type of response has been observed for other functions, such as chloroaniline degradation in water [@swenson2000], but is the first, to our knowledge, of selection being performed on soil biogeochemical cycling.

A previous study near our sample site observed zero CH~4~ emissions from soils (cite laurel's paper). This suggests that little or no methanogenesis is happening in soils in this region and so the soil microbial communities are not "primed" to consume CH~4~. With no soil methanogenesis, the only type of methanotrophy happening would be high-affinity CH~4~ oxidation or consumption of atmospheric CH~4~, which is on the order of *some number* ppm CH~4~. By contrast, our incubations were at ~1000 ppm CH~4~. This is disparity is reflected in our results in that within the first passage we generated, only four of the 24 jars consumed more than 10% of the headspace CH~4~ within 48 hours and the biggest consumer only consumed 60% By contrast, in the final generation there were eight jars that consumed more than 10% and the top jar consumed over 95% of the headspace CH~4~.

If this experiment had been run for longer, we might be expected to see crashes in methane oxidation as the process of random sampling at each generation

Ecosystem heritability sounds confusing because it is difficult to imagine heritability without discrete generations. And how could a soil have "generations" (in the absence of coalescence events (cite Williams and Lenton))? However, this definition of heritability is conflated with the colloquial use of the term as referring to *inheritance*. But looking at the definition of narrow-sense heritability, it is simply the additive genetic variation associated with a trait [@falconer1996]. In this case, additive genetic effect would refer to the identity and relative abundance of microbial taxa in the metagenome. This notably excludes any interactions among taxa and interactions with the environment (i.e., GxE effects or context-dependent effects).

The heritability we observed was quite large at h^2^ = `r myround(tidy(herit_model)[4, 2], 2)`. However, given the small sample size and and wide confidence intervals (95% CI: `r myround(confint(herit_model)[4, ], 2)`) it is likely that a larger study could more precisely estimate the true heritability. Given these data, though, it would seem that genetic variation within the soil microbial community has a very strong effect on soil CH~4~ oxidation rates independent of any environmental variation. Granted, we intentionally eliminated much of the environmental variation that would be present in natural ecosystems.

The negative heritability of CH~4~ oxidation rate in the Neutral selection
treatment may result from selection on alternative characters in those
populations.

## Future Directions

This study focused on the practical implications of using AES as a tool to deduce how microbes map onto functions at the whole-ecosystem level. If one were interested in the evolutionary dynamics underlying this process, one could design a study to distinguish between individual selection on microbial strains and group selection or multilevel selection. That is, selection that cannot be recapitulated through selection at the individual level (Williams and Lenton 2007). This could be accomplished by culturing differentially abundant taxa in the positive selection treatment and then inoculating them in isolation (or in a non-selected background community) to see if the same level of ecosystem function could be recapitulated as in the final generation of the positive selection treatment.

Another avenue for research could be to look at correlated responses of ecosystem-level traits to selection. This would address questions such as, "Is methane oxidation rate correlated with other ecosystem processes due to pleiotropy or linkage disequilibrium?" At the community level, pleiotropy might be defined as "the production by a single taxon of two or more unrelated effects." Further, linkage disequilbrium would represent correlated traits due to taxa who tend to co-associate due to shared environmental preferences or coupled metabolic pathways.

These experiments would further help us understand the nature of the "map" between community structure and ecosystem function. This would not only reveal how best to predict ecosystem function in earth system models, but also demonstrate the mechanism underlying that relationship. This is akin to deducing the genetic control of trait variation in molecular biology.

Community epistasis is analogous to functional redundancy.

# Acknowledgments
This project was supported by the National Science Foundation Graduate Research Fellowship Program (grant no. DGE 1255832) and the ARCS Foundation Florence and Mike Nudelman Scholarship.

# Competing Interests
We declare we have no competing interests.

# Supplement

First, we found an effect of treatment by comparing the full model to the reduced model `r print_lrt(compare_response_mods)`.

# Junk

To identify communities of bacteria and archaea that collectively perform a
high rate of methane oxidation, we performed an artificial ecosystem selection
experiment  (Swenson et al. 2000).  We generated twenty-four soil ecosystems in the lab using a sterilized
potting mix as the substrate. We then inoculated these ecosystems with a small
amount of living soil to generate variation in community structure across the
twenty-four ecosystems. These ecosystems were then assigned to one of two
selection treatments: neutral selection or positive selection. Each ecosystem
was maintained at 1000 ppm CH~4~ over several weeks to allow for colonization and
growth of the microbial community and to enrich for methane oxidizers. We then
determined the methane oxidation rate for each ecosystem. To impose selection,
we selected three jars from the positive treatment with the highest rate of
methane oxidation, homogenized them, and used that soil to inoculate the next
set of jars. For the neutral treatment, we selected three jars at random to
inoculate the next set of jars. This process continued for five passages. At
the end of the experiment, we extracted DNA from soils in the second and fifth
passage for both the neutral and positive treatments. By comparing the response
to selection within the community between the two treatments, we can identify
microbial species that contribute to variation in ecosystem methane emissions.

This approach is powerful because by using a common soil substrate and
maintaining a constant headspace concentration of methane, we can eliminate
much of the environmental variation that would be present in an observational
study. In addition, by applying selection on methane oxidation rate
in the positive treatment, we can enrich for taxa involved in methane cycling
making it easier to perform association mapping between microbial taxa and
ecosystem methane emissions. In this paper, we address whether there is a
response to selection on methane oxidation rate at the whole ecosystem level. We
then quantify the amount of variation in methane oxidation attributable to
variation in community composition within our laboratory environment. Finally,
we identify microbial markers of soil methane cycling using association mapping
in order to deduce the mapping between microbial community structure and
ecosystem function.

(Goodnight et al. 1997, Williams and Lenton
2007)

# References