### 2022-04-07 R/08-community.Rmd

---
title: "Community Analysis"
author: "Andrew Morris"
date: "9/15/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir=normalizePath('../'))
```

```{r initialize_packages}
library(tidyverse)
library(phyloseq)
library(decontam)
library(microViz)
library(knitr)
library(corncob)
library(DivNet)
library(magrittr)
library(vegan)
library(ggpubr)
library(SpiecEasi)
library(ggthemes)
library(broom)

source('R/functions.R')
```

```{r ggplot-settings}
```

```{r import-data}
load(file = "Output/phyloseq.RData")
ps <-
  ps %>% 
  ps_mutate(pas_by_treat = paste0(passage, treat))
ps_melt <-
  ps %>% 
  ps_melt()
```

# Exploratory

```{r plot-depth}

depth_data <-
  ps %>% 
  ps_melt %>% 
  group_by(Sample, passage, treat) %>% 
  summarize(depth = sum(Abundance))

summary(aov(depth ~ passage * treat, data = depth_data))
depth_data %>% 
  ggplot(aes(x = interaction(treat, passage), y = depth)) +
  theme_tufte() +
  geom_rangeframe() +
  # coord_cartesian(clip="off") +
  geom_tufteboxplot()

```
There is no systematic difference in sequencing depth/library size between treatments. It does appear that sequencing depth decreases between passage 2 and 5.

```{r plot-richness}

plot_richness(ps, x = "pas_by_treat", color="treat", measures=c("Observed", "Shannon", "Simpson"), title = "Alpha Diversity ASV") +
    scale_color_manual(name = "Treatment", 
                         labels = c('Neutral', 'Positive'),
                         values = c('gray40', 'darkorange2'))

```

Richness decreases with passage, but seems unchanged between 5n and 5p.
Initial richness appears similar between the treatments.

```{r stacked_bar, eval=FALSE}
level = "Family"
ps_phylum <- speedyseq::tax_glom(ps, taxrank=level)
top10 = names(sort(taxa_sums(ps_phylum), TRUE)[1:10])
ps_phylum10 <- prune_taxa(top10, ps_phylum)
ps_phylum10 %>% 
  ps_filter(passage == 5) %>% 
  speedyseq::plot_bar(facet_grid = 'treat', fill = level) + 
  geom_bar(aes_string( color=level, fill=level), stat="identity", position="stack") +
  geom_bar(aes(x = reorder(rep, rank_ch4)))

```

# Hypothesis Testing

## Richness

```{r, eval=FALSE}
richness_soil <- ps %>% breakaway
plot(richness_soil, physeq=ps, color="passage")
summary(richness_soil) %>% as_tibble

meta <- ps %>%
  sample_data %>%
  as_tibble %>%
  mutate("sample_names" = ps %>% sample_names )

combined_richness <- meta %>%
  left_join(summary(richness_soil),
            by = "sample_names")

bt_passage <- betta(chats = combined_richness$estimate,
                      ses = combined_richness$error,
                      X = model.matrix(~passage, data = combined_richness))
bt_passage$table

save(bt_passage, file = 'Output/richness_model_passage.Rdata')

ps_5 <- ps_filter(ps, passage == "5")
richness_soil <- ps_5 %>% breakaway
plot(richness_soil, physeq=ps_5, color="treat")
summary(richness_soil) %>% as_tibble

meta <- ps_5 %>%
  sample_data %>%
  as_tibble %>%
  mutate("sample_names" = ps_5 %>% sample_names )

combined_richness <- meta %>%
  left_join(summary(richness_soil),
            by = "sample_names")

bt_treat5 <- betta(chats = combined_richness$estimate,
                      ses = combined_richness$error,
                      X = model.matrix(~treat, data = combined_richness))
bt_treat5$table

save(bt_treat5, file = 'Output/richness_model_treat5.Rdata')

bt_ch4 <- betta(chats = combined_richness$estimate,
                      ses = combined_richness$error,
                      X = model.matrix(~ch4, data = combined_richness))
bt_ch4$table

save(bt_ch4, file = 'Output/richness_model_ch4.Rdata')

```

Test confirms the above observation: Richness decreases between Passage 2
and Passage 5. No difference in richness by treatment in passage 5 or by
methane.

# Beta Diversity

```{r}

ps_fixed <-
ps %>%
 tax_fix(
  min_length = 4,
  unknowns = c(""),
  sep = " ", anon_unique = TRUE,
  suffix_rank = "classified"
 )

ps_fixed <- 
  ps_fixed %>% 
  ps_mutate(treat_bin = factor(ifelse(treat=='p', 1,0))) %>% 
  ps_mutate(passage_bin = factor(ifelse(passage==5, 1,0))) %>% 
  ps_mutate(treat_pas = interaction(treat_bin, passage_bin))

ps_fixed
```

## Beta diversity test passage by treatment

```{r}
plot_beta(ps_fixed)
```

```{r calc-beta, results=FALSE}

ord_calc_aitch <-
  ps_fixed %>% 
  tax_transform(trans = 'clr') %>% 
  ord_calc(method = "PCA")

ord_calc_bray <-
  ps_fixed %>%
  dist_calc("bray") %>%
  ord_calc(method = "NMDS")

ord_calc_jacc <-
  ps_fixed %>%
  dist_calc("jaccard", binary = TRUE) %>%
  ord_calc(method = "NMDS")

```

```{r plot-beta, results=FALSE, fig.width = 6, fig.height = 2}
bray <-
  ord_calc_bray %>% 
  ord_plot(color = "treat") +
  theme_tufte() +
  geom_rangeframe() +
  coord_cartesian(clip="off") +
  scale_color_manual(name = "Treatment", 
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2'))

jacc <- 
  ord_calc_jacc %>% 
  ord_plot(color = "treat") +
  theme_tufte() +
  geom_rangeframe() +
  coord_cartesian(clip="off") +
  scale_color_manual(name = "Treatment", 
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2'))

ggarrange(bray, jacc)

```

```{r plot-aitch, fig.width = 4, fig.height = 4}

ord_calc_aitch %>% 
  ord_plot(color = "treat",
           auto_caption = FALSE) +
  theme_tufte() +
  geom_rangeframe() +
  coord_fixed(5.5/43.6, clip="off") +
  scale_color_manual(name = "Treatment", 
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2'))

```

```{r permanova-aitch, eval=FALSE}
dist_aitch <-
  ps %>% 
  dist_calc(dist = "aitchison")

dist_permanova(dist_aitch, c("passage", "treat"), "passage * treat")

passage_groups <- factor(c(rep(1,22), rep(2,24)), labels = c("2","5"))

dist_aitch %>% 
  dist_get %>% 
  vegan::betadisper(passage_groups) %>% 
  permutest()

treat_groups <- factor(c(rep(1,12), rep(2,10), rep(1,12), rep(2,12)), labels = c("n","p"))

dist_aitch %>% 
  dist_get %>% 
  vegan::betadisper(treat_groups) %>% 
  permutest()

ps %>%
  ps_filter(passage == 2) %>% 
  dist_calc(dist = "aitchison") %>% 
  dist_permanova("treat")

ps %>%
  ps_filter(passage == 5) %>% 
  dist_calc(dist = "aitchison") %>% 
  dist_permanova("treat")

adonis2(dist_get(dist_aitch) ~ passage * treat, data = as(sample_data(ps_fixed), 'data.frame'))

dist_aitch2 <-
  ps %>% 
  ps_filter(passage == 2) %>% 
  dist_calc(dist = "aitchison")
dist_permanova(dist_aitch2, "treat")
groups <- factor(c(rep(1,12), rep(2,10)), labels = c("N","P"))

dist_aitch2 %>% 
  dist_get %>% 
  vegan::betadisper(groups) %>% 
  permutest()
dist_aitch5 <-
  ps %>% 
  ps_filter(passage == 5) %>% 
  dist_calc(dist = "aitchison")
dist_permanova(dist_aitch5, "treat")
groups <- factor(c(rep(1,12), rep(2,12)), labels = c("N","P"))

dist_aitch5 %>% 
  dist_get %>% 
  vegan::betadisper(groups) %>% 
  permutest()

```

There is a significant main effect and interaction of passage and treatment on
taxonomic dissimilarity of the soil microbiome. In addition, I reject the null
hypothesis of homogeneity of group dispersions for passage, but not for treatment.
Inspecting the NMDS plot, from this I conclude that passaging between Passage 2
and 5 both altered the soil microbiome taxonomic composition and reduced its 
variability. In addition, there is a difference in centroid between the two
treatments in both Passage 2 and Passage 5, but no difference in dispersion.

```{r beta-test}
# ps_prop <- transform_sample_counts(ps, function(otu) otu/sum(otu))
# 
# fit_bray <- adonis2(otu_table(ps_prop) ~ passage * treat, data = as(sample_data(ps_fixed), 'data.frame'), method = "bray")
# 
# fit_bray
# 
# fit_jacc <- adonis2(otu_table(ps_prop) ~ passage * treat, data = as(sample_data(ps_fixed), 'data.frame'), method = "jacc", binary = TRUE)
# 
# fit_jacc
ps_clr <- microbiome::transform(ps, "clr")

fit_bin_jacc <- adonis2(otu_table(ps) ~ passage * treat, data = as(sample_data(ps), 'data.frame'), method = "jaccard", binary = TRUE)

fit_aitch <- adonis2(otu_table(ps_clr) ~ passage * treat, data = as(sample_data(ps_clr), 'data.frame'), method = "euclidean")

fit_adonis <- rbind(tidy(fit_aitch), tidy(fit_bin_jacc))
fit_adonis$term <- c('Passage', 'Treatment', 'Pass:Treat', 'Residual', 'Total',
      'Passage', 'Treatment', 'Pass:Treat', 'Residual', 'Total')
save(fit_adonis, file = 'Output/beta_model.Rdata')

############

clr_dist_matrix <- phyloseq::distance(ps_clr, method = "euclidean")

dispr <- vegan::betadisper(clr_dist_matrix, phyloseq::sample_data(ps_clr)$passage)
disper_test <- permutest(dispr)

dispr <- vegan::betadisper(clr_dist_matrix, phyloseq::sample_data(ps_clr)$treat)
disper_test <- permutest(dispr)

dist_aitch5 <-
  ps %>% 
  ps_filter(passage == 5) %>% 
  dist_calc(dist = "aitchison")
dist_permanova(dist_aitch5, "treat")
groups <- factor(c(rep(1,12), rep(2,12)), labels = c("N","P"))

dist_aitch5 %>% 
  dist_get %>% 
  vegan::betadisper(groups) %>% 
  permutest()

dist_aitch2 <-
  ps %>% 
  ps_filter(passage == 2) %>% 
  dist_calc(dist = "aitchison")
dist_permanova(dist_aitch5, "treat")
groups <- factor(c(rep(1,12), rep(2,10)), labels = c("N","P"))

dist_aitch2 %>% 
  dist_get %>% 
  vegan::betadisper(groups) %>% 
  permutest()

###########3

# adonis2(otu_table(ps_filter(ps_prop, passage == 5)) ~ treat, 
#         data = 
#           as(
#             sample_data(
#               ps_filter(
#                 ps_prop, 
#                 passage == 5)
#             ), 
#             'data.frame'), 
#         method = "bray")
# adonis2(otu_table(ps_filter(ps_prop, passage == 5)) ~ treat, 
#         data = 
#           as(
#             sample_data(
#               ps_filter(
#                 ps_prop, 
#                 passage == 5)
#             ), 
#             'data.frame'), 
#         method = "jaccard",
#         binary = TRUE)

```

# Differential Abundance

## ALDex2

```{r}
library(ALDEx2)
data(selex)
#subset only the last 400 features for efficiency
selex.sub <- selex[1:400,]
conds <- c(rep("NS", 7), rep("S", 7))
x <- aldex.clr(selex.sub, conds, mc.samples=16, denom="iqlr", verbose=F)
x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
x.effect <- aldex.effect(x, CI=T, verbose=FALSE)
head(x.effect)

```

```{r}
ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")

ps_fam_pas5 <- microViz::ps_filter(ps_fixed_family, passage == 5)

ps_fam_pas5_prev <- microViz::tax_filter(ps_fam_pas5, min_prevalence = 0.10)
ps_pass_5 <- microViz::ps_filter(ps_fixed, passage == 5)
otu_table <- as(t(otu_table(ps_fam_pas5_prev)), 'matrix')
conds <- c(rep("n", 12), rep("p", 12))

# Based on effect confidence interval

x <- aldex.clr(otu_table, conds, mc.samples=128, denom="all", verbose=F)
x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
x.effect <- aldex.effect(x, CI=T, verbose=FALSE)

sgn <- sign(x.effect$effect.low) == sign(x.effect$effect.high)
par(mfrow=c(1,2))
plot(x.effect$rab.all, x.effect$diff.btw, pch=19, cex=0.3, col="grey", xlab="Abundance", ylab="Difference", main="Bland-Altman")
points(x.effect$rab.all[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
points(x.effect$rab.all[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")

plot(x.effect$diff.win, x.effect$diff.btw, pch=19, cex=0.3, col="grey",xlab="Dispersion", ylab="Difference", main="Effect")
points(x.effect$diff.win[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
points(x.effect$diff.win[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")
abline(0,2, lty=2, col="grey")
abline(0,-2, lty=2, col="grey")

merge(as(tax_table(ps_fam_pas5_prev)[rownames(x.effect[sgn,]), ], 'data.frame'), x.effect[sgn,], by = 0)
```

```{r}

# Based on effect size

x.all <- aldex(otu_table, conds, mc.samples=128, test="t", effect=TRUE,
     include.sample.summary=FALSE, denom="all", verbose=FALSE)
effect_gt_2 <- merge(as(tax_table(ps_pass_5)[rownames(x.all[abs(x.all$effect)>2,]), ], 'data.frame'), x.all[abs(x.all$effect)>2,], by = 0)

effect_gt_2 %>% 
  ggplot(aes(x = effect, y = reorder(Family,effect))) +
  geom_point()

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
    ylab="Difference")
aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
    ylab="Difference")

par(mfrow=c(1,2))
plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")
legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))

plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")

# identify which values are significant in both the t-test and glm tests
found.by.all <- which(x.all$we.eBH < 0.05 & x.all$wi.eBH < 0.05)

# identify which values are significant in fewer than all tests
found.by.one <- which(x.all$we.eBH < 0.05 | x.all$wi.eBH < 0.05)

# plot the within and between variation of the data
plot(x.all$diff.win, x.all$diff.btw, pch=19, cex=0.3, col=rgb(0,0,0,0.3),
 xlab="Dispersion", ylab="Difference")
points(x.all$diff.win[found.by.one], x.all$diff.btw[found.by.one], pch=19,
 cex=0.7, col=rgb(0,0,1,0.5))
points(x.all$diff.win[found.by.all], x.all$diff.btw[found.by.all], pch=19,
 cex=0.7, col=rgb(1,0,0,1))
abline(0,1,lty=2)
abline(0,-1,lty=2)

#View(tax_table(ps_fam_pas5_prev)[rownames(otu_table[found.by.all, ]), ])
```

```{r, eval=FALSE}
boxData <-
  ps_fixed %>% 
  ps_filter(passage ==5) %>% 
  tax_agg('Family') %>% 
  ps_get %>% 
  ps_melt()

Family_order<-
  boxData %>% 
  group_by(Family) %>% 
  summarize(median = median(Abundance)) %>% 
  arrange(median)
boxData$Family <- factor(boxData$Family, levels = Family_order$Family)
boxData$treat <- factor(boxData$treat, levels = c('p', 'n'))
```

```{r, eval=FALSE, fig.height=3, fig.width=12}

boxData %>% 
  ggplot(aes(x = Family, 
             y = Abundance)) +
  facet_grid(treat ~ .) +
  theme_tufte() +
  geom_rangeframe(color = "black") +
  # scale_x_reordered() +
  scale_y_continuous(trans = "pseudo_log") +
  geom_tufteboxplot() +
  theme(axis.text.x = element_text(angle = 90))
  # geom_boxplot() +
  # scale_color_discrete(guide = 'none') +
  # coord_flip()

```

```{r, eval=FALSE}
g5p<- 
  ps_fixed %>% 
  ps_filter(treat == "p", passage == 5) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()
g5n<- 
  ps_fixed %>% 
  ps_filter(treat == "n", passage == 5) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()

# Families unique to p in passage 5
g5p_unique <-
  g5p %>% 
  filter(Family %in% unique(g5p$Family)[!(unique(g5p$Family) %in% unique(g5n$Family))]) %>% 
  mutate(Prevalence = as.numeric(Abundance > 0))

# Abundance
Family_order<-
  g5p_unique %>% 
  group_by(OTU) %>% 
  summarize(median = median(Abundance), max = max(Abundance)) %>% 
  arrange(median, max)
g5p_unique$OTU <- factor(g5p_unique$OTU, levels = Family_order$OTU)

g5p_unique %>% 
  ggplot(aes(y = Abundance, x = OTU)) +
  geom_tufteboxplot() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 24))

# Prevalence
Family_order<-
  g5p_unique %>% 
  group_by(OTU) %>% 
  summarize(sum = sum(Prevalence)) %>% 
  arrange(sum)
g5p_unique$OTU <- factor(g5p_unique$OTU, levels = Family_order$OTU)

g5p_unique %>% 
  group_by(OTU) %>% 
  summarize(Prevalence = sum(Prevalence)) %>% 
  ggplot(aes(y = Prevalence, x = OTU)) +
  geom_point() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(1, 2, 3, 5, 10, 12))

# Families unique to n in passage 5
g5n_unique <-
  g5n %>% 
  filter(Family %in% unique(g5n$Family)[!(unique(g5n$Family) %in% unique(g5p$Family))]) %>% 
  mutate(Prevalence = as.numeric(Abundance > 0))

# Abundance
Family_order<-
  g5n_unique %>% 
  group_by(OTU) %>% 
  summarize(median = median(Abundance), max = max(Abundance)) %>% 
  arrange(median, max)
g5n_unique$OTU <- factor(g5n_unique$OTU, levels = Family_order$OTU)

g5n_unique %>% 
  ggplot(aes(y = Abundance, x = OTU)) +
  geom_tufteboxplot() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(0, 10, 20, 30, 37))

# Prevalence
Family_order<-
  g5n_unique %>% 
  group_by(OTU) %>% 
  summarize(sum = sum(Prevalence)) %>% 
  arrange(sum)
g5n_unique$OTU <- factor(g5n_unique$OTU, levels = Family_order$OTU)

g5n_unique %>% 
  group_by(OTU) %>% 
  summarize(Prevalence = sum(Prevalence)) %>% 
  ggplot(aes(y = Prevalence, x = OTU)) +
  geom_point() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(1, 2, 6))
```

First, I looked at which taxa are unique to treatments N and P. Several families
present in P and absent in N. Most of them had low prevalance (present in 3 or
fewer samples) and low abundance (median < 5 reads across samples). There were two families unique
to the Positive treatment with relatively high prevalence. This included an ASV
that was a member of the Bacteroidia Class with no lower taxonomic designation. 
This ASV had a prevalence of 10/12 and a median abundance of ~2 reads. The other
prevalent family was a member of the Silvanigrellaceae, a newly described family places in
its own order. This family was present in
all 12 samples and was represented by three OTUs: one with no lower designation, one from the genus
Silvanigrella and one from the genus Candidatus Turbobacter. Silvanigrella was
isolated from a temperate fresh water lake (https://www.microbiologyresearch.org/content/journal/ijsem/10.1099/ijsem.0.001965).

Of the families unique to the Neutral treatment, only one had a prevalence greater than 2/12. 
This family, Armatimonadaceae, was present in half of the Neutral samples (6/12). This
family was represented by two ASVs in the genus Armatimonas. The type strain for
Armatimonas was isolated from the rhizoplane of Phragmites australis (https://www.microbiologyresearch.org/content/journal/ijsem/10.1099/ijs.0.025643-0)

```{r, eval=FALSE}

g5<- 
  ps_fixed %>% 
  ps_filter(passage == 5) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()

g2<- 
  ps_fixed %>% 
  ps_filter(passage == 2) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()

unique(g5$Family)[!(unique(g5$Family) %in% unique(g2$Family))]
unique(g5$Family)[unique(g5$Family) %in% unique(g2$Family)]

unique(g2$Family)[!(unique(g2$Family) %in% unique(g5$Family))]
unique(g2$Family)[unique(g2$Family) %in% unique(g5$Family)]

```

## Differential Abundance Treatment in passage 5

```{r corncob_treatment, fig.hight=40}
ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")

ps_fam_pas5 <- microViz::ps_filter(ps_fixed_family, passage == 5)

ps_fam_pas5_prev <- microViz::tax_filter(ps_fam_pas5, min_prevalence = 0.10)

sample_data(ps_fam_pas5_prev) =
  sample_data(ps_fam_pas5_prev)[,c('sample_id', 'treat')] %>% 
  sample_data()
ps = ps_fam_pas5_prev

save(ps, file = 'ps.RData')
# Differential abundance test
corncob <- bbdml(formula = ASV2954 ~ treat,
                                 phi.formula = ~ 1,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ 1,
                                 test = "lrt", boot = TRUE,
                                 data = ps_fam_pas5_prev,
                                 fdr_cutoff = 0.05)
da_analysis <- differentialTest(formula = ~ treat,
                                 phi.formula = ~ 1,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ 1,
                                 test = "Wald", boot = FALSE,
                                 data = ps_fam_pas5_prev,
                                 fdr_cutoff = 0.05, B = 50)
plot(da_analysis, level = c("Family"))
da_analysis$significant_models
dt_data <- plot(da_analysis, level = c("Family"), data_only = TRUE)

sig_taxa_treat <- tax_select(ps_fam_pas5_prev, da_analysis$significant_taxa)
ntaxa(sig_taxa_treat)
save(da_analysis, sig_taxa_treat, file = 'Output/da_corncob.Rdata')

save(dt_data, file = 'Output/dt_data.Rdata')

```

```{r corncob-plot, fig.width=6, fig.height=6}
dt_plot(dt_data)

```

## Faprotax on DA taxa

```{r export_to_faprotax, eval=FALSE}
faprotax_taxonomy <-
  sig_taxa_treat %>% 
  tax_table %>% 
  as_tibble %>% 
  mutate(.taxonomy = 
           faprotax_taxonomy %>% 
           select(Kingdom:Species) %>% 
           apply(1 , paste , collapse = "; " )) %>% 
  select(.otu, .taxonomy)

otu<-t(as(otu_table(sig_taxa_treat),"matrix")) 
  # 't' to transform if taxa_are_rows=FALSE
  #if taxa_are_rows=TRUE
  #otu<-as(otu_table(GlobalPatterns),"matrix"))
rownames(otu) <- faprotax_taxonomy$.taxonomy
otu_biom <- biomformat::make_biom(data=otu)
biomformat::write_biom(otu_biom,"otu_biom.biom")

```

```{bash faprotax, eval=FALSE}
conda activate faprotax

rm Output/func_table_from_biom.biom \
    Output/faprotax_report_from_biom.txt

python python/FAPROTAX_1.2.4/collapse_table.py \
    -i otu_biom.biom \
    -g python/FAPROTAX_1.2.4/FAPROTAX.txt \
    -o Output/func_table_from_biom.biom \
    -r Output/faprotax_report_from_biom.txt \
    -v

biom convert \
    -i Output/func_table_from_biom.biom \
    -o Output/func_table_from_biom_json.biom \
    --table-type="OTU table" \
    --to-json

conda deactivate
```

```{r}
biom_file <- "Output/func_table_from_biom_json.biom"

ps_faprotax <-
  biom_file %>% 
  biomformat::read_biom() %>% 
  biomformat::biom_data() %>% 
  as("matrix") %>% 
  otu_table(taxa_are_rows=TRUE) %>% 
  phyloseq(sample_data(ps_gen_pas5_prev))

da_faprotax <- differentialTest(formula = ~ treat,
                                 phi.formula = ~ treat,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ treat,
                                 test = "Wald", boot = FALSE,
                                 data = ps_faprotax,
                                 fdr_cutoff = 0.05)
plot(da_faprotax)

```

## Picrust on DA taxa

```{r}
x = NULL
for (i in seq_along(da_analysis$significant_models)) {
  x <- c(x, da_analysis$significant_models[[i]]$coefficients[2, 1])
}
x > 0
sig_taxa_pos_neg <-
  da_analysis$significant_taxa %>% 
  as_tibble_col(column_name = ".otu") %>% 
  mutate(estimate = x) %>% 
  mutate(pos_or_neg = if_else(x > 0, "pos", "neg"))

top_da_taxa <- 
  sig_taxa_treat %>% 
  otu_table %>% 
  as_tibble %>% 
  filter(.abundance > 2000) %>% 
  distinct(.otu) %$% .otu

taxonomy <-
  sig_taxa_treat %>% 
  tax_table %>% 
  as_tibble %>%
  filter(.otu %in% top_da_taxa)

taxonomy$.taxonomy <- taxonomy %>% 
           select(Kingdom:Species) %>% 
           apply(1 , paste , collapse = "; " )

taxonomy <-
  taxonomy %>% 
  select(.otu, .taxonomy)

df_top_da_taxa <-
  sig_taxa_treat %>% 
  otu_table %>% 
  as_tibble %>%
  left_join(as_tibble(sample_data(sig_taxa_treat))) %>% 
  left_join(sig_taxa_pos_neg, by = ".otu") %>% 
  filter(.otu != 'ASV4174') %>% 
  filter(.otu %in% top_da_taxa) %>% 
  left_join(taxonomy)

df_top_da_taxa %>% 
  ggplot(., aes(y = Genus, x = .abundance, color = treat)) +
  geom_jitter(height = 0.1) + 
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~ pos_or_neg)
ggsave('Output/dif_abund_results.png', width = 6, height = 4)

major_otus <- 
sig_taxa_treat %>% 
  otu_table %>% 
  as_tibble %>%
  left_join(as_tibble(sample_data(sig_taxa_treat))) %>% 
  left_join(sig_taxa_pos_neg, by = ".otu") %>% 
  filter(.otu != 'ASV4174') %>% 
  filter(.abundance > 300) %>% 
  select(.otu, pos_or_neg) %>% 
  distinct %>% 
  mutate(tax = otu_to_taxonomy(.otu, sig_taxa_treat))

otu_to_taxonomy(major_otus$.otu, sig_taxa_treat)

otu_to_taxonomy('ASV192', sig_taxa_treat)

sig_taxa_treat %>% 
  otu_table %>% 
  as_tibble %>%
  left_join(as_tibble(sample_data(sig_taxa_treat))) %>% 
  left_join(sig_taxa_pos_neg, by = ".otu") %>% 
  filter(.otu == 'ASV192') %>% 
  ggplot(., aes(x = .otu, y = .abundance, color = treat)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~ pos_or_neg)

beijer_otu <-
  ps %>% 
  otu_table %>% 
  as_tibble %>% 
  filter(.otu == 'ASV192')

beijer_metadata <-
  ps %>% 
  sample_data %>% 
  as_tibble %>% 
    left_join(beijer_otu)

ggplot(beijer_metadata, aes(x = .abundance, y = log_ch4, color = treat)) +
  geom_point() +
  facet_wrap(~ passage)

df_top_da_taxa %>% 
filter(pos_or_neg == "pos") %>% 
select(Phylum:Genus) %>% 
distinct %>% 
  write_tsv(., 'top_positive_taxa.txt')
```

```{bash, eval=FALSE}
mkdir picrust2
```

```{r export_to_picrust}

top_da_taxa_pos <-
  df_top_da_taxa %>% 
  select(.otu, pos_or_neg) %>% 
  distinct(.otu, pos_or_neg) %>% 
  filter(pos_or_neg == "pos") %$%
  .otu
sig_taxa_treat %>% 
  tax_select(top_da_taxa_pos) %>% 
  otu_table %>% 
  as("matrix") %>% 
  t %>% 
  biomformat::make_biom() %>% 
  biomformat::write_biom("Output/study_seqs.biom")

sequences <-
  sig_taxa_treat %>% 
  tax_select(top_da_taxa_pos) %>% 
  refseq %>% 
  as.vector %>% 
  as.list
seqinr::write.fasta(sequences, names(sequences), file.out = "Output/study_seqs.fna")

```

```{bash picrust, eval=FALSE}
conda activate picrust2
cd picrust3c

place_seqs.py -s ../Output/study_seqs.fna \
    -o out.tre \
    -p 1 \
    --intermediate intermediate/place_seqs \
    --chunk_size 500
hsp.py -i 16S -t out.tre -o marker_predicted_and_nsti.tsv.gz -p 1 -n

hsp.py -i EC -t out.tre -o EC_predicted.tsv.gz -p 1

metagenome_pipeline.py -i ../Output/study_seqs.biom -m marker_predicted_and_nsti.tsv.gz -f EC_predicted.tsv.gz \
                       -o EC_metagenome_out --strat_out

pathway_pipeline.py -i EC_metagenome_out/pred_metagenome_contrib.tsv.gz \
                    -o pathways_out -p 1

add_descriptions.py -i EC_metagenome_out/pred_metagenome_unstrat.tsv.gz -m EC \
                    -o EC_metagenome_out/pred_metagenome_unstrat_descrip.tsv.gz

add_descriptions.py -i pathways_out/path_abun_unstrat.tsv.gz -m METACYC \
                    -o pathways_out/path_abun_unstrat_descrip.tsv.gz

#####

picrust2_pipeline.py -s Output/study_seqs.fna -i Output/study_seqs.biom -o picrust2_out_pipeline -p 4 --stratified

conda deactivate
```

```{r}
se.mb.amgut <- spiec.easi(amgut1.filt, method='mb', lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=50))
se.gl.amgut <- spiec.easi(amgut1.filt, method='glasso', lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=50))
sparcc.amgut <- sparcc(amgut1.filt)
## Define arbitrary threshold for SparCC correlation matrix for the graph
sparcc.graph <- abs(sparcc.amgut$Cor) >= 0.3
diag(sparcc.graph) <- 0
library(Matrix)
sparcc.graph <- Matrix(sparcc.graph, sparse=TRUE)
## Create igraph objects
ig.mb     <- adj2igraph(getRefit(se.mb.amgut))
ig.gl     <- adj2igraph(getRefit(se.gl.amgut))
ig.sparcc <- adj2igraph(sparcc.graph)

```

```{r}
library(igraph)
## set size of vertex proportional to clr-mean
vsize    <- rowMeans(clr(amgut1.filt, 1))+6
am.coord <- layout.fruchterman.reingold(ig.mb)

par(mfrow=c(1,3))
plot(ig.mb, layout=am.coord, vertex.size=vsize, vertex.label=NA, main="MB")
plot(ig.gl, layout=am.coord, vertex.size=vsize, vertex.label=NA, main="glasso")
plot(ig.sparcc, layout=am.coord, vertex.size=vsize, vertex.label=NA, main="sparcc")

```

```{r}

library(phyloseq)
## Load round 2 of American gut project
data('amgut2.filt.phy')
se.mb.amgut2 <- spiec.easi(amgut2.filt.phy, method='mb', lambda.min.ratio=1e-2,
                           nlambda=20, pulsar.params=list(rep.num=50))
ig2.mb <- adj2igraph(getRefit(se.mb.amgut2),  vertex.attr=list(name=taxa_names(amgut2.filt.phy)))
plot_network(ig2.mb, amgut2.filt.phy, type='taxa', color="Rank3")

ps_p <-
  test %>% 
  speedyseq::tax_glom("Family") %>% 
  ps_filter(passage == "5") %>% 
  ps_filter(treat == "p")

se.mb.ps_p <- spiec.easi(ps_p, method='mb', lambda.min.ratio=1e-2,
                           nlambda=200, pulsar.params=list(rep.num=50))
ig2.mb.ps_p <- adj2igraph(getRefit(se.mb.ps_p),  vertex.attr=list(name=taxa_names(ps_p)))
plot_network(ig2.mb.ps_p, ps_p, type='taxa', color = "Class")

ps_n <-
  test %>% 
  speedyseq::tax_glom("Family") %>% 
  ps_filter(passage == "5") %>% 
  ps_filter(treat == "n")

se.mb.ps_n <- spiec.easi(ps_n, method='mb', lambda.min.ratio=1e-2,
                           nlambda=200, pulsar.params=list(rep.num=50))
ig2.mb.ps_n <- adj2igraph(getRefit(se.mb.ps_n),  vertex.attr=list(name=taxa_names(ps_n)))
plot_network(ig2.mb.ps_n, ps_n, type='taxa', color="Class")
getOptMerge(se.mb.ps_p)
getOptMerge(se.mb.ps_n)
getStability(se.mb.ps_n)
###

se1.mb.amgut <- spiec.easi(amgut1.filt, method='mb', lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))
se2.mb.amgut <- spiec.easi(amgut2.filt.phy, method='mb', lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))

se1.slr.amgut <- spiec.easi(amgut1.filt, method='slr', r=10, lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))
se2.slr.amgut <- spiec.easi(amgut2.filt.phy, method='slr', r=10, lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))

otu1 <- colnames(amgut1.filt)
otu2 <- taxa_names(amgut2.filt.phy)
edge.diss(getRefit(se1.mb.amgut), getRefit(se2.mb.amgut), 'jaccard', otu1, otu2)
edge.diss(getRefit(se1.slr.amgut), getRefit(se2.slr.amgut), 'jaccard', otu1, otu2)

X <- se2.slr.amgut$est$data
L <- se2.slr.amgut$est$resid[[getOptInd(se2.slr.amgut)]]
pca <- robustPCA(X, L)
age <- as.numeric(as.character(sample_data(amgut2.filt.phy)$AGE))
bmi <- as.numeric(as.character(sample_data(amgut2.filt.phy)$BMI))
depth <- colSums(otu_table(amgut2.filt.phy))

cor(age, pca$scores, use='pairwise')
cor(bmi, pca$scores, use='pairwise')
cor(depth, pca$scores, use='pairwise')

```

## Differential Abundance Methane

```{r corncob_ch4}

ps_gen_pas5_trep <- ps_filter(ps_gen_pas5, treat == "p")

ps_gen_prev_trep_prev <- tax_filter(ps_gen_pas5_trep, min_prevalence = 0.10)

da_analysis <- differentialTest(formula = ~ ch4,
                                 phi.formula = ~ ch4,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ ch4,
                                 test = "Wald", boot = FALSE,
                                 data = ps_gen_prev_trep_prev,
                                 fdr_cutoff = 0.05)
# otu_to_taxonomy(OTU = da_analysis$significant_taxa, data = ps_gen_prev)

plot(da_analysis, level = c("Family", "Genus"))

# Check unfitted taxa
# otu_to_taxonomy(which(is.na(da_analysis$p)) %>% names, data = ps_gen_prev)
# otu_table(ps_gen_prev)[,which(is.na(da_analysis$p))]

# Differential variability test

dv_analysis <- differentialTest(formula = ~ passage + treat + ch4,
                                 phi.formula = ~ passage + treat + ch4,
                                 formula_null = ~ passage + treat + ch4,
                                 phi.formula_null = ~ passage + treat,
                                 data = ps_gen_prev,
                                 test = "Wald", boot = FALSE,
                                 fdr_cutoff = 0.05)
# otu_to_taxonomy(OTU = dv_analysis$significant_taxa, data = ps_gen_prev)
plot(dv_analysis, level = c("Family", "Genus"))

# Check unfitted taxa

# otu_to_taxonomy(which(is.na(dv_analysis$p)) %>% names, data = ps_gen_prev)
# otu_table(ps_gen_prev)[,which(is.na(dv_analysis$p))]

```

## Alpha Diversity Methane

```{r diversity_ch4, eval = FALSE}
ps_phylum <- speedyseq::tax_glom(ps, taxrank="Phylum")

divnet_phylum_ch4 <- ps_phylum %>%
  divnet(ncores = 4, tuning = "careful", formula = ~ rank_ch4)

divnet_phylum_ch4$shannon %>% 
  plot(ps_phylum, color = "rank_ch4") +
  xlab("Rank Methane") +
  ylab("Shannon diversity estimate\n(phylum level)") #+
  #coord_cartesian(ylim = c(0,2))

estimates <- divnet_phylum_ch4$shannon %>% summary %$% estimate
ses <- sqrt(divnet_phylum_ch4$`shannon-variance`)
X <- breakaway::make_design_matrix(ps_phylum, "rank_ch4")
betta(estimates, ses, X)$table

betta(estimates, ses, X)$global[2]

```

## Alpha Diversity Treatment in Passage 2

```{r, results=FALSE, eval = FALSE}
ps_class <- speedyseq::tax_glom(ps, taxrank="Class")
ps_class

ps_c_2 <- ps_filter(ps_class, passage == 2)

divnet_c_2 <- ps_c_2 %>%
  divnet(ncores = 4, tuning = "careful", formula = ~ treat)
```

```{r, eval = FALSE}
divnet_c_2$shannon %>% 
  plot(ps_c_2, color = "treat") +
  xlab("Selection treatment") +
  ylab("Shannon diversity estimate\n(class level)") +
  labs(title="Test alpha diversity between P and N treatments within Passage 2")

estimates <- divnet_c_2$shannon %>% summary %$% estimate
ses <- sqrt(divnet_c_2$`shannon-variance`)
X <- breakaway::make_design_matrix(ps_c_2, "treat")
betta(estimates, ses, X)$table
betta(estimates, ses, X)$global

divnet_c_2$simpson %>% 
  plot(ps_c_2, color = "treat") +
  xlab("Selection treatment") +
  ylab("Simpson diversity estimate\n(class level)") +
  labs(title="Test alpha diversity between P and N treatments within Passage 2")

estimates <- divnet_c_2$simpson %>% summary %$% estimate
ses <- sqrt(divnet_c_2$`simpson-variance`)
X <- breakaway::make_design_matrix(ps_c_2, "treat")
betta(estimates, ses, X)$table
betta(estimates, ses, X)$global

```

## Alpha Diversity Passage 2 vs 5 (all treatment)

```{r, results=FALSE, eval=FALSE}
ps_phylum <- speedyseq::tax_glom(ps, taxrank="Phylum")

divnet_p <- ps_phylum %>%
  divnet(ncores = 4, tuning = "careful", formula = ~ passage)
```

```{r, eval = FALSE}
divnet_p$shannon %>% 
  plot(ps_phylum, color = "passage") +
  xlab("Selection treatment") +
  ylab("Shannon diversity estimate\n(phylum level)") +
  labs(title="Test alpha diversity between Passage 2 and 5 all treatments")

estimates <- divnet_p$shannon %>% summary %$% estimate
ses <- sqrt(divnet_p$`shannon-variance`)
X <- breakaway::make_design_matrix(ps_phylum, "passage")
betta(estimates, ses, X)$table
betta(estimates, ses, X)$global[2]

divnet_p$simpson %>% 
  plot(ps_phylum, color = "passage") +
  xlab("Selection treatment") +
  ylab("Simpson diversity estimate\n(phylum level)") +
  labs(title="Test alpha diversity between Passage 2 and 5 all treatments")

estimates <- divnet_p$simpson %>% summary %$% estimate
ses <- sqrt(divnet_p$`simpson-variance`)
X <- breakaway::make_design_matrix(ps_phylum, "passage")
betta(estimates, ses, X)$table
betta(estimates, ses, X)$global[2]

```

## Alpha Diversity Treatment in Passage 5

```{r, results=FALSE, eval = FALSE}

ps_class <- speedyseq::tax_glom(ps, taxrank="Class")

ps_c_5 <- ps_filter(ps_class, passage == 5)

ps_5 <- ps_filter(ps_phylum, passage == 5)

divnet_p_5 <- ps_5 %>%
  divnet(ncores = 4, tuning = "careful", formula = ~ treat)
```

```{r, eval = FALSE}
divnet_p_5$shannon %>% 
  plot(ps_5, color = "treat") +
  xlab("Selection treatment") +
  ylab("Shannon diversity estimate\n(phylum level)") +
  labs(title="Test alpha diversity between P and N treatments within Passage 5")

estimates <- divnet_p_5$shannon %>% summary %$% estimate
ses <- sqrt(divnet_p_5$`shannon-variance`)
X <- breakaway::make_design_matrix(ps_5, "treat")
betta(estimates, ses, X)$table
betta(estimates, ses, X)$global
#betta_pic(estimates, ses)

divnet_p_5$simpson %>% 
  plot(ps_5, color = "treat") +
  xlab("Selection treatment") +
  ylab("Simpson diversity estimate\n(phylum level)") +
  labs(title="Test alpha diversity between P and N treatments within Passage 5")

estimates <- divnet_p_5$simpson %>% summary %$% estimate
ses <- sqrt(divnet_p_5$`simpson-variance`)
X <- breakaway::make_design_matrix(ps_5, "treat")
betta(estimates, ses, X)$table
betta(estimates, ses, X)$global
#betta_pic(estimates, ses)

```

## Beta Diversity passage + treatment

```{r, eval=FALSE}

ps_f_c <- speedyseq::tax_glom(ps_fixed, "Class")

X <- diag(nrow(sample_data(ps_f_c)))
colnames(X) <- rownames(X) <-  as.character(1:ncol(X))

divnet_class <-  divnet(W = ps_f_c,
                         X = X,
                         variance = "none")

# Median composition
set.seed(454)
bc_test <- testBetaDiversity(divnet_class, 
                             h0 = "bray-curtis",
                             n_boot = 1000, # set this to 10000
                             sample_specimen_matrix = divnet_class$X,
                             groups = sample_data(ps_f_c)$treat_pas)
bc_test$p_value

# Euclidean distance
set.seed(3532)
euc_test <- testBetaDiversity(dv = divnet_class , 
                              h0 = "euclidean",
                              n_boot = 1000,
                              sample_specimen_matrix = divnet_class$X,
                              groups = sample_data(ps_f_c)$treat_pas)
euc_test$p_value

# Aitchison Distance
set.seed(3423)
ait_test <- testBetaDiversity(dv = divnet_class, 
                              sample_specimen_matrix = divnet_class$X,
                              h0 = "aitchison",
                              n_boot = 1000,
                    groups = sample_data(ps_f_c)$treat_pas)

ait_test$p_value
```

## Beta Diversity gen 2 treatment

```{r, results=FALSE, eval = FALSE}
ps_f_2 <- ps_filter(ps_fixed, passage == "2")
ps_f_2_c <- speedyseq::tax_glom(ps_f_2, "Class")
X <- diag(nrow(sample_data(ps_f_2_c)))
colnames(X) <- rownames(X) <-  as.character(1:ncol(X))

divnet_c_2 <-  divnet(W = ps_f_2_c,
                         X = X,
                         variance = "none")

```

```{r, eval = FALSE}

# Median composition
set.seed(454)
bc_test <- testBetaDiversity(dv = divnet_c_2, 
                             h0 = "bray-curtis",
                             n_boot = 1000, # set this to 10000
                             sample_specimen_matrix = divnet_c_2$X,
                             groups = sample_data(ps_f_2_c)$treat_bin)
bc_test$p_value
set.seed(3532)
# Euclidean distance
euc_test <- testBetaDiversity(dv = divnet_c_2 , 
                              h0 = "euclidean",
                              n_boot = 1000,
                              sample_specimen_matrix = divnet_c_2$X,
                              groups = sample_data(ps_f_2_c)$treat_bin)
euc_test$p_value

# Aitchison Distance
set.seed(3423)
ait_test <- testBetaDiversity(dv = divnet_c_2, 
                              sample_specimen_matrix = divnet_c_2$X,
                              h0 = "aitchison",
                              n_boot = 1000,
                              groups = sample_data(ps_f_2_c)$treat_bin)

ait_test$p_value

```

## Beta Diversity gen 5 treatment

```{r, eval = FALSE}
ps_f_5 <- ps_filter(ps_fixed, passage == "5")
ps_f_5_c <- speedyseq::tax_glom(ps_f_5, "Class")
X <- diag(nrow(sample_data(ps_f_5_c)))
colnames(X) <- rownames(X) <-  as.character(1:ncol(X))

divnet_c_5 <-  divnet(W = ps_f_5_c,
                         X = X,
                         variance = "none")

# Median composition
set.seed(454)
bc_test <- testBetaDiversity(dv = divnet_c_5, 
                             h0 = "bray-curtis",
                             n_boot = 1000, # set this to 10000
                             sample_specimen_matrix = divnet_c_5$X,
                             groups = sample_data(ps_f_5_c)$treat_bin)
bc_test$p_value
set.seed(3532)
# Euclidean distance
euc_test <- testBetaDiversity(dv = divnet_c_5 , 
                              h0 = "euclidean",
                              n_boot = 1000,
                              sample_specimen_matrix = divnet_c_5$X,
                              groups = sample_data(ps_f_5_c)$treat_bin)
euc_test$p_value

# Aitchison Distance
set.seed(3423)
ait_test <- testBetaDiversity(dv = divnet_c_5, 
                              sample_specimen_matrix = divnet_c_5$X,
                              h0 = "aitchison",
                              n_boot = 1000,
                              groups = sample_data(ps_f_5_c)$treat_bin)

ait_test$p_value

```