### 2022-05-25 R/07-process-community.Rmd
remove phyloseq and replace with tibbles

---
title: "Sequence data processing"
author: "Andrew Morris"
date: "9/13/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir=normalizePath('../'))
```

```{r initialize-packages}
library(tidyverse)
#library(phyloseq)
library(decontam)
#library(microViz)
library(knitr)
library(vegan)
```

```{r import-data}
load(file = "HPC_Output/dada2_output.RData")

samples <- read_tsv('Output/barcode_master.tsv')
fluxes <- read_tsv('Output/fluxes.tsv')
qubit <- read_tsv('Data/post_pcr_qubit.tsv')

```

```{r data-munging}

metadata <-
  samples %>% 
  rename(sample = sample_id) %>% 
  left_join(qubit, by = c("pcr_well" = "well")) %>% 
  mutate(treat = (Vectorize(function (x) {if(x %in% c('PCRP', 'PCRN')) x
    else substr(x, 1, 3)}))(sample)) %>% 
  mutate(col = as.character(col)) %>% 
  select(sample, row, col, treat, qubit_ngml)

rm(qubit)
rm(samples)

# Check if taxa names match between the otu and taxa tables
sum(colnames(seqtab.nochim) != rownames(taxa))
# They match if = 0

colnames(seqtab.nochim) <- paste0("ASV", seq(ncol(seqtab.nochim)))
rownames(seqtab.nochim) <- metadata$sample
rownames(taxa) <- colnames(seqtab.nochim)

asvs <-
  seqtab.nochim %>% 
  as_tibble(rownames = "sample")

rm(seqtab.nochim)

taxa <-
  taxa %>% 
  as_tibble(rownames = "asv")

aes_data <-
  asvs %>% 
  pivot_longer(-sample, names_to = "asv", values_to = "count") %>% 
  inner_join(metadata, by = "sample") %>% 
  inner_join(taxa, by = "asv")

```

# Alpha Diversity for All Data

Here is alpha diversity for each treatment, positive/negative control, etc.
In addition, there is alpha diversity by row and column of the PCR plate. I don't see any relationship between PCR plate location and alpha diversity, which is good. It looks like the PCR positive control, which was a pure culture of E. Coli, has particularly low diversity, which is good. Let's pull that sample out and look at it.

```{r alpha-diversity-treatment}

aes_data %>%
  select(sample, treat, asv, count) %>% 
  filter(count > 0) %>% 
  count(sample, treat, count) %>%
  nest(data = -c(sample,treat)) %>%
  mutate(n_seqs = map_dbl(data, ~sum(.x$count * .x$n)),
         sobs = map_dbl(data, ~sum(.x$n))) %>%
  select(-data) %>%
  select(-sample) %>% 
  group_by(treat) %>% 
  rename(Reads = n_seqs, Richness = sobs) %>% 
  pivot_longer(-c(treat)) %>%
  ggplot(aes(x=treat, y=value)) +
  facet_wrap(~ name, nrow = 3, scales = "free") +
  geom_boxplot() +
  geom_jitter() +
  theme_classic() +
  labs(x = "Treatment", y = "Value")

```

```{r alpha-diversity-row-col}

aes_data %>%
  select(row, col, asv, count) %>% 
  filter(count > 0) %>% 
  mutate(row = as.numeric(factor(row)),
         col = as.numeric(col)) %>% 
  count(row, col, count) %>%
  nest(data = -c(row, col)) %>%
  mutate(n_seqs = map_dbl(data, ~sum(.x$count * .x$n)),
         sobs = map_dbl(data, ~sum(.x$n))) %>%
  select(-data) %>%
  rename(Column = col, Row = row, Richness = sobs, Sequences = n_seqs) %>% 
  pivot_longer(c(Row, Column), names_to = "row_or_column", values_to = "position") %>%
  pivot_longer(c(Sequences, Richness), names_to = "name", values_to = "count") %>%
  ggplot(aes(x=position, y=count)) +
  facet_grid(name ~ row_or_column, scales = "free") +
  geom_point() +
  theme_classic() +
  geom_smooth() + 
  scale_x_continuous(breaks = 1:9)

```

## Top 20 taxa in controls

```{r top20-function}
top20_taxa_in_treatment <- function(x, treatment) {
x %>% 
  filter(treat == treatment) %>% 
  select(-sample, -row, -col, -qubit_ngml) %>% 
  group_by(asv) %>% 
  mutate(count = sum(count)) %>%  
  ungroup() %>% 
  distinct() %>% 
  arrange(desc(count)) %>% 
  slice_head(n = 20)
}
```

### Potting Mix

```{r top20-potting-mix}
top20_taxa_in_treatment(aes_data, "PM0")

```

### Negative Control

```{r top20-negative-control}
top20_taxa_in_treatment(aes_data, "NC0")

```

### PCR Positive Control

```{r top20-pcr-pos}
top20_taxa_in_treatment(aes_data, "PCRP")

```

### PCR Negative Control

```{r top20-pcr-neg}
top20_taxa_in_treatment(aes_data, "PCRN")

```

### Soil Core

```{r top20-soil-core}
top20_taxa_in_treatment(aes_data, "SC0")

```

# Sequence Depth Distribution + Sample Composition

```{r}

# Look at distribution of sequencing depths
aes_data %>%
  select(sample, asv, count) %>% 
  group_by(sample) %>% 
  summarize(total = sum(count)) %>% 
  ggplot(aes(x = total)) +
  geom_histogram() +
  labs(x = "Sequences", y = "Count")

```

```{r}

aes_df <- 
  aes_data %>%
  select(sample, asv, count) %>% 
  pivot_wider(names_from=asv, values_from=count, values_fill = 0) %>%
  as.data.frame()
rownames(aes_df) <- aes_df$sample
aes_df <- aes_df[,-1]

dist <- vegdist(aes_df, method="bray")
pcoa <- cmdscale(dist, eig = TRUE, add = TRUE)
positions = pcoa$points
colnames(positions) <- paste0("pcoa", 1:ncol(positions))

percent_explained <- 100 * pcoa$eig / sum(pcoa$eig)

tibble(pe = percent_explained, 
       axis=1:length(percent_explained)) %>% 
  ggplot(aes(x = axis, y = pe)) +
  geom_line() +
  coord_cartesian(xlim = c(1,10)) +
  labs(title = "Bray PCoA Scree Plot", y = "Precent Explained", x = "PCo Axis")

labs <- c(paste0("PCo 1 (", format(round(percent_explained[1], 1), nsmall = 1), "%)"),
          paste0("PCo 2 (", format(round(percent_explained[2], 1), nsmall = 1), "%)"))

positions <- 
  positions %>% 
  as_tibble(rownames = "sample")

bray_data <-
  aes_data %>% 
  select(sample, treat, row, col) %>% 
  distinct %>% 
  inner_join(positions, by = "sample")

bray_data %>% 
  ggplot(aes(x = pcoa1, y = pcoa2, color = treat)) +
  geom_point() +
  labs(x = labs[1], y = labs[2], title = "Bray PCoA by Treatment") +
  coord_fixed(percent_explained[2]/percent_explained[1], clip="off")

bray_data %>% 
  ggplot(aes(x = pcoa1, y = pcoa2, color = col)) +
  geom_point() +
  labs(x = labs[1], y = labs[2], title = "Bray PCoA by PCR plate column") +
  coord_fixed(percent_explained[2]/percent_explained[1], clip="off")

bray_data %>% 
  ggplot(aes(x = pcoa1, y = pcoa2, color = row)) +
  geom_point() +
  labs(x = labs[1], y = labs[2], title = "Bray PCoA by PCR plate row") +
  coord_fixed(percent_explained[2]/percent_explained[1], clip="off")

```

# DECONTAM

```{r decontam-n-seqs}

aes_data %>% 
  filter(!treat %in% c('PM0', 'SC0', 'PCRN', 'PCRP')) %>% 
  mutate(ctrl_smpl = case_when(treat == "NC0" ~ "control", 
                               TRUE ~ "sample")) %>% 
  add_count(sample, count) %>%
  nest(data = -c(sample, ctrl_smpl)) %>%
  mutate(n_seqs = map_dbl(data, ~sum(.x$count * .x$n))) %>% 
  arrange(n_seqs) %>% 
  mutate(index = seq(nrow(.))) %>% 
  ggplot(aes(x=index, y=n_seqs, color=ctrl_smpl)) + 
  geom_point() +
  labs(y = "Library Size", x = "Index")

```

My negative controls have a lower library size than my samples.

## Decontam Combined Prevalence and Frequency

Here, I use DNA concentration values from Qubit and prevalence in negative
controls to identify potential contaminants. Several negative controls
had DNA concentration below the detection limit of the instrument and these
were set to 0.50 ng/mL, which is close to the DNA concentration of the
negative controls that were within the detection limit and is at the limit of the 
Qubit's detection limit.

In the plots below, ASV1 is a true sequence. The others are the ASVs identified
by decontam as potential contaminants.

```{r decontam-prep}

otu_table <-
  aes_data %>% 
  filter(!treat %in% c('PM0', 'SC0', 'PCRN', 'PCRP')) %>% 
    mutate(control = case_when(treat == "NC0" ~ TRUE, 
                               TRUE ~ FALSE)) %>% 
  select(qubit_ngml, control, sample, asv, count) %>% 
  pivot_wider(names_from = asv, values_from = count) %>% 
  mutate(qubit_ngml = case_when(qubit_ngml == 0 ~ 0.5, 
                               TRUE ~ qubit_ngml)) %>% 
  as.data.frame
concentration <- otu_table$qubit_ngml
otu_table <- otu_table[, -1]
control <- otu_table$control
otu_table <- otu_table[, -1]
rownames(otu_table) <- otu_table$sample
otu_table <- otu_table[, -1]
otu_mat <- as.matrix(otu_table)

```

```{r decontam}
contamdf <- isContaminant(otu_mat, conc = concentration, neg=control)

alleged_contaminants <- which(contamdf$contaminant)

quant_test <- 
  otu_mat[,c(1,alleged_contaminants)] %>% 
  as.data.frame %>% 
  as_tibble(rownames = "sample") %>% 
  mutate(conc = concentration) %>% 
  mutate(cont = control) %>% 
  pivot_longer(starts_with("ASV"))

ggplot(quant_test, aes(x = conc, y = value, color = cont)) +
  facet_wrap(~ name, scales = "free") +
  geom_point()

contaminants <- 
  which(contamdf$contaminant)[which(contamdf$contaminant) %in% 
  c(570, 810, 1283, 1859, 2233, 2777, 2926, 5687, 6696)]

```

Decontam identified `r length(which(contamdf$contaminant))` potential contaminants based
on prevalence and frequency. Visual inspection of abundance ~ concentration
plots indicated that `r length(contaminants)` were likely contaminants. These
ASVs were removed.

```{r remove-contaminants}
aes_data_no_contam <-
 aes_data %>%
  nest(data = -asv) %>% 
  filter(!asv %in% paste0("ASV", contaminants)) %>% 
  unnest(data)

```

# Combine ASV table and flux data

```{r combine-data}

flux_sd <- 
  fluxes %>% 
  mutate(sample = paste0('G', passage, toupper(treat), rep)) %>% 
  select(sample, flux_date:selected) %>% 
  mutate(passage = as.character(passage))

flux_sd <-
  flux_sd %>% 
  select(sample, ch4, selected)

full_data <-
  aes_data %>% 
  left_join(flux_sd, by = "sample")

write_tsv(full_data, 'Output/communty_data.tsv')

```

```{r old-stuff, eval = FALSE}

samples_filtered <- 
filter(samples, sample_id %in% flux_sd$sample)

fluxes_filtered <- 
  filter(flux_sd, sample_id %in% samples$sample_id) %>% 
  as.data.frame()

```