### 2022-07-06 Manuscript/manuscript.Rmd
start to update text and figures with new methods, results

---
title: Artificial ecosystem selection reveals relationships between microbiome composition and ecosystem function
author: "Andrew H. Morris^1^* and Brendan J. M. Bohannan^1^"
output:
  word_document:
    reference_docx: reference.docx
bibliography: bibliography.bib
csl: vancouver.csl
---

^1^ Institute of Ecology and Evolution, University of Oregon, Eugene, OR, USA

\* Correspondence: Andrew H. Morris, amorris3@uoregon.edu

Competing Interests: We declare we have no competing interests.

```{r options, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

```

```{r initialize-packages}
library(tidyverse)
library(broom)
library(broman)
library(morris)
library(phyloseq)
library(ggpubr)
library(knitr)
library(corncob)
library(ggthemes)
library(microViz)
library(captioner)
library(knitr)
library(kableExtra)
library(flextable)
library(ggtext)
figures <- captioner()
tables <- captioner(prefix = "Table")
```

```{r load-data}
heritability <- read_tsv('../Output/heritability.tsv')
fluxes <- read_tsv('../Output/fluxes.tsv')

# CH4 models
load(file = '../Output/response_model.RData')
load(file = '../Output/response_plot.RData')
load(file = '../Output/herit_model.RData')
load(file = '../Output/herit_plot.RData')

# 16S models
load(file = '../Output/richness_models.RData')
load(file = '../Output/phyloseq.RData')
load(file = '../Output/beta_model.RData')
load(file = '../Output/beta_fig.RData')
load(file = '../Output/da_corncob.RData')
load(file = '../Output/da_fig.RData')

source('../R/functions.R')

```

# Abstract

Microbiomes mediate important ecosystem functions, yet it has proven difficult
to determine the relationship between microbiome composition and the rate of
any ecosystem function. This challenge remains because microbiomes are often
complex, it is difficult to know a priori which microbiome members limit the
rate of an ecosystem function, and it is usually challenging to manipulate
microbiome composition directly. We propose that artificial ecosystem selection
can overcome these challenges. Artificial ecosystem selection involves creating
replicate ecosystems, selecting replicates with a desired ecosystem-level
trait, and using the microbiomes from the selected ecosystems to inoculate a
new set of ecosystems. We performed artificial ecosystem selection on replicate
soil microcosms to select for increased soil methane oxidation rate. We
observed a strong response of methane oxidation rate to artificial ecosystem
selection with a 51% increase in methane oxidation rate per round over five
rounds of selection. This suggests that a large portion of the variation in
methane oxidation rate in these soils can be attributed to microbiome
variation. We also found that selection did not enrich for known methane
oxidizers; rather microbes that were enriched after selection included 24
families such as Puniceicoccaceae, Elsteraceae, and Cytophagaceae. This result
is in contrast to the typical assumption that the rate of an ecosystem function
is limited by the final step in the associated microbial pathway. Our study
demonstrates that artificial ecosystem selection is a useful strategy for
revealing relationships between microbiome composition and ecosystem function.

# Introduction

Microbiomes mediate a variety of important functions in ecosystems, and there
is great interest in understanding how attributes of microbiomes may influence
variation in ecosystem functions [@crowther2019]. Biodiversity-ecosystem
function relationships have been described for a variety of macroorganismal
communities. For example, plant species richness correlates with productivity
and marine community diversity correlates with functions such as production,
consumption, and nutrient release
[@cardinale2006; @hooper2012; @gamfeldt2015]. While there is evidence that
variation in the microbiome contributes to variation in ecosystem functions,
describing this relationship has been elusive [@rocca2015; @graham2016].

One reason for this difficulty is that it is challenging to directly manipulate
microbiome composition and diversity the way we can for macroorganismal
communities. Instead, we must use genetics through marker gene or metagenomic
sequencing to characterize the microbiome.  However, studies that measure the
taxonomic or functional diversity of a microbiome through these approaches
rarely find a correlation between diversity and function [@graham2016]. Another
approach is to quantify cells or genes using  quantitative
PCR. Studies such as this still rarely find a correlation between the abundance
of a functional group and the rate of the corresponding process [@rocca2015].
As a result, it remains unclear how and when microbiome variation contributes
to variation in the rate of ecosystem functions.

Another likely reason that microbiome-ecosystem function relationships have been
difficult to document is that the most common approaches for looking for a
relationship requires some prior knowledge of the likely causal links between
microbiomes and ecosystem functions. For example, many studies have looked for
relationships with the abundance or diversity of a "marker gene" for a
particular function. Marker genes are microbial genes that code for an enzyme or
enzyme-subunit known to be involved in a particular function. But this approach
assumes that one can identify the "right" marker gene a priori which, given
how little is understood about microbial diversity, seems highly unlikely. What
is needed are approaches to microbiome-ecosystem
function mapping that do not require this degree of prior knowledge.

Yet another likely reason that microbiome-ecosystem function relationships have
been difficult to document is that microbiomes can be related to ecosystem
functions in two related but distinct ways that historically have been difficult
to separate. One way is that microbiomes may simply be conduits through which
the environment drives function. That is to say that the environmental
conditions completely regulate the attributes of the microbiome and a shift in
the environment causes a shift in microbiome attributes, such as the abundance
of a microbial functional group. This shift results in a change in the rate of
an ecosystem function. In this case, microbiome attributes would be
statistically associated with variation in ecosystem function, but also strongly
covary with environmental conditions. Therefore, there would be no association
between microbiomes and ecosystem functions independent of the environmental
conditions. Under this scenario, knowing anything about the microbiome would not
necessarily provide you with information regarding the causes of the change in
ecosystem function or improve one's ability to predict changes in ecosystem
function. Furthermore, attempting to alter ecosystem function by altering the
microbiome without changing the underlying environmental conditions would likely
be fruitless.

Another possibility is that the microbiome itself alters the rate of ecosystem
function independent of the environment. In this case, one cannot predict the
rate of ecosystem function without understanding changes in the microbiome. In
addition, managing the microbiome through inoculation or selective antibiotics
would potentially be an effective approach for enhancing or mitigating certain
changes to ecosystem functions. These two scenarios are not mutually
exclusive, nor are they likely unique to microbes, but methods to determine the
relative important of these two scenarios have not until recently been employed.

The problem of connecting microbiome attributes to the rate of an ecosystem
function is analogous to the problem of connecting genomic variation to
phenotypic traits in organisms, and solutions to this problem may be found by
exploring this analogy [@morris2020]. For example, one approach to the problem of
genotype-phenotype mapping is to perform artificial selection on a trait of
interest and identify genes that respond to selection. Such genes are likely to
be causally linked to the trait under selection. An analogous approach for
exploring microbiome-ecosystem function relationships is gaining acceptance
among microbiologists [@swenson2000; @panke-buisse2015]. Microbiomes have been
shown to respond to selection at the ecosystem-trait level [@swenson2000],
and some ecosystem-scale microbiome traits have been shown to be transferable
through microbiome inoculations [@panke-buisse2015]. Applying artificial
selection to whole microbiomes could be a novel approach for quantifying how
much of the variation in the rate of an ecosystem function could be attributed
independently to microbiome variation [@goodnight2000]. In addition, this
approach could be used to identify microbiome attributes, such as genes or taxa,
associated with the rate of an ecosystem function, although, to my knowledge
this has not previously been accomplished [@arias-sanchez2019; @morris2020]. 
Such an approach has the potential to expand our ability to model
microbial ecosystem-function relationships more accurately and allow us to
manage ecosystems for particular outcomes [@fuhrman2009; @schimel1998]. Here I
use this approach to estimate how much of the variation in ecosystem function
(the flux of CH~4~ gas between the soil and atmosphere) can be attributed to
variation in the microbiome and to identify microbial markers associated with
this ecosystem function.

Soil CH~4~ oxidation is a suitable function for this study because there is
evidence that it may vary with microbiome attributes. For example, variation in
CH~4~ emissions in arctic permafrost is correlated with the transcriptional
activity of certain methanogens [@freitag2009; @freitag2010]. In addition,
soil microbiome phylogenetic variation is a strong predictor of CH~4~ oxidation in
forests and pastures of the Brazilian Amazon [@meyer2020]. Finally,
methanogenesis and methanotrophy are two of the most deeply conserved microbial
physiologies and are represented in a narrow range of taxa and so the taxonomic
composition of the microbiome is more likely to be associated with the rate of
CH~4~ flux than other broader or more shallowly conserved functions
[@schimel1998; @martiny2013]. This suggests that microbiome variation might
drive variation in CH~4~ oxidation in soils.

For the current project, we applied artificial ecosystem selection on soil
microbiomes by selecting on the ecosystem-scale CH~4~ oxidation rate in order
to estimate the amount of microbiome variation underlying variation in soil CH~4~ oxidation rate
[@swenson2000]. To identify which microorganisms could be markers of CH~4~
oxidation rate, we compared the composition of the artificially selected
microbiomes to a control set of microbiomes without selection. Artificial
ecosystem selection has a similar effect to enrichment culturing by amplifying
the population of interest [@swenson2000]. This will reduce the diversity of
the soil microbiome and allow for greater power in detecting significant
markers of CH~4~ oxidation rate. We then evaluated whether these markers meet
our underlying assumptions about which taxa limit the rate of ecosystem
function.

# Materials and Methods

## Experimental design

We performed an artificial ecosystem selection experiment [sensu @swenson2000]
by passaging replicate soil microbiomes. The trait we selected on was
CH~4~ oxidation rate. This experiment had two selection lines with twelve jars
each for a total of twenty-four jars per passage. One line was a positive
selection line where the two or three jars with the highest CH~4~ oxidation rate
were chosen to inoculate the next set of positive jars. The other line was a
neutral selection line where an equal number of jars as the positive line were
chosen at random to inoculate the next set of neutral jars. The number of jars
chosen was based on the distribution of fluxes among the positive jars, i.e., we
chose the top three jars unless only two jars had considerably greater CH~4~ oxidation
rates based on visual inspection of histograms. The experiment was carried out
over five passages until a significant divergence in functional rates was 
observed between the two selection treatments.

The initial soil microbiome was sampled from the top 10 cm of an upland mineral
soil under a deciduous forest ecosystem near the University of Oregon campus in
Eugene, OR, USA. Incubations were performed in 500 mL mason jars with rubber
septa installed in the lids. Each jar was sterilized with 70\% ethanol to which
was added 45 g of autoclaved potting mix, 5 g of living soil, and 3.5 mL of
sterile deionized water to bring them to 60\% of field capacity. The potting mix was
Lane Potting Mix from Lane Forest Products (Eugene, OR, USA). The jars were
then homogenized, capped, and injected with 4.3 mL of 99\% CH~4~ to bring
the headspace concentration to approximately 1000 ppm CH~4~. To create the
two treatment groups, twenty-four jars were created in an identical manner and
then randomly assigned to either the positive or neutral selection line. Jars
were flushed and respiked twice per week to maintain aerobic conditions and
elevated CH~4~ concentrations and were incubated at ambient temperature for
approximately three weeks. Methane oxidation rates were determined at the end
of the incubation period. For the positive treatment, the three jars with the
greatest CH~4~ oxidation rates were chosen to inoculate the next generation.
For the neutral treatment, three jars were randomly selected to inoculate the
next generation. For each treatment, these three jars were homogenized and 5 g 
of this homogenized soil was used to inoculate the next set of jars and which
represents a 10% subsampling. The
next set of jars were created in an identical manner to the first generation
with fresh autoclaved potting mix and the same moisture and CH~4~ content.

## Methane oxidation rate

Methane oxidation rates were determined after flushing and spiking jars to
approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from each
jar immediately after spiking and then at time points 3, 6, 24, and 48 hours for
a 5-point curve. Samples were immediately injected into a SRI model 8610C gas
chromatograph equipped with a flame ionization detector (SRI Instruments, Torrance, CA, USA) to determine the headspace CH~4~
concentration. Fluxes were calculated from a first-order exponential decay
function as decay constant $k$ with units day^-1^. Oxidation rates are presented
as the additive inverse of $k$ (i.e., $-k$) so that a more positive value
represents a greater oxidation rate. The jars selected for the positive
treatment in passage 2 had the lowest CH~4~ oxidation rate of the twelve jars
due to a calculation error in the CH~4~ oxidation rate. All other passages
correctly used the top three jars.

## Soil DNA extraction and sequencing

A subsample of soil from the starting inoculum and from every jar in passages 2
and 5 was collected and stored at $-80^\circ$C. Soil DNA was extracted from 0.25
g soil. Negative controls were extracted from autoclaved potting mix and
DNase-free water. Extractions were performed using the DNeasy PowerSoil kit
(QIAGEN, Düsseldorf, Germany) and quantified using Qubit dsDNA HS Assay Kit (Thermo Fisher
Scientific, Inc., Waltham, MA, USA). To estimate the diversity and relative abundance of
the bacterial and archaeal taxa in our soil ecosystems, we sequenced the V4
region of the 16S rRNA gene using the using the 515F - 806R primer combination
[@caporaso2011]. PCR mixtures were: 10 $\mu$l NEBNext Q5 Hot Start HiFi PCR
master mix, 9.2 $\mu$l primer mixture (1.09 $\mu$M concentration), and 0.8
$\mu$l of DNA template. Reaction conditions were: $98^\circ$C for 30 s
(initialization); 35 cycles of $98^\circ$C for 10 s (denaturation), $61^\circ$C
for 20 s (annealing), and $72^\circ$C for 20 s (extension); and $72^\circ$C for
2 m (final extension). Reactions were performed in triplicate and then combined.
Amplicons were purified twice using 0.8x ratio Mag-Bind
RxnPure Plus isolation beads (Omega Bio-Tek, Norcross, GA, USA). Sequencing libraries were
prepared using a dual-indexing approach [@fadrosh2014; @kozich2013]. Amplicon
concentrations were quantified using Qubit and multiplexed at equimolar concentration. Sequencing
was performed at the University of Oregon Genomics Core Facility on the Illumina
NovaSeq 6000 with paired-end 150 bp reads (Illumina, Inc., San Diego, CA, USA).

## Bioinformatics

Bioinformatics processing was performed in 'R' [@rcoreteam2018].
Demultiplexed sequencing reads were denoised using 'DADA2' to generate a table
of amplicon sequence variants (ASVs)
[@callahan2016]. Taxonomic assignment for the 16S reads was performed
using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The presence of
contaminants was evaluated using both the prevalence and frequency methods from
'DECONTAM' by comparing samples to negative controls [@davis2018]. We did not identify
any obvious contamination using these methods.

## Statistical Analysis

Statistical analyses were performed in 'R' [@rcoreteam2018]. To test
whether there was a significant change in CH~4~ oxidation rate ($k$) as a
response to selection, we tested a difference in slopes between the positive
and neutral selection lines. The CH~4~ oxidation rates were strongly 
right-skewed with most values close to zero and few large, positive values. This
resulted in residuals that did not meet the assumptions of constant variance
and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
transformed to better meet the assumptions of a linear model and to make
figures easier to interpret. First, we tested the difference of slopes between
the positive and neutral treatments by
fitting two nested models with and without the interaction between treatment
and passage using 'lm'. We compared these
models using an F-test with the 'anova' function. We
then fit the full model with the interaction term to estimate the slopes for each treatment, which
represented the change in CH~4~ oxidation rate per passage as a response to
selection.

To estimate the proportion of variance in CH~4~ oxidation rate due to variation in
the microbiome, we estimated narrow-sense heritability ($h^2$) as the
regression of mid-offspring on mid-parent [@falconer1996]. The mid-parent was the
mean for all three selected jars and the mid-offspring was the mean for all jars
produced by those parents. First, we tested if there was an effect of treatment
on the heritability estimate. We compared these models using an F-test
test with the 'anova' function. We then fit the full model to estimate the
heritability of the Positive and Neutral treatments.

Richness was estimated using the method from @hurlbert1971 with a subsample size of `r min_seqs` using the 'rarefy'
function in 'vegan' [@oksanen2019]. We
tested a difference in richness between Passages 2 and 5 with both treatments
combined. Then we subset the samples from Passage 5 and compared richness between
the Positive and Neutral treatment. Next, we compared beta-diversity as the 
Bray-Curtis dissimilarity by averaging random subsets with a subsample size of `r min_seqs` using the 'avgdist' function in 'vegan' [@bray1957; @oksanen2019].
We tested a difference in centroid and dispersion of beta
diversity between the Positive and Neutral treatment and Passage 2 and 5 with
permutational analysis of variance (PERMANOVA) using the 'adonis2' function from
'vegan' [@mcardle2001; @oksanen2019].

To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
differential abundance between the two treatments in passage 5. We first
grouped ASVs at the family level keeping ASVs that lacked a family-level taxonomic
assignment by grouping them at a higher level. We
then subset the samples in Passage 5 and removed all families with a prevalence
of less than 10\% in either treatment. We used three methods for testing differential abundance: ANCOM-II, ALDEx2, and CORNCOB. We then
pulled out the consensus taxa that were significant
with all three tests and plotted their relative abundances. For ANCOM-II, we used the 'ancom' function in the 'ANCOM-BC' package with a cutoff of W = 0.7 [@kaul2017; @lin2020]. For ALDEx2, we used the 'aldex' function in the 'ALDEx2' package with the Welch's t test and we used an effect size of
1 as our significance threshold [@fernandes2014]. Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob' package with the Wald test and without bootstrapping [@martin2020].

# Results

## Response to selection on methane oxidation rate

We observed a response to selection on whole-ecosystem soil CH~4~ oxidation
rate (`r figures("response", display = "cite")`; difference of slopes:
F~`r compare_response_mods[2,3]`,`r compare_response_mods[2,1]`~ = `r myround(compare_response_mods[2,5], 2)`, p = `r myround(compare_response_mods[2,6], 2)`). At the start of the experiment, the Positive
treatment had a mean CH~4~ oxidation rate that was
`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`\% lower than the
Neutral treatment `r print_lm(response_model, 3, 'difference of y-intercepts')`. 
There was no change in CH~4~ oxidation rate of the Neutral
treatment over the five passages `r print_lm(response_model, 2)`. By contrast,
the Positive treatment had a `r round(per_slope)`\% increase in CH~4~
oxidation rate per passage `r print_lm(response_model, 4)`.

## Heritability of methane oxidation

We estimated heritability as the regression of mid-offspring on mid-parent (`r figures("herit", display = "cite")`).
Offspring CH~4~ oxidation rates were correlated with parental CH~4~ oxidation
rates in both the Positive treatment `r print_lm(herit_model, 4)` and the
Neutral treatment `r print_lm(herit_model, 2)`. Notably, the sign of the effect
was reversed with a positive heritability for the Positive treatment and a negative heritability for the Neutral treatment.

## Taxonomic richness

Richness of ASVs decreased between Passage 2 (median = `r myround(rare_median[1, 2], 1)`, IQR = `r myround(rare_median[1, 3], 1)`) and 5 (median = `r myround(rare_median[2, 2], 1)`, IQR = `r myround(rare_median[2, 3], 1)`; Kruskal-Wallace test: $\chi^2$ = 
`r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; pairwise Wilcoxon test: p < 0.001). However, there was no
difference in richness between the Positive and the Neutral treatment in
Passage 2 or 5 (pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
correlation between richness and CH~4~ oxidation rate across all passages
and treatments (Spearman's rho = `r myround(s[1,1], 1)`, p = `r myround(s[1,3], 1)`).

## Community dissimilarity

Bray-Curtis dissimilarity of the soil microbiome varied strongly by passage and
weakly by treatment and the interaction between passage and treatment (`r figures("beta", display = "cite")`; `r tables("tidy_permanova", display = "cite")`). 
Passage explained `r myround(100*tidy_permanova[1, 4])`\% of the
variation in Bray-Curtis dissimilarity.  Treatment explained 
`r myround(100*tidy_permanova[2, 4])`\%
of the variation. There was also an interaction
between treatment and passage, which explained 
`r myround(100*tidy_permanova[3, 4])`\% of the variation. We found there was 
no difference in dispersion between treatments or passages (F~3,44~ = 0.91, p = 0.45).

## Taxa that responded to selection

To investigate which taxa responded to selection on soil CH~4~ oxidation rate, we
first looked at which taxa were unique to the Positive and Neutral treatments
aggregated at the family level. Several taxa were present in the Positive
treatment and absent in the Neutral treatment. Most of these families had low
prevalence (present in 3 or fewer samples) and low abundance (median < 5 reads).
There were two families unique to the Positive treatment with relatively high
prevalence. This included an ASV that was a member of the Bacteroidia Class with
no lower taxonomic designation. This ASV had a prevalence of 10/12 and a median
abundance of ~2 reads. The other prevalent family was a member of the
*Silvanigrellaceae*, a newly described family placed in its own order. This family
was present in all 12 samples and had a median abundance of 8 reads.
*Silvanigrella* is the only cultivated member of
*Silvanigrellaceae* and was isolated from a temperate fresh water lake
[@hahn2017]. Of the families unique to the Neutral treatment, only one had a
prevalence greater than 2/12. This family, *Armatimonadaceae*, was present in half
of the Neutral samples (prevalence = 6/12). The type strain for *Armatimonadaceae* was isolated from
the rhizosphere of *Phragmites australis* [@tamaki2011].

The remaining families were shared between the Positive and Neutral treatment,
but some differed in their relative abundance. To identify taxa that responded
to selection on soil CH~4~ oxidation rate, we tested the differential relative
abundance of families in the Positive jars relative to the Neutral jars within
Passage 5 using three methods and then plotted the taxa identified by all three methods. We identified `r length(da_fams)`
families that were enriched or depleted in the Positive treatment relative to
the Neutral treatment (`r figures("da", display = "cite")`).

Overall, none of the families enriched in the Positive selection treatment
are known methanotrophs. Several taxa identified had a 
higher taxonomic designation that contains methanotrophs, for example, the
Gammaproteobacteria class had a large effect size.
This class is one of the groups containing methanotrophic families such as
*Methylococcaceae.* However, Gammaproteobacteria is among the most diverse 
groups in the Prokaryotes, so this is not strong evidence for a selection response
by methanotrophs [@garrity2005].
The next family
was *Puniceicoccaceae*, which is a member of the phylum Verrucomicrobia and
contains marine microbes. The Verrucomicrobia is a diverse group that contain
known methanotrophs as well as ammonia-oxidizing bacteria [@freitag2003]. 
An ASV from the order Armatimonadales was also enriched in the Positive treatment.
This order also contains the family *Armatimonadaceae*, which was
found to be unique to the Neutral treatment. *Cytophagaceae* was also enriched in
the Positive treatment and contains a number of mainly aerobic heterotrophs
that can digest a variety of macromolecules [@mcbride2014]. Notably, the family
*Gemmatimonadaceae* has a single bacterial bin, Candidatus '*Methylotropicum kingii*',
that contains methanotrophy genes, but this family was depleted in the Positive
selection treatment [@bay2021].

# Discussion

We observed a response to
selection on whole-ecosystem soil CH~4~ oxidation rate. This type of response
has been observed for other functions, such as chloroaniline degradation in
water [@swenson2000], but this is the first instance, to our knowledge, of
selection being performed directly on soil biogeochemical cycling. With this
experimental design, we attempted to create multiple ecosystems that were very
similar with respect to the type and quantity of soil, soil moisture content,
and the headspace concentration of CH~4~. we further controlled for environmental
variation across replicates by having twelve replicates per treatment, using
controlled laboratory conditions, and by having a random selection treatment as
a control. Therefore, the observed response to selection on
CH~4~ oxidation rate is likely due to changes in the microbiome.

To investigate the relationship between microbiome structure and
ecosystem function, we wanted to determine how much of the variation in soil CH4
oxidation was attributable to variation in the microbiome. This ecological
question is analogous to the problem of estimating heritability in quantitative
genetics [@morris2020]. Variation in an organismal trait is
determined by the sum of genetic variation, environmental variation, and the
interaction between the two [@falconer1996]. One goal of quantitative
genetics is to determine the proportion of phenotypic variation attributable to
genetic variation. This is commonly estimated as the narrow-sense heritability
defined as the additive genetic variance. This goal can be achieved through
artificial selection experiments with selection on the character of interest
[@falconer1996].

The heritability we observed in the Positive treatment was large ($h^2$ =
`r myround(tidy(herit_model)[4, 2], 2)`) relative to other studies of biological
heritability [@visscher2008]. However, given the small sample
size and wide confidence intervals (95\% CI: 
`r myround(confint(herit_model)[4, ], 2)`) it is likely that a larger study could
more precisely estimate the true heritability. We intentionally eliminated much of the
environmental variation that would be present in natural ecosystems, but it is
unclear how much jar-to-jar variation there is in the composition of the soil
substrate. Given these data, though, we conclude that variation within
the soil microbiome has a very strong effect on soil CH~4~
oxidation rates independent of any environmental variation.

The negative heritability observed in the Neutral treatment could indicate a
negative genetic correlation between CH~4~ oxidation rate and traits
associated with persistence in a jar [@conner2003]. The conditions in the
jar favor organisms that can grow in potting mix, reproduce within a three week
incubation, and survive passaging between jars. These traits might be
negatively correlated with a community's ability to oxidize CH~4~ indicating
a potential trade-off [@conner2003]. This trade-off would result in a
negative heritability of CH~4~ oxidation rate in the Neutral treatment,
which did not experience selection on CH~4~ oxidation rate.

Since there was a response to selection, we wanted to identify
which taxa responded to selection. That is, the taxa that
were enriched in the Positive treatment relative to the Neutral treatment. 
Studies of microbiome structure-function
relationships typically quantify marker genes for the final enzyme in a pathway
to try to predict the rate of flux through that pathway. However, of the 
`r ntaxa(sig_taxa_treat)` taxa identified by the differential abundance test, none
were members of a family known to oxidize CH~4~. This suggests
that, in this system, the taxa that perform the final step in the CH~4~ oxidation pathway
do not
limit the rate of flux through the pathway. Traits that we assume are important
for regulating function (such as the diversity or relative abundance of
methanotrophs estimated from phylogenetic marker genes) may not be important for
determining variation in function.
Perhaps once a sufficient number of methanotrophs are present, the
process itself is limited by other metabolic processes in the ecosystem, such
as nitrogen cycling mediated by non-methanotrophic organisms.

We found that `r nrow(dt_data[dt_data$x > 0,])` taxa were enriched in the Positive treatment,
which might indicate that the soil CH~4~ oxidation rate in these jars is controlled by
multiple taxa. 
This conclusion is analogous to the conclusion in quantitative genetics that
most traits at the organismal level are the result of the interaction between
many independent genes. Indeed, very few traits or genetic diseases are the
result of a single gene or mutation [@kerem1989; @macdonald1992]. This results in the use of polygenic risk scores for predicting
phenotype or disease risk in humans [@reich2001; @hindorff2009]. An analogous conclusion can be drawn for predicting
the rate of ecosystem function from the presence or relative abundance of 
microbial taxa. The rate of an ecosystem function at the whole-ecosystem level is 
the result of interactions among a variety of disparate taxa with different traits.
To better understand microbiome structure-function relationships, our results suggest that we
should move away from single marker genes and instead investigate the role of
multiple functional groups in determining the function of ecosystems whether it
is CH~4~ flux from soil or host-microbiome health.

Taxonomic richness dropped precipitously
between Passage 2 and Passage 5. There are several explanations for this drop
in diversity. The initial community was a diverse microbiome sampled from a natural soil.
With the initial inoculation and at
each subsequent passaging, the ecosystem was subsampled to 10\% (5 g of living
soil was combined with 45 g of sterilized substrate). This subsampling likely
explains a large part of the drop in diversity. In addition, beyond the imposed
selection regime based on CH~4~ oxidation rate, these soil microbiomes were under
a variety of selection pressures such as persistence in a jar, survival during the transfer
from one jar to the next, and the ability to colonize a new jar in a couple of
weeks, to name a few. These selection pressures would have further filtered
the microbiome for taxa that could persist in this laboratory environment.

In addition to the drop in richness, there was also a decrease in beta
diversity among the jars as a result of passaging. This could be due to the
homogenization step between each passage. Once the jars were selected to generate
the next set of jars, the soil from those jars was homogenized and this homogenate
was used to inoculate the next set of jars. The goal of this step was to "shuffle"
membership in the microbiomes among the jars in order to test different
combinations and relative abundances of taxa in terms of their effect on CH~4~
oxidation rate. This likely also made the communities more similar in their
membership by increasing the likelihood that each taxon was represented in each
jar. Another explanation for this biotic homogenization is the aforementioned
selection for persistence in a jar. These selection pressures would have 
favored the subset of taxa suited to these laboratory conditions regardless
of whether they were in the Positive or Neutral selection line. This can be
seen by the relative similarity of the Positive and Neutral jars in Passage 5.
Despite the Positive treatment undergoing selection on CH~4~ oxidation rate,
both the Positive and Neutral jars were experiencing selection on a variety
of other traits that made the overall soil microbiomes appear similar.

One caveat that is important to keep in mind is that 16S rRNA genes are not
markers of functional characteristics of microorganisms. They are simply
phylogenetic markers useful for determining the relative position of taxa in a
phylogenetic tree. In addition, assigning functions to taxa based solely on
their taxonomy is a fraught exercise because many microbial functions are not
very deeply conserved, are spread across disparate lineages of the tree of life,
or are easily transferred between distantly related taxa. However,
methanotrophy and methanogenesis are two of the most deeply conserved microbial
traits [@martiny2013]. This results in part from the fact that
these functions require multiple genes to perform and are therefore not easy to
evolve independently or to transfer between lineages through horizontal gene
transfer. However, in order to be confident that the response we observed was not
simply due to a relative increase in methanotrophs, we would need to apply
a functional approach to characterizing the microbiome, for example, by sequencing
whole metagenomes.

Future research in this area should use methods that are able to detect
functions and metabolic pathways within the community that respond to
selection. For example, genome-resolved metagenomics could identify pathways
that are enriched in the Positive selection treatment. This would further
advance our understanding of the exact traits that are enriched in high-flux
ecosystems and therefore could be markers of ecosystem function. In addition,
future studies could build on experimental results such as these to sample a
natural gradient of ecosystem CH~4~ fluxes to determine whether the taxa identified
in our selection experiment could be useful indicators of CH~4~ flux in the
field. These studies would allow us to build a picture of the multiple traits
that drive variation in ecosystem function at the whole-ecosystem level.

# Conclusion

We performed an artificial ecosystem selection experiment on whole-ecosystem
CH~4~ oxidation rate to determine whether variation in microbiome
composition can contribute to variation in ecosystem function. We observed
a significant response to selection on CH~4~ oxidation rate - the first
example of an experiment performing selection on soil biogeochemical cycling.
In addition, we observed strong heritability of CH~4~ flux between
passages, suggesting that variation in microbiome composition could be
a major source of variation in CH~4~ flux in ecosystems. Surprisingly, 
the taxonomic groups that experienced selection on CH~4~
oxidation rate were not enriched in methanotrophs. This suggests that
understanding a single functional group is insufficient for predicting the
effect of microbiome composition on the rate of ecosystem function.
Instead, we may need to investigate alternative functional groups or multiple
interacting taxa to understand the role of microbiomes in ecosystem function. While it has often been assumed that
microorganisms play a major role in determining variation in ecosystem function, 
this has not previously been shown empirically. This experiment demonstrates that
variation in microbiome composition can contribute to considerable variation
in ecosystem function.

# Acknowledgments

This project was supported by the National Science Foundation Graduate Research Fellowship Program (grant no. DGE 1255832) and the ARCS Foundation Florence and Mike Nudelman Scholarship.

# Competing Interests

We declare we have no competing interests.

# Data Availability Statement

The 16S rRNA sequencing data generated during the current study are available in the NCBI Sequence Read Archive (SRA) under BioProject accession number PRJNA832314, https://www.ncbi.nlm.nih.gov/sra/PRJNA832314. The metadata generated during the current study as well as the scripts to recreate the analysis are available on Github, https://github.com/amorris28/artificial_ecosystem_selection.

# Supplement

# References

::: {#refs}
:::

# Figures

```{r response, fig.width=4, fig.height=3}

p_response

```

`r figures("response", "Response to selection on soil CH~4~ oxidation rate fit by ordinary least squares regression. The y-axis is the log~10~ of the additive inverse of the first-order exponential decay constant $k$ (i.e., $-k$) with units day^-1^. The x-axis is the passage number. The orange line and points are the Positive treatment and the gray line and points are the Neutral treatment.")`

```{r herit, fig.width=5, fig.height=3}

p_herit

```

`r figures("herit", "Ordinary least squares regression of mid-offspring CH~4~ flux on mid-parent CH~4~ flux ($-k$ with units day^-1^). Mid-parent is the mean of the jars selected to inoculate the next passage. Mid-offspring is the mean of all twelve jars produced in one passage. The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment")`

```{r beta, results=FALSE, fig.height=2.5}

p_beta

```

`r figures("beta", "Principal component analysis plot of beta diversity for all jars. Dissimilarities are based on rarefied Bray-Curtis dissimilarity. Colors represent treatment and shapes represent passage number.")`

```{r da, results=FALSE}

p_log10

```

`r figures("da", "Differentially abundant family-level taxa identified by ANCOM-II, ALDEx2, and CORNCOB. Values on the x-axis are relative abundances on a log~10~ scale. Points are colored by treatment. Large points with error bars are the median relative abundance and interquartile range. Taxa in the top panel are enriched in the Positive treatment and taxa in the bottom panel are depleted in the Positive treatment. Taxa are sorted by their effect size with taxa at the top having the largest positive effect size and taxa at the bottom with the largest negative effect size.")`

`r tables("adonis", "PERMANOVA on Bray-Curtis dissimilarities.")`

```{r adonis}

tidy_permanova %>%
	flextable()

```