### 2022-07-14 R/08-community-analysis.Rmd
Add new decontam results, add dbrda results, add new discussion draft, update methods

---
title: "Community Analysis"
author: "Andrew Morris"
date: "6/28/2022"
output: 
  html_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = TRUE)
knitr::opts_knit$set(root.dir=normalizePath('../'))
set.seed(1337)

```

```{r initialize_packages}

library(phyloseq)
library(knitr)
library(magrittr)
library(vegan)
library(ggpubr)
library(ggthemes)
library(broom)
library(ggtext)
library(ALDEx2)
library(corncob)
library(ANCOMBC)
library(tidyverse)

source('R/functions.R')

```

```{r import-data}

aes_data <- read_tsv('Output/community_data.tsv')

```

```{r fix-treatment}

aes_data <-
  aes_data %>% 
  mutate(passage_treat = treat) %>% 
  separate(passage_treat, c('passage', "treat"), sep = 2, remove = FALSE) %>% 
  select(sample, passage_treat, passage, treat, selected, ch4, asv, count, Kingdom:Species, seq)

```

```{r fix-taxonomy}

# There are "Unknown Family" taxa with Genus and Class, but Order is Incertae Sedis.
# I'm just going to remove everything below Class and call it "Unclassified"

fill_taxonomy <- function(tab) {
    if(is.na(tab[[1]])) tab[[1]] = "Unclassified"
    for(i in 2:length(tab)) {
      if(is.na(tab[[i]])){
        if(grepl("unclassified", tab[[i-1]])) tab[[i]] = tab[[i-1]]
        else tab[[i]] = paste(tab[[i-1]], "unclassified", sep = "_")
      }
    }
    tab
}

aes_data <-
  aes_data %>% 
  nest(data = -c(asv, Kingdom:Species)) %>% 
  nest(taxa = Kingdom:Species) %>% 
  mutate(taxa = map(taxa, ~fill_taxonomy(.x))) %>% 
  unnest(taxa) %>% 
  unnest(data)

```

# Exploratory

```{r n-seqs}

aes_data %>% 
  group_by(sample) %>% 
  summarize(total = sum(count)) %>% 
  ggplot(aes(x = total)) +
  geom_histogram() +
  theme_classic() +
  labs(title = "Distribution sequencing depths",
       x = "Library size",
       y = "Frequency")

seq_range <-
  aes_data %>% 
  group_by(sample) %>% 
  summarize(total = sum(count)) %>% 
  summarize(seq_range = range(total)) %$% 
  seq_range

min_seqs <- seq_range[1]

a_analysis <- 
  aes_data %>%
  count(sample, count) %>%
  nest(data = -sample) %>%
  mutate(n_seqs = map_dbl(data, ~sum(.x$count * .x$n)),
         sobs = map_dbl(data, ~sum(.x$n)),
         rare = map_dbl(data, ~rarefy(rep(.x$count, .x$n), sample=seq_range[1]))) %>%
  select(-data)

a_analysis %>%
  select(sample, n_seqs, sobs, rare) %>%
  pivot_longer(-c(sample, n_seqs)) %>%
  ggplot(aes(x=n_seqs, y=value, color=name)) +
  geom_point() +
  geom_smooth(se=FALSE) +
  coord_cartesian(ylim=c(0, 6000)) +
  theme_classic() +
  labs(x = "Sequencing Depth", y = "Alpha Diversity")

a_analysis %>%
  select(sample, n_seqs, sobs, rare) %>%
  mutate(treat = substr(sample, 1, 3)) %>% 
  select(-sample) %>% 
  group_by(treat) %>% 
  pivot_longer(-c(treat)) %>%
  ggplot(aes(x=treat, y=value)) +
  facet_wrap(~ name, nrow = 3, scales = "free") +
  geom_boxplot() +
  geom_jitter() +
  theme_classic() +
  labs(title = "AES", x = "Treatment", y = "Value")

```

This is the number of reads (n_seqs), the rarefied richness (rare), and 
the observed richness (sobs for "species observed") for the four treatments:
generations 2 and 5 and selection treatments neutral and positive. As you can
see, there were more sequences from G2 than G5 and also greater richness
in G2 than G5. However, that richness difference is robust to rarefying
to the level of the shallowest sample. There doesn't seem to be a systemic
difference in sequencing depth or richness between the neutral and positive 
treatments in either generation.

## Richness

### Plot

```{r alpha-diversity-plot}

# aes_data %>% 
#   group_by(sample) %>% 
#   summarize(n_seqs = sum(count)) %>% 
#   summarize(min = min(n_seqs))

aes_df <-
  aes_data %>% 
  select(sample, asv, count) %>%
  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
  as.data.frame()

rownames(aes_df) <- aes_df$sample
aes_df <- aes_df[, -1]

rare <- 
  rarefy(aes_df, seq_range[1]) %>% 
  as_tibble(rownames = "sample") %>% 
  rename(rare = value)

aes_data %>% 
  inner_join(rare, by = 'sample') %>% 
  select(sample, passage_treat, rare) %>% 
  distinct %>% 
  ggplot(aes(x = passage_treat, y = rare)) +
  geom_boxplot() +
  geom_jitter() +
  labs(x = "Passage x Treatment", y = "Rarefied Richness") +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())

```

```{r compare_richness_methods}

aes_df_long <-
  aes_df %>% 
  as_tibble(rownames = "sample") %>% 
  pivot_longer(starts_with("ASV")) %>%     
  group_by(sample) %>% 
  uncount(value)

subsample <- function(data, sample_size) {
  data %>% 
    sample_n(sample_size) %>% 
    summarize(s = n_distinct(name))
}

empirical <- map_dfr(1:10, ~subsample(aes_df_long, seq_range[1]), .id = "iters")
empirical <-
empirical %>% 
  group_by(sample) %>% 
  summarize(s = mean(s))

inner_join(rare, empirical, by = "sample") %>% 
  ggplot(aes(x = rare, y = s)) +
  geom_abline(intercept = 0, slope = 1, color = "gray") +
  geom_point()
```

This plot shows the correlation between Hurlbert's (1971) approach to
richness estimation (rare) and empirically rarefying 10 times and taking
the mean richness (s).

### Hypothesis Test

```{r alpha-diversity-test}

rare_data <-
  aes_data %>% 
  select(sample, passage_treat, passage, treat, ch4) %>% 
  distinct %>% 
  inner_join(rare, by = "sample")

kt <- kruskal.test(rare ~ passage_treat, data = rare_data)
kt <- tidy(kt)

wt <- pairwise.wilcox.test(rare_data$rare, 
                           rare_data$passage_treat, 
                           p.adjust.method = "BH")
wt <- tidy(wt)

rare_median <-
  rare_data %>% 
  group_by(passage) %>% 
  summarize(median = median(rare),
            IQR = quart(rare)[2] - quart(rare)[1])

```

There is an effect of passage x treatment on richness (Kruskal-Wallace test: chi-squared = 
`r kt[1,1]`, df = `r kt[1,3]`, p = `r kt[1, 2]`). The pairwise test shows that
G2 is different from G5, but N and P are not different in either generation
(Wilcoxon/Mann-Whitney Test).

```{r}

kable(wt)

```

```{r alpha-diversity-ch4-test}

s <- cor.test(~ ch4 + rare, data = rare_data, method = "spearman", alternative = "two.sided") %>% 
  tidy

s5 <-
  rare_data %>% 
  filter(passage == "G5") %>% 
  cor.test(~ ch4 + rare, data = ., method = "spearman", alternative = "two.sided")

min_seqs <- paste(substr(as.character(min_seqs), 1, 3), substr(as.character(min_seqs), 4, 6), sep = ",")
save(kt, wt, rare_median, s, min_seqs, file = "Output/richness_models.RData")

```

A Spearman's rank correlation test shows no correlation between
methane oxidation rate and richness (Spearman's rho = `r s[1,1]`, p = `r s[1,3]`).

# Beta Diversity

```{r bray-curtis-nmds, messages=FALSE}

dist <- vegdist(aes_df, method="bray")

if(file.exists("Output/rare_dist_bray.rds")) {
  rare_dist_bray <- 
    readRDS("Output/rare_dist_bray.rds")
} else {
  rare_dist_bray <- avgdist(aes_df, dmethod="bray", sample=seq_range[1])
  rare_dist_bray %>% 
    saveRDS("Output/rare_dist_bray.rds")
}

nmds <- metaMDS(rare_dist_bray)

nmds_scores <-
  scores(nmds, display = "sites") %>% 
  as_tibble(rownames = "sample")

nmds_data <-
  aes_data %>% 
  select(sample, passage, treat) %>% 
  distinct %>% 
  inner_join(nmds_scores, by = "sample")

centroids <-
  nmds_data %>% 
  group_by(passage, treat) %>% 
  summarize(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2),
            .groups = "drop")
p_beta <-
  nmds_data %>%   
  ggplot(aes(x = NMDS1, y = NMDS2, shape = passage, color = treat)) +
  geom_point() +
  stat_ellipse(show.legend = FALSE) +
  coord_fixed() +
  scale_shape_discrete(name = "Passage") +
  scale_color_manual(name = "Treatment",
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2')) #+
  # theme_bw() +
  # theme(
  #   plot.background = element_blank(),
  #   panel.grid.major = element_blank(),
  #   panel.grid.minor = element_blank())

save(p_beta, file = "Output/beta_fig.RData")

ggsave('Manuscript/fig3.jpg', p_beta, height = 2.5, width = 4)

```

## Beta diversity test passage by treatment

```{r permanova-passage-treatment, warnings=FALSE}

dist_df <-
  rare_dist_bray %>% 
  as.matrix %>% 
  as_tibble(rownames = "sample")

meta_dist <-
  aes_data %>% 
  select(sample, passage_treat, passage, treat, ch4) %>% 
  distinct %>% 
  inner_join(dist_df, by = "sample")

permanova <- adonis2(rare_dist_bray ~ passage * treat, data = meta_dist, permutations = 999)

betadisper(rare_dist_bray, meta_dist$passage_treat) %>% 
  permutest()

tidy_permanova <- 
  tidy(permanova)

save(tidy_permanova, file = "Output/beta_model.RData")

```

There is a significant difference of centroid between passage, treatment, and
their interaction. Testing a difference of dispersion using betadisper
shows no difference in variance between the four groups.

# Distance-based RDA Methane

```{r full-dbrda}

meta_dist %>% 
  select(passage_treat, ch4) %>% 
  mutate(log_ch4 = log_methane(ch4)) %>% 
  ggplot(aes(x = passage_treat, y = log_ch4)) +
  geom_boxplot() +
  theme_classic() +
  labs(x = "Passage x Treatment", y = "Log10 Methane")

dbrda_ch4 <- dbrda(rare_dist_bray ~ rank(ch4), 
                      data = meta_dist, 
                      na.action = "na.omit",
                      add = TRUE)

dbrda_ch4 %>% 
  permutest(permutations = 9999)

percent_explained <- 100 * eigenvals(dbrda_ch4) / sum(eigenvals(dbrda_ch4))

dbrda_scores <-
  scores(dbrda_ch4, display = "sites") %>% 
  as_tibble(rownames = "sample")

dbrda_data <-
  aes_data %>% 
  select(sample, passage, treat, ch4) %>% 
  distinct %>% 
  inner_join(dbrda_scores, by = "sample")

dbrda_data %>%   
  ggplot(aes(x = dbRDA1, y = MDS1, shape = passage, color = treat)) +
  geom_point() +
  stat_ellipse(show.legend = FALSE) +
  coord_fixed() +
  scale_shape_discrete(name = "Passage") +
  scale_color_manual(name = "Treatment",
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2')) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())

```

```{r pre-dbrda-data-subsets}

# aes_data %>% 
#   filter(treat == "P") %>% 
#   group_by(asv) %>% 
#   mutate(total = sum(count)) %>% 
#   filter(total != 0) %>% 
#   ungroup() %>% 
#   select(-total) %>% 
#   group_by(sample) %>% 
#   summarize(n_seqs = sum(count)) %>% 
#   summarize(min = min(n_seqs))

treatP_df <-
  aes_data %>%  
  filter(treat == "P") %>% 
  group_by(asv) %>% 
  mutate(total = sum(count)) %>% 
  filter(total != 0) %>% 
  ungroup() %>% 
  select(-total) %>% 
  select(sample, asv, count) %>%
  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
  as.data.frame()

rownames(treatP_df) <- treatP_df$sample
treatP_df <- treatP_df[, -1]

# aes_data %>% 
#   filter(passage == "G5") %>% 
#   group_by(asv) %>% 
#   mutate(total = sum(count)) %>% 
#   filter(total != 0) %>% 
#   ungroup() %>% 
#   select(-total) %>% 
#   group_by(sample) %>% 
#   summarize(n_seqs = sum(count)) %>% 
#   summarize(min = min(n_seqs))

passage2_df <-
  aes_data %>%  
  filter(passage == "G2") %>% 
  group_by(asv) %>% 
  mutate(total = sum(count)) %>% 
  filter(total != 0) %>% 
  ungroup() %>% 
  select(-total) %>% 
  select(sample, asv, count) %>%
  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
  as.data.frame()

rownames(passage2_df) <- passage2_df$sample
passage2_df <- passage2_df[, -1]

passage5_df <-
  aes_data %>%  
  filter(passage == "G5") %>% 
  group_by(asv) %>% 
  mutate(total = sum(count)) %>% 
  filter(total != 0) %>% 
  ungroup() %>% 
  select(-total) %>% 
  select(sample, asv, count) %>%
  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
  as.data.frame()

rownames(passage5_df) <- passage5_df$sample
passage5_df <- passage5_df[, -1]

```

```{r treat-dbrda}

if(file.exists("Output/treat_rare_dist_bray.rds")) {
  treat_rare_dist_bray <- 
    readRDS("Output/treat_rare_dist_bray.rds")
} else {
treat_rare_dist_bray <- avgdist(treatP_df, dmethod="bray", sample=seq_range[1])
  treat_rare_dist_bray %>% 
    saveRDS("Output/treat_rare_dist_bray.rds")
}

treat_dist_df <-
  treat_rare_dist_bray %>% 
  as.matrix %>% 
  as_tibble(rownames = "sample")

treat_meta_dist <-
  aes_data %>% 
  select(sample, passage_treat, passage, treat, ch4) %>% 
  distinct %>% 
  inner_join(treat_dist_df, by = "sample")

treat_dbrda_ch4 <- dbrda(treat_rare_dist_bray ~ rank(ch4), 
                      data = treat_meta_dist, 
                      na.action = "na.omit",
                      add = TRUE)

treat_dbrda_ch4 %>% 
  permutest(permutations = 9999)

percent_explained <- 100 * eigenvals(treat_dbrda_ch4) / sum(eigenvals(treat_dbrda_ch4))

treat_dbrda_scores <-
  scores(treat_dbrda_ch4, display = "sites") %>% 
  as_tibble(rownames = "sample")

treat_dbrda_data <-
  aes_data %>% 
  select(sample, passage, treat, ch4) %>% 
  distinct %>% 
  inner_join(treat_dbrda_scores, by = "sample")

treat_dbrda_data %>%   
  ggplot(aes(x = dbRDA1, y = MDS1, shape = passage, color = treat)) +
  geom_point() +
  stat_ellipse(show.legend = FALSE) +
  coord_fixed() +
  scale_shape_discrete(name = "Passage") +
  scale_color_manual(name = "Treatment",
                       labels = c('Positive'),
                       values = c('darkorange2')) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())

```

```{r passage2-dbrda}
if(file.exists("Output/passage2_rare_dist_bray.rds")) {
  passage2_rare_dist_bray <- 
    readRDS("Output/passage2_rare_dist_bray.rds")
} else {
passage2_rare_dist_bray <- avgdist(passage2_df, dmethod="bray", sample=seq_range[1])
  passage2_rare_dist_bray %>% 
    saveRDS("Output/passage2_rare_dist_bray.rds")
}

passage2_dist_df <-
  passage2_rare_dist_bray %>% 
  as.matrix %>% 
  as_tibble(rownames = "sample")

passage2_meta_dist <-
  aes_data %>% 
  select(sample, passage_treat, passage, treat, ch4) %>% 
  distinct %>% 
  inner_join(passage2_dist_df, by = "sample")

passage2_dbrda_ch4 <- dbrda(passage2_rare_dist_bray ~ rank(ch4), 
                      data = passage2_meta_dist, 
                      na.action = "na.omit",
                      add = TRUE)

passage2_dbrda_permtest <- 
passage2_dbrda_ch4 %>% 
  permutest(permutations = 9999)

passage2_dbrda_permtest <- 
passage2_dbrda_ch4 %>% 
  anova(permutations = 9999) %>% 
  tidy()

passage2_percent_explained <- 100 * eigenvals(passage2_dbrda_ch4) / sum(eigenvals(passage2_dbrda_ch4))

passage2_dbrda_scores <-
  scores(passage2_dbrda_ch4, display = "sites") %>% 
  as_tibble(rownames = "sample")

passage2_dbrda_data <-
  aes_data %>% 
  select(sample, passage, treat, ch4) %>% 
  distinct %>% 
  inner_join(passage2_dbrda_scores, by = "sample")

passage2_dbrda_data %>%   
  ggplot(aes(x = dbRDA1, y = MDS1, shape = passage, color = treat)) +
  geom_point() +
  stat_ellipse(show.legend = FALSE) +
  coord_fixed() +
  scale_shape_discrete(name = "Passage") +
  scale_color_manual(name = "Treatment",
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2')) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())

```

```{r passage5-dbrda}
if(file.exists("Output/passage5_rare_dist_bray.rds")) {
  passage5_rare_dist_bray <- 
    readRDS("Output/passage5_rare_dist_bray.rds")
} else {
passage5_rare_dist_bray <- avgdist(passage5_df, dmethod="bray", sample=seq_range[1])
  passage5_rare_dist_bray %>% 
    saveRDS("Output/passage5_rare_dist_bray.rds")
}

passage5_dist_df <-
  passage5_rare_dist_bray %>% 
  as.matrix %>% 
  as_tibble(rownames = "sample")

passage5_meta_dist <-
  aes_data %>% 
  select(sample, passage_treat, passage, treat, ch4) %>% 
  distinct %>% 
  inner_join(passage5_dist_df, by = "sample")

passage5_dbrda_ch4 <- dbrda(passage5_rare_dist_bray ~ rank(ch4), 
                      data = passage5_meta_dist, 
                      na.action = "na.omit",
                      add = TRUE)

passage5_dbrda_permtest <-
passage5_dbrda_ch4 %>% 
  anova(permutations = 9999) %>% 
  tidy

passage5_percent_explained <- 100 * eigenvals(passage5_dbrda_ch4) / sum(eigenvals(passage5_dbrda_ch4))

passage5_dbrda_scores <-
  scores(passage5_dbrda_ch4, display = "sites") %>% 
  as_tibble(rownames = "sample")

passage5_dbrda_data <-
  aes_data %>% 
  select(sample, passage, treat, ch4) %>% 
  distinct %>% 
  inner_join(passage5_dbrda_scores, by = "sample")

passage5_dbrda_data %>%   
  ggplot(aes(x = dbRDA1, y = MDS1, shape = passage, color = treat)) +
  geom_point() +
  stat_ellipse(show.legend = FALSE) +
  coord_fixed() +
  scale_shape_discrete(name = "Passage") +
  scale_color_manual(name = "Treatment",
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2')) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())

```

```{r}
save(passage2_dbrda_permtest, 
     passage5_dbrda_permtest, 
     passage5_percent_explained,
     file = "Output/dbrda.RData")
```

The overall dbRDA is significant and rank of methane explains 5% of the variation.
Subsetting the Positive treatment in both G2 and G5, it is not significant: no
correlation between dissimilarity and rank of methane between G2 and G5. This
makes sense, because most of that variation would be due to selection in the jar.
However, subsetting passage G5 and comparing N and P, 10% of dissimilarity is
explained by rank of CH4.

# Differential Abundance

## Unique to each treatment

### Family

```{r}
g5n <- 
  aes_data %>% 
  filter(passage_treat == "G5N") %>% 
  select(Family) %>% 
  distinct() %$%
  Family

g5p <- 
  aes_data %>% 
  filter(passage_treat == "G5P") %>% 
  select(Family) %>% 
  distinct() %$%
  Family

p_not_n <- g5p[!g5p %in% g5n]

n_not_p <- g5n[!g5n %in% g5p]

n_not_p_prevalence <-
  aes_data %>% 
  filter(passage_treat == "G5N") %>% 
  filter(Family %in% n_not_p) %>% 
  select(sample, Family, count) %>% 
  mutate(prevalence = as.numeric(count > 0)) %>% 
  select(-count) %>% 
  group_by(Family) %>% 
  summarize(prevalence = sum(prevalence))

n_not_p_prevalence %>% 
  ggplot(aes(x = prevalence)) +
  geom_histogram() +
  scale_x_continuous(breaks = 1:12) +
  theme_classic() +
  labs(title = "Unique to Positive")

# n_not_p_prevalence %>%
#   filter(prevalence == 1) %>% 
#   nrow

# n_not_p_prevalence %>%
#   filter(prevalence == 2) %>% 
#   nrow
# n_not_p_prevalence %>%
#   filter(prevalence > 2) %$%
#   Family

p_not_n_prevalence <-
  aes_data %>% 
  filter(passage_treat == "G5P") %>% 
  filter(Family %in% p_not_n) %>% 
  select(sample, Family, count) %>% 
  mutate(prevalence = as.numeric(count > 0)) %>% 
  select(-count) %>% 
  group_by(Family) %>% 
  summarize(prevalence = sum(prevalence))

p_not_n_prevalence %>% 
  ggplot(aes(x = prevalence)) +
  geom_histogram() +
  scale_x_continuous(breaks = 1:12) +
  theme_classic() +
  labs(title = "Unique to Neutral")
# 
# p_not_n_prevalence %>%
#   filter(prevalence == 1) %>% 
#   nrow
# 
# p_not_n_prevalence %>%
#   filter(prevalence == 2) %>% 
#   nrow
p_not_n_over_6 <- p_not_n_prevalence %>%
  filter(prevalence > 6)

```

First, I looked at which Families are unique to each treatment in passage 5.
There were `r length(n_not_p)` families unique to N and `r length(p_not_n)`
unique to P. However, of the  `r length(n_not_p)` unique to P, 
`r n_not_p_prevalence %>% filter(prevalence == 1) %>% nrow` had a prevalence
of 1 and `r n_not_p_prevalence %>% filter(prevalence == 2) %>% nrow` had
a prevalence of 2 leaving only a single family, which had a prevalence of
`r n_not_p_prevalence %>% filter(prevalence > 2) %$% prevalence`, which was `r n_not_p_prevalence %>% filter(prevalence > 2) %$% Family`.
Of the `r length(p_not_n)` Families unique to P, 2 had prevalences greater than
6, which were `r p_not_n_over_6$Family[1]` with a prevalence of
`r p_not_n_over_6$prevalence[1]` and `r p_not_n_over_6$Family[2]` with
a prevalence of `r p_not_n_over_6$prevalence[2]`.

### Genus

```{r}
g5n <- 
  aes_data %>% 
  filter(passage_treat == "G5N") %>% 
  select(Genus) %>% 
  distinct() %$%
  Genus

g5p <- 
  aes_data %>% 
  filter(passage_treat == "G5P") %>% 
  select(Genus) %>% 
  distinct() %$%
  Genus

p_not_n <- g5p[!g5p %in% g5n]

n_not_p <- g5n[!g5n %in% g5p]

n_not_p_prevalence <-
  aes_data %>% 
  filter(passage_treat == "G5N") %>% 
  filter(Genus %in% n_not_p) %>% 
  select(sample, Genus, count) %>% 
  mutate(prevalence = as.numeric(count > 0)) %>% 
  select(-count) %>% 
  group_by(Genus) %>% 
  summarize(prevalence = sum(prevalence))

n_not_p_prevalence %>% 
  ggplot(aes(x = prevalence)) +
  geom_histogram() +
  scale_x_continuous(breaks = 1:12) +
  theme_classic() +
  labs(title = "Unique to Positive")

p_not_n_prevalence <-
  aes_data %>% 
  filter(passage_treat == "G5P") %>% 
  filter(Genus %in% p_not_n) %>% 
  select(sample, Genus, count) %>% 
  mutate(prevalence = as.numeric(count > 0)) %>% 
  select(-count) %>% 
  group_by(Genus) %>% 
  summarize(prevalence = sum(prevalence))

p_not_n_prevalence %>% 
  ggplot(aes(x = prevalence)) +
  geom_histogram() +
  scale_x_continuous(breaks = 1:12) +
  theme_classic() +
  labs(title = "Unique to Neutral")

p_not_n_over_6 <- p_not_n_prevalence %>%
  filter(prevalence > 6)

n_not_p_over_6 <- n_not_p_prevalence %>%
  filter(prevalence > 6)

```

### Species

```{r}
g5n <- 
  aes_data %>% 
  filter(passage_treat == "G5N") %>% 
  select(Species) %>% 
  distinct() %$%
  Species

g5p <- 
  aes_data %>% 
  filter(passage_treat == "G5P") %>% 
  select(Species) %>% 
  distinct() %$%
  Species

p_not_n <- g5p[!g5p %in% g5n]

n_not_p <- g5n[!g5n %in% g5p]

n_not_p_prevalence <-
  aes_data %>% 
  filter(passage_treat == "G5N") %>% 
  filter(Species %in% n_not_p) %>% 
  select(sample, Species, count) %>% 
  mutate(prevalence = as.numeric(count > 0)) %>% 
  select(-count) %>% 
  group_by(Species) %>% 
  summarize(prevalence = sum(prevalence))

n_not_p_prevalence %>% 
  ggplot(aes(x = prevalence)) +
  geom_histogram() +
  scale_x_continuous(breaks = 1:12) +
  theme_classic() +
  labs(title = "Unique to Positive")

p_not_n_prevalence <-
  aes_data %>% 
  filter(passage_treat == "G5P") %>% 
  filter(Species %in% p_not_n) %>% 
  select(sample, Species, count) %>% 
  mutate(prevalence = as.numeric(count > 0)) %>% 
  select(-count) %>% 
  group_by(Species) %>% 
  summarize(prevalence = sum(prevalence))

p_not_n_prevalence %>% 
  ggplot(aes(x = prevalence)) +
  geom_histogram() +
  scale_x_continuous(breaks = 1:12) +
  theme_classic() +
  labs(title = "Unique to Neutral")

p_not_n_over_6 <- p_not_n_prevalence %>%
  filter(prevalence > 3)

n_not_p_over_6 <- n_not_p_prevalence %>%
  filter(prevalence > 3)

```

## Setup Differential Abundance Data

### Family

```{r setup-differential-abundance-data-family}

aes_G5 <-
  aes_data %>% 
  filter(passage == "G5")

aes_family_G5 <-
  aes_G5 %>% 
  select(asv, sample, treat, count, Family) %>%
  group_by(sample, Family) %>% 
  mutate(total = sum(count)) %>% 
  ungroup() %>% 
  select(sample, treat, Family, total) %>% 
  distinct()

prevalent_families <- 
  aes_family_G5 %>% 
  group_by(Family, treat) %>% 
  summarize(prev = sum(total > 0), .groups = "drop") %>% 
  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
  filter(N/12 > 0.1 & P/12 > 0.1) %$%
  Family

otu_table <- 
  aes_family_G5 %>% 
  filter(Family %in% prevalent_families) %>% 
  select(sample, Family, total) %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

rownames(otu_table) <- otu_table[,1]
otu_table <- otu_table[,-1]

sample_data <- aes_family_G5 %>% 
  select(sample, treat) %>% 
  distinct() %>% 
  as.data.frame()

rownames(sample_data) <- sample_data[,1]

```

#### Corncob

```{r}

da_analysis <- differentialTest(formula = ~ treat,
                                phi.formula = ~ 1,
                                formula_null = ~ 1,
                                phi.formula_null = ~ 1,
                                test = "Wald", 
                                boot = FALSE,
                                data = otu_table,
                                sample_data = sample_data,
                                fdr_cutoff = 0.05, B = 100,
                                taxa_are_rows = TRUE)

dt_data <- plot(da_analysis, 
                # level = c("Family"), 
                data_only = TRUE)

dt_plot <- function(dt_data) {
  ggplot(dt_data, aes(x = x, xmin = xmin, xmax = xmax, y = reorder(taxa, x))) +
    geom_pointrange() +
    geom_vline(xintercept = 0) +
    labs(x = "Estimate", y = NULL) +
    coord_cartesian(clip="off") +
    # theme(panel.border = element_blank()) +
    theme_classic() +
      theme(axis.text.y = element_markdown())

}

dt_data %>%
  mutate(taxa = str_replace(taxa, "(.*)", "*\\1*"),
         taxa = str_replace(taxa, "\\*(.*)_unclassified\\*",
                                "Unclassified *\\1*")) %>% 
  dt_plot()

```

#### ANCOM

```{r}

# Setup phyloseq object
aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE), sample_data(sample_data))

if(file.exists("Output/ancom_out.rds")) {
  out <- 
    readRDS("Output/ancom_out.rds")
} else {
  # Run Ancom
  out = ancom(phyloseq = aes_ps,  
              p_adj_method = "BH",
              prv_cut = 0.1, 
              lib_cut = 0, 
              main_var = "treat",
              rand_formula = NULL, 
              lme_control = NULL,
              struc_zero = TRUE, 
              neg_lb = FALSE, 
              alpha = 0.05, 
              n_cl = 1)  
  out %>% 
    saveRDS("Output/ancom_out.rds")
}

res = out$res
q_val = out$q_data
beta_val = out$beta_data
# Only consider the effect sizes with the corresponding q-value less than alpha
beta_val = beta_val * (q_val < 0.05) 
# Choose the maximum of beta's as the effect size
beta_pos = apply(abs(beta_val), 2, which.max) 
beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
                  FUN.VALUE = double(1))
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(out$zero_ind), 
                nrow(feature_table), 
                sum(apply(out$zero_ind, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = 0.7 * (n_taxa - 1)

df_fig_w = res %>%
  dplyr::mutate(beta = beta_max,
                direct = case_when(
                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
                  TRUE ~ "Not Significant"
                  )) %>%
  dplyr::arrange(W)
df_fig_w$taxon_id = factor(df_fig_w$taxon_id, levels = df_fig_w$taxon_id)
df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
df_fig_w$direct = factor(df_fig_w$direct, 
                     levels = c("Negative", "Positive", "Not Significant"))

p_w = df_fig_w %>%
  ggplot(aes(x = taxon_id, y = W, color = direct)) +
  geom_point(size = 2, alpha = 0.6) +
  labs(x = "Taxon", y = "W") +
  scale_color_discrete(name = NULL) + 
  geom_hline(yintercept = cut_off, linetype = "dotted", 
             color = "blue", size = 1.5) +
  geom_text(aes(x = 2, y = cut_off + 0.5, label = "W[0.7]"), 
            size = 5, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank())
p_w

```

#### Aldex2

```{r}

conds <- colnames(otu_table) %>% 
  substr(3, 3)

if(file.exists("Output/aldex_x_all.rds")) {
  x.all <- 
    readRDS("Output/aldex_x_all.rds")
} else {
  x.all <- aldex(otu_table, 
                 conds, mc.samples=1000, test="t", effect=TRUE,
                 include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
  x.all %>% 
    saveRDS("Output/aldex_x_all.rds")
}

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
    ylab="Difference")
aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
    ylab="Difference")

par(mfrow=c(1,2))
plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")
legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))

plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")

x.all_effect_1 <-
  x.all %>% 
  filter(abs(effect) > 1) %>% 
  as_tibble(rownames = "Family")

# x.all_effect_1 %>% 
#   # inner_join(taxa_table, by = "asv")
#   ggplot(aes(y = reorder(Family, effect), x = effect)) +
#   geom_point()

# x.all_effect_1 %>% 
#   select(Family, effect)

```

#### Comparison

##### All Significant

```{r}

# CORNCOB
corncob_taxa <- dt_data$taxa

# ANCOM
ancom_taxa <- as.character(df_fig_w[df_fig_w$detected_0.7, ]$taxon_id)

# ALDEX2
aldex_taxa <- x.all_effect_1$Family

at_least_one_taxa <-
	aes_family_G5 %>%
	group_by(sample) %>%
	mutate(rel_abund = (total / sum(total)) * 100) %>%
	filter(Family %in% ancom_taxa | Family %in% aldex_taxa | Family %in% corncob_taxa) %>%
	mutate(Family = str_replace(Family, "(.*)", "*\\1*"),
	       Family = str_replace(Family, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))

at_least_one_iqm <- 
	at_least_one_taxa %>%
	group_by(treat, Family) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")

sort_order <-
	at_least_one_taxa %>%
	group_by(Family) %>%
	summarize(median = median(rel_abund)) %>%
	arrange(median)

at_least_one_taxa$Family <- factor(at_least_one_taxa$Family, levels = sort_order$Family)

at_least_one_taxa %>%
	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
			data = at_least_one_iqm, 
			position = position_dodge(width = 0.75), 
			shape = 21, 
			color = "black") +
	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Relative Abundance (%)", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2'))

```

##### Consensus Taxa

```{r}
aldex_effect <-
  x.all_effect_1 %>% 
  select(Family, effect)

all_taxa <-
	aes_family_G5 %>%
	group_by(sample) %>%
	mutate(rel_abund = (total / sum(total)) * 100) %>%
  left_join(aldex_effect, by = "Family") %>% 
	filter(Family %in% ancom_taxa & Family %in% aldex_taxa & Family %in% corncob_taxa) %>% 
  ungroup()

da_fams <-
  all_taxa %>% 
  select(Family) %>% 
  distinct() %$%
  Family

all_taxa <-
  all_taxa %>%
	mutate(Family = str_replace(Family, "(.*)", "*\\1*"),
	       Family = str_replace(Family, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))
all_taxa$Family[all_taxa$Family == "Unclassified *0319-6G20*"] <- "Uncultured family 0319-6G20"
all_taxa$Family[all_taxa$Family == "Unclassified *Chloroplast*"] <- "Chloroplast"
all_taxa$Family[all_taxa$Family == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
all_taxa$Family[all_taxa$Family == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
all_taxa$Family[all_taxa$Family == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"

all_iqm <- 
	all_taxa %>%
	group_by(treat, Family) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")

sort_order <-
	all_taxa %>%
  select(Family, effect) %>% 
  distinct() %>% 
  arrange(effect)
	# group_by(Family) %>%
	# summarize(median = median(rel_abund)) %>%
	# arrange(median)

all_taxa$Family <- factor(all_taxa$Family, levels = sort_order$Family)
all_iqm$Family <- factor(all_iqm$Family, levels = sort_order$Family)
all_taxa <-
  all_taxa %>% 
  mutate(direction = if_else(effect > 0, 0, 1))
all_iqm <-
all_taxa %>% 
  select(Family, direction) %>% 
  distinct() %>% 
  full_join(all_iqm, by = "Family")
(p <-
	all_taxa %>%
	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
			data = all_iqm, 
			position = position_dodge(width = 0.75), 
			shape = 21, 
			color = "black") +
	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Relative Abundance (%)", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
    facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
theme(
  strip.background = element_blank(),
  strip.text.y = element_blank()
))

(p_log10 <- 
	p +
	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
		      labels = c(0.001, 0.01, 0.1, 1)))

save(p_log10, da_fams, file = "Output/da_fig.RData")

ggsave('Manuscript/fig4.jpg', p_log10, width = 5, height = 5)

## All
# ggsave("shared_taxa.png", p)
ggsave("shared_taxa_log.png", p_log10, width = 5, height = 5)

# aes_family_rclr_plot <-
#   aes_family_rclr %>% 
# 	filter(Family %in% ancom_taxa & Family %in% aldex_taxa & Family %in% corncob_taxa) %>% 
#   pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
#   select(Family, treat, rep, passage, rclr) %>% 
#   distinct() %>% 	
#   mutate(Family = str_replace(Family, "(.*)", "*\\1*"),
# 	       Family = str_replace(Family, "\\*(.*)_unclassified\\*",
# 				    "Unclassified *\\1*"))
# 
# aes_family_rclr_plot$Family[aes_family_rclr_plot$Family == "Unclassified *0319-6G20*"] <- "Uncultured family 0319-6G20"
# aes_family_rclr_plot$Family[aes_family_rclr_plot$Family == "Unclassified *Chloroplast*"] <- "Chloroplast"
# aes_family_rclr_plot$Family <- factor(aes_family_rclr_plot$Family, levels = sort_order$Family)
# ggplot(aes_family_rclr_plot, aes(x = rclr, y = Family, color = treat)) +
#   facet_grid( ~ passage) +
#   geom_point(position = position_jitterdodge())

p_effect <-
	all_taxa %>%
	ggplot(aes(x = effect, y = Family)) +
	geom_point() +

	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Aldex2 Effect Size", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2'))
p_effect

```

## Methanogens and Methanotrophs

Methanotrophy is (mostly) conserved at the family level. The Type I and Type X methanotrophs are in the Gammaproteobacteria in the families _Methylococcaceae_ and _Methylothermaceae_. There are also methanotrophs in the Verrumicrobiota in the family _Methylacidiphilaceae_. The Type II methanotrophs are in the _Alphaproteobacteria_ and are found in the families _Methylocystaceae_ and _Beijerinckiaceae_. However, the _Beijerinckiaceae_ also contain non-methanotrophs including N2-fixers and photoautotrophs.

```{r methanotroph-lists}

methanotrophs <- c('Methylococcaceae', 
                   'Methylothermaceae', 
                   'Methylocystaceae', 
                   'Beijerinckiaceae', 
                   'Methylacidiphilaceae')
methanotrophs <- c(methanotrophs, paste(methanotrophs, "unclassified", sep = "_"))

troph_orders <- c('Methylococcales')

methanogen_classes <- c("Methanococci", "Methanomicrobia", "Methanobacteria")

methanogen_classes <- c(methanogen_classes, paste(methanogen_classes, "unclassified", sep = "_"))

```

```{r}

aes_data %>% 
  select(Family, Genus, count, sample) %>% 
  group_by(sample, Genus, Family) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  group_by(sample) %>% 
  mutate(rel_abund = total / sum(total)) %>% 
  mutate(rclr = log(total/gm(total))) %>% 
  filter(Family %in% methanotrophs) %>% 
  select(-Family) %>% 
  mutate(treat = substr(sample, 3, 3), passage = substr(sample, 1, 2)) %>% 
  pivot_longer(rel_abund:rclr, names_to = "method", values_to = "value") %>% 
  ggplot(aes(x = value, y = Genus, color = treat)) +
  geom_jitter() +
  facet_grid(passage ~ method, scales = "free") +
  theme_bw()

aes_data %>% 
  select(Family, count, sample) %>% 
  group_by(sample, Family) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  group_by(sample) %>% 
  mutate(rel_abund = total / sum(total)) %>% 
  mutate(rclr = log(total/gm(total))) %>% 
  filter(Family %in% methanotrophs) %>% 
  mutate(treat = substr(sample, 3, 3), passage = substr(sample, 1, 2)) %>% 
  pivot_longer(rel_abund:rclr, names_to = "method", values_to = "value") %>% 
  ggplot(aes(x = value, y = Family, color = treat)) +
  geom_jitter() +
  facet_grid(passage ~ method, scales = "free") +
  theme_bw()

troph_asvs <-
  aes_data %>% 
  select(Kingdom:Species, asv) %>% 
  distinct() %>% 
  filter(Family %in% methanotrophs) %$%
  asv

```

Of the five methanotroph families, two are represented in this dataset:
_Methylacidiphilaceae_ and _Beijerinckiaceae_. From these two families,
there are eleven genera represented. These are spread across
`r length(troph_asvs)` ASVs

```{r}

aes_data %>% 
  select(Class, count, sample) %>% 
  group_by(sample, Class) %>% 
  mutate(total = sum(count)) %>% 
  select(-count) %>% 
  distinct() %>% 
  filter(Class %in% methanogen_classes)

```

A grand total of 10 reads map to methanogen classes across the entire dataset,
including 2 reads in Methanobacteria (sample G2N02) and 8 reads of
Methanomicrobia (sample G2N07).

### Methanotroph differential abundance

```{r}

troph_data <-
  aes_data %>% 
  select(sample, asv, treat, passage, count, Genus, Family) %>% 
  filter(passage == "G5") %>% 
  select(-passage) %>% 
  filter(asv %in% troph_asvs)

troph_data_family <-
  troph_data %>% 
  select(sample, Family, count) %>% 
  group_by(sample, Family) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

rownames(troph_data_family) <- troph_data_family[, 1]
troph_data_family <- troph_data_family[, -1]

troph_data_genus <-
  troph_data %>% 
  select(sample, Genus, count) %>% 
  group_by(sample, Genus) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

rownames(troph_data_genus) <- troph_data_genus[, 1]
troph_data_genus <- troph_data_genus[, -1]

sample_data <- 
  troph_data %>% 
  select(sample, treat) %>% 
  distinct() %>% 
  as.data.frame()
conds <- sample_data[, 2]

```

```{r corncob-methanotrophs}
if(file.exists("Output/cc_family.rds")) {
  cc_family <- 
    readRDS("Output/cc_family.rds")
} else {
cc_family <- differentialTest(formula = ~ treat,
                                phi.formula = ~ 1,
                                formula_null = ~ 1,
                                phi.formula_null = ~ 1,
                                test = "Wald", 
                                boot = TRUE,
                                data = troph_data_family,
                                sample_data = sample_data,
                                fdr_cutoff = 0.05, B = 1000)  
cc_family %>% 
    saveRDS("Output/cc_family.rds")
}

plot(cc_family)

if(file.exists("Output/cc_genus.rds")) {
  cc_genus <- 
    readRDS("Output/cc_genus.rds")
} else {
cc_genus <- differentialTest(formula = ~ treat,
                                phi.formula = ~ 1,
                                formula_null = ~ 1,
                                phi.formula_null = ~ 1,
                                test = "Wald", 
                                boot = TRUE,
                                data = troph_data_genus,
                                sample_data = sample_data,
                                fdr_cutoff = 0.05, B = 1000) 
cc_genus %>% 
    saveRDS("Output/cc_genus.rds")
}

plot(cc_genus)

```

```{r}

knitr::opts_chunk$set(eval = FALSE)

```

### Genus

```{r setup-differential-abundance-data-genus}

aes_G5 <-
  aes_data %>% 
  filter(passage == "G5")

aes_Genus_G5 <-
  aes_G5 %>% 
  select(asv, sample, treat, count, Genus) %>%
  group_by(sample, Genus) %>% 
  mutate(total = sum(count)) %>% 
  ungroup() %>% 
  select(sample, treat, Genus, total) %>% 
  distinct()

prevalent_families <- 
  aes_Genus_G5 %>% 
  group_by(Genus, treat) %>% 
  summarize(prev = sum(total > 0), .groups = "drop") %>% 
  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
  filter(N/12 > 0.1 & P/12 > 0.1) %$%
  Genus

otu_table <- 
  aes_Genus_G5 %>% 
  filter(Genus %in% prevalent_families) %>% 
  select(sample, Genus, total) %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

rownames(otu_table) <- otu_table[,1]
otu_table <- otu_table[,-1]

sample_data <- aes_Genus_G5 %>% 
  select(sample, treat) %>% 
  distinct() %>% 
  as.data.frame()

rownames(sample_data) <- sample_data[,1]

```

#### Corncob

```{r}

da_analysis <- differentialTest(formula = ~ treat,
                                phi.formula = ~ 1,
                                formula_null = ~ 1,
                                phi.formula_null = ~ 1,
                                test = "Wald", 
                                boot = FALSE,
                                data = otu_table,
                                sample_data = sample_data,
                                fdr_cutoff = 0.05, B = 100,
                                taxa_are_rows = TRUE)

dt_data <- plot(da_analysis, 
                # level = c("Genus"), 
                data_only = TRUE)

dt_plot <- function(dt_data) {
  ggplot(dt_data, aes(x = x, xmin = xmin, xmax = xmax, y = reorder(taxa, x))) +
    geom_pointrange() +
    geom_vline(xintercept = 0) +
    labs(x = "Estimate", y = NULL) +
    coord_cartesian(clip="off") +
    # theme(panel.border = element_blank()) +
    theme_classic() +
      theme(axis.text.y = element_markdown())

}

dt_data %>%
  mutate(taxa = str_replace(taxa, "(.*)", "*\\1*"),
         taxa = str_replace(taxa, "\\*(.*)_unclassified\\*",
                                "Unclassified *\\1*")) %>% 
  dt_plot()

```

#### ANCOM

```{r}

# Setup phyloseq object
aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE), sample_data(sample_data))

if(file.exists("Output/ancom_out.rds")) {
  out <- 
    readRDS("Output/ancom_out.rds")
} else {
  # Run Ancom
  out = ancom(phyloseq = aes_ps,  
              p_adj_method = "BH",
              prv_cut = 0.1, 
              lib_cut = 0, 
              main_var = "treat",
              rand_formula = NULL, 
              lme_control = NULL,
              struc_zero = TRUE, 
              neg_lb = FALSE, 
              alpha = 0.05, 
              n_cl = 1)  
  out %>% 
    saveRDS("Output/ancom_out.rds")
}

res = out$res
q_val = out$q_data
beta_val = out$beta_data
# Only consider the effect sizes with the corresponding q-value less than alpha
beta_val = beta_val * (q_val < 0.05) 
# Choose the maximum of beta's as the effect size
beta_pos = apply(abs(beta_val), 2, which.max) 
beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
                  FUN.VALUE = double(1))
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(out$zero_ind), 
                nrow(feature_table), 
                sum(apply(out$zero_ind, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = 0.7 * (n_taxa - 1)

df_fig_w = res %>%
  dplyr::mutate(beta = beta_max,
                direct = case_when(
                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
                  TRUE ~ "Not Significant"
                  )) %>%
  dplyr::arrange(W)
df_fig_w$taxon_id = factor(df_fig_w$taxon_id, levels = df_fig_w$taxon_id)
df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
df_fig_w$direct = factor(df_fig_w$direct, 
                     levels = c("Negative", "Positive", "Not Significant"))

p_w = df_fig_w %>%
  ggplot(aes(x = taxon_id, y = W, color = direct)) +
  geom_point(size = 2, alpha = 0.6) +
  labs(x = "Taxon", y = "W") +
  scale_color_discrete(name = NULL) + 
  geom_hline(yintercept = cut_off, linetype = "dotted", 
             color = "blue", size = 1.5) +
  geom_text(aes(x = 2, y = cut_off + 0.5, label = "W[0.7]"), 
            size = 5, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank())
p_w

```

#### Aldex2

```{r}

conds <- colnames(otu_table) %>% 
  substr(3, 3)

if(file.exists("Output/aldex_x_all.rds")) {
  x.all <- 
    readRDS("Output/aldex_x_all.rds")
} else {
  x.all <- aldex(otu_table, 
                 conds, mc.samples=1000, test="t", effect=TRUE,
                 include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
  x.all %>% 
    saveRDS("Output/aldex_x_all.rds")
}

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
    ylab="Difference")
aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
    ylab="Difference")

par(mfrow=c(1,2))
plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")
legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))

plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")

x.all_effect_1 <-
  x.all %>% 
  filter(abs(effect) > 1) %>% 
  as_tibble(rownames = "Genus")

# x.all_effect_1 %>% 
#   # inner_join(taxa_table, by = "asv")
#   ggplot(aes(y = reorder(Genus, effect), x = effect)) +
#   geom_point()

# x.all_effect_1 %>% 
#   select(Genus, effect)

```

#### Comparison

##### All Significant

```{r}

# CORNCOB
corncob_taxa <- dt_data$taxa

# ANCOM
ancom_taxa <- as.character(df_fig_w[df_fig_w$detected_0.7, ]$taxon_id)

# ALDEX2
aldex_taxa <- x.all_effect_1$Genus

at_least_one_taxa <-
	aes_Genus_G5 %>%
	group_by(sample) %>%
	mutate(rel_abund = (total / sum(total)) * 100) %>%
	filter(Genus %in% ancom_taxa | Genus %in% aldex_taxa | Genus %in% corncob_taxa) %>%
	mutate(Genus = str_replace(Genus, "(.*)", "*\\1*"),
	       Genus = str_replace(Genus, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))

at_least_one_iqm <- 
	at_least_one_taxa %>%
	group_by(treat, Genus) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")

sort_order <-
	at_least_one_taxa %>%
	group_by(Genus) %>%
	summarize(median = median(rel_abund)) %>%
	arrange(median)

at_least_one_taxa$Genus <- factor(at_least_one_taxa$Genus, levels = sort_order$Genus)

at_least_one_taxa %>%
	ggplot(aes(x = rel_abund, y = Genus, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
			data = at_least_one_iqm, 
			position = position_dodge(width = 0.75), 
			shape = 21, 
			color = "black") +
	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Relative Abundance (%)", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2'))

```

##### Consensus Taxa

```{r}
aldex_effect <-
  x.all_effect_1 %>% 
  select(Genus, effect)

all_taxa <-
	aes_Genus_G5 %>%
	group_by(sample) %>%
	mutate(rel_abund = (total / sum(total)) * 100) %>%
  left_join(aldex_effect, by = "Genus") %>% 
	filter(Genus %in% ancom_taxa & Genus %in% aldex_taxa & Genus %in% corncob_taxa) %>% 
  ungroup()

da_fams <-
  all_taxa %>% 
  select(Genus) %>% 
  distinct() %$%
  Genus

all_taxa <-
  all_taxa %>%
	mutate(Genus = str_replace(Genus, "(.*)", "*\\1*"),
	       Genus = str_replace(Genus, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))
all_taxa$Genus[all_taxa$Genus == "Unclassified *0319-6G20*"] <- "Uncultured Genus 0319-6G20"
all_taxa$Genus[all_taxa$Genus == "Unclassified *Chloroplast*"] <- "Chloroplast"
all_taxa$Genus[all_taxa$Genus == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
all_taxa$Genus[all_taxa$Genus == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
all_taxa$Genus[all_taxa$Genus == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"

all_iqm <- 
	all_taxa %>%
	group_by(treat, Genus) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")

sort_order <-
	all_taxa %>%
  select(Genus, effect) %>% 
  distinct() %>% 
  arrange(effect)
	# group_by(Genus) %>%
	# summarize(median = median(rel_abund)) %>%
	# arrange(median)

all_taxa$Genus <- factor(all_taxa$Genus, levels = sort_order$Genus)
all_iqm$Genus <- factor(all_iqm$Genus, levels = sort_order$Genus)
all_taxa <-
  all_taxa %>% 
  mutate(direction = if_else(effect > 0, 0, 1))
all_iqm <-
all_taxa %>% 
  select(Genus, direction) %>% 
  distinct() %>% 
  full_join(all_iqm, by = "Genus")
(p <-
	all_taxa %>%
	ggplot(aes(x = rel_abund, y = Genus, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
			data = all_iqm, 
			position = position_dodge(width = 0.75), 
			shape = 21, 
			color = "black") +
	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Relative Abundance (%)", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
    facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
theme(
  strip.background = element_blank(),
  strip.text.y = element_blank()
))

(p_log10 <- 
	p +
	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
		      labels = c(0.001, 0.01, 0.1, 1)))

save(p_log10, da_fams, file = "Output/da_fig.RData")

ggsave('Manuscript/fig4.jpg', p_log10, width = 5, height = 5)

## All
# ggsave("shared_taxa.png", p)
ggsave("shared_taxa_log.png", p_log10, width = 5, height = 5)

# aes_Genus_rclr_plot <-
#   aes_Genus_rclr %>% 
# 	filter(Genus %in% ancom_taxa & Genus %in% aldex_taxa & Genus %in% corncob_taxa) %>% 
#   pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
#   select(Genus, treat, rep, passage, rclr) %>% 
#   distinct() %>% 	
#   mutate(Genus = str_replace(Genus, "(.*)", "*\\1*"),
# 	       Genus = str_replace(Genus, "\\*(.*)_unclassified\\*",
# 				    "Unclassified *\\1*"))
# 
# aes_Genus_rclr_plot$Genus[aes_Genus_rclr_plot$Genus == "Unclassified *0319-6G20*"] <- "Uncultured Genus 0319-6G20"
# aes_Genus_rclr_plot$Genus[aes_Genus_rclr_plot$Genus == "Unclassified *Chloroplast*"] <- "Chloroplast"
# aes_Genus_rclr_plot$Genus <- factor(aes_Genus_rclr_plot$Genus, levels = sort_order$Genus)
# ggplot(aes_Genus_rclr_plot, aes(x = rclr, y = Genus, color = treat)) +
#   facet_grid( ~ passage) +
#   geom_point(position = position_jitterdodge())

p_effect <-
	all_taxa %>%
	ggplot(aes(x = effect, y = Genus)) +
	geom_point() +

	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Aldex2 Effect Size", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2'))
p_effect

```

### Species

```{r setup-differential-abundance-data-species}

aes_G5 <-
  aes_data %>% 
  filter(passage == "G5")

aes_Species_G5 <-
  aes_G5 %>% 
  select(asv, sample, treat, count, Species) %>%
  group_by(sample, Species) %>% 
  mutate(total = sum(count)) %>% 
  ungroup() %>% 
  select(sample, treat, Species, total) %>% 
  distinct()

prevalent_families <- 
  aes_Species_G5 %>% 
  group_by(Species, treat) %>% 
  summarize(prev = sum(total > 0), .groups = "drop") %>% 
  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
  filter(N/12 > 0.1 & P/12 > 0.1) %$%
  Species

otu_table <- 
  aes_Species_G5 %>% 
  filter(Species %in% prevalent_families) %>% 
  select(sample, Species, total) %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

rownames(otu_table) <- otu_table[,1]
otu_table <- otu_table[,-1]

sample_data <- aes_Species_G5 %>% 
  select(sample, treat) %>% 
  distinct() %>% 
  as.data.frame()

rownames(sample_data) <- sample_data[,1]

```

#### Corncob

```{r}

da_analysis <- differentialTest(formula = ~ treat,
                                phi.formula = ~ 1,
                                formula_null = ~ 1,
                                phi.formula_null = ~ 1,
                                test = "Wald", 
                                boot = FALSE,
                                data = otu_table,
                                sample_data = sample_data,
                                fdr_cutoff = 0.05, B = 100,
                                taxa_are_rows = TRUE)

dt_data <- plot(da_analysis, 
                # level = c("Species"), 
                data_only = TRUE)

dt_plot <- function(dt_data) {
  ggplot(dt_data, aes(x = x, xmin = xmin, xmax = xmax, y = reorder(taxa, x))) +
    geom_pointrange() +
    geom_vline(xintercept = 0) +
    labs(x = "Estimate", y = NULL) +
    coord_cartesian(clip="off") +
    # theme(panel.border = element_blank()) +
    theme_classic() +
      theme(axis.text.y = element_markdown())

}

dt_data %>%
  mutate(taxa = str_replace(taxa, "(.*)", "*\\1*"),
         taxa = str_replace(taxa, "\\*(.*)_unclassified\\*",
                                "Unclassified *\\1*")) %>% 
  dt_plot()

```

#### ANCOM

```{r}

# Setup phyloseq object
aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE), sample_data(sample_data))

if(file.exists("Output/ancom_out.rds")) {
  out <- 
    readRDS("Output/ancom_out.rds")
} else {
  # Run Ancom
  out = ancom(phyloseq = aes_ps,  
              p_adj_method = "BH",
              prv_cut = 0.1, 
              lib_cut = 0, 
              main_var = "treat",
              rand_formula = NULL, 
              lme_control = NULL,
              struc_zero = TRUE, 
              neg_lb = FALSE, 
              alpha = 0.05, 
              n_cl = 1)  
  out %>% 
    saveRDS("Output/ancom_out.rds")
}

res = out$res
q_val = out$q_data
beta_val = out$beta_data
# Only consider the effect sizes with the corresponding q-value less than alpha
beta_val = beta_val * (q_val < 0.05) 
# Choose the maximum of beta's as the effect size
beta_pos = apply(abs(beta_val), 2, which.max) 
beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
                  FUN.VALUE = double(1))
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(out$zero_ind), 
                nrow(feature_table), 
                sum(apply(out$zero_ind, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = 0.7 * (n_taxa - 1)

df_fig_w = res %>%
  dplyr::mutate(beta = beta_max,
                direct = case_when(
                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
                  TRUE ~ "Not Significant"
                  )) %>%
  dplyr::arrange(W)
df_fig_w$taxon_id = factor(df_fig_w$taxon_id, levels = df_fig_w$taxon_id)
df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
df_fig_w$direct = factor(df_fig_w$direct, 
                     levels = c("Negative", "Positive", "Not Significant"))

p_w = df_fig_w %>%
  ggplot(aes(x = taxon_id, y = W, color = direct)) +
  geom_point(size = 2, alpha = 0.6) +
  labs(x = "Taxon", y = "W") +
  scale_color_discrete(name = NULL) + 
  geom_hline(yintercept = cut_off, linetype = "dotted", 
             color = "blue", size = 1.5) +
  geom_text(aes(x = 2, y = cut_off + 0.5, label = "W[0.7]"), 
            size = 5, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank())
p_w

```

#### Aldex2

```{r}

conds <- colnames(otu_table) %>% 
  substr(3, 3)

if(file.exists("Output/aldex_x_all.rds")) {
  x.all <- 
    readRDS("Output/aldex_x_all.rds")
} else {
  x.all <- aldex(otu_table, 
                 conds, mc.samples=1000, test="t", effect=TRUE,
                 include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
  x.all %>% 
    saveRDS("Output/aldex_x_all.rds")
}

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
    ylab="Difference")
aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
    ylab="Difference")

par(mfrow=c(1,2))
plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")
legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))

plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")

x.all_effect_1 <-
  x.all %>% 
  filter(abs(effect) > 1) %>% 
  as_tibble(rownames = "Species")

# x.all_effect_1 %>% 
#   # inner_join(taxa_table, by = "asv")
#   ggplot(aes(y = reorder(Species, effect), x = effect)) +
#   geom_point()

# x.all_effect_1 %>% 
#   select(Species, effect)

```

#### Comparison

##### All Significant

```{r}

# CORNCOB
corncob_taxa <- dt_data$taxa

# ANCOM
ancom_taxa <- as.character(df_fig_w[df_fig_w$detected_0.7, ]$taxon_id)

# ALDEX2
aldex_taxa <- x.all_effect_1$Species

at_least_one_taxa <-
	aes_Species_G5 %>%
	group_by(sample) %>%
	mutate(rel_abund = (total / sum(total)) * 100) %>%
	filter(Species %in% ancom_taxa | Species %in% aldex_taxa | Species %in% corncob_taxa) %>%
	mutate(Species = str_replace(Species, "(.*)", "*\\1*"),
	       Species = str_replace(Species, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))

at_least_one_iqm <- 
	at_least_one_taxa %>%
	group_by(treat, Species) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")

sort_order <-
	at_least_one_taxa %>%
	group_by(Species) %>%
	summarize(median = median(rel_abund)) %>%
	arrange(median)

at_least_one_taxa$Species <- factor(at_least_one_taxa$Species, levels = sort_order$Species)

at_least_one_taxa %>%
	ggplot(aes(x = rel_abund, y = Species, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
			data = at_least_one_iqm, 
			position = position_dodge(width = 0.75), 
			shape = 21, 
			color = "black") +
	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Relative Abundance (%)", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2'))

```

##### Consensus Taxa

```{r}
aldex_effect <-
  x.all_effect_1 %>% 
  select(Species, effect)

all_taxa <-
	aes_Species_G5 %>%
	group_by(sample) %>%
	mutate(rel_abund = (total / sum(total)) * 100) %>%
  left_join(aldex_effect, by = "Species") %>% 
	filter(Species %in% ancom_taxa & Species %in% aldex_taxa & Species %in% corncob_taxa) %>% 
  ungroup()

da_fams <-
  all_taxa %>% 
  select(Species) %>% 
  distinct() %$%
  Species

all_taxa <-
  all_taxa %>%
	mutate(Species = str_replace(Species, "(.*)", "*\\1*"),
	       Species = str_replace(Species, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))
all_taxa$Species[all_taxa$Species == "Unclassified *0319-6G20*"] <- "Uncultured Species 0319-6G20"
all_taxa$Species[all_taxa$Species == "Unclassified *Chloroplast*"] <- "Chloroplast"
all_taxa$Species[all_taxa$Species == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
all_taxa$Species[all_taxa$Species == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
all_taxa$Species[all_taxa$Species == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"

all_iqm <- 
	all_taxa %>%
	group_by(treat, Species) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")

sort_order <-
	all_taxa %>%
  select(Species, effect) %>% 
  distinct() %>% 
  arrange(effect)
	# group_by(Species) %>%
	# summarize(median = median(rel_abund)) %>%
	# arrange(median)

all_taxa$Species <- factor(all_taxa$Species, levels = sort_order$Species)
all_iqm$Species <- factor(all_iqm$Species, levels = sort_order$Species)
all_taxa <-
  all_taxa %>% 
  mutate(direction = if_else(effect > 0, 0, 1))
all_iqm <-
all_taxa %>% 
  select(Species, direction) %>% 
  distinct() %>% 
  full_join(all_iqm, by = "Species")
(p <-
	all_taxa %>%
	ggplot(aes(x = rel_abund, y = Species, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
			data = all_iqm, 
			position = position_dodge(width = 0.75), 
			shape = 21, 
			color = "black") +
	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Relative Abundance (%)", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
    facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
theme(
  strip.background = element_blank(),
  strip.text.y = element_blank()
))

(p_log10 <- 
	p +
	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
		      labels = c(0.001, 0.01, 0.1, 1)))

save(p_log10, da_fams, file = "Output/da_fig.RData")

ggsave('Manuscript/fig4.jpg', p_log10, width = 5, height = 5)

## All
# ggsave("shared_taxa.png", p)
ggsave("shared_taxa_log.png", p_log10, width = 5, height = 5)

# aes_Species_rclr_plot <-
#   aes_Species_rclr %>% 
# 	filter(Species %in% ancom_taxa & Species %in% aldex_taxa & Species %in% corncob_taxa) %>% 
#   pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
#   select(Species, treat, rep, passage, rclr) %>% 
#   distinct() %>% 	
#   mutate(Species = str_replace(Species, "(.*)", "*\\1*"),
# 	       Species = str_replace(Species, "\\*(.*)_unclassified\\*",
# 				    "Unclassified *\\1*"))
# 
# aes_Species_rclr_plot$Species[aes_Species_rclr_plot$Species == "Unclassified *0319-6G20*"] <- "Uncultured Species 0319-6G20"
# aes_Species_rclr_plot$Species[aes_Species_rclr_plot$Species == "Unclassified *Chloroplast*"] <- "Chloroplast"
# aes_Species_rclr_plot$Species <- factor(aes_Species_rclr_plot$Species, levels = sort_order$Species)
# ggplot(aes_Species_rclr_plot, aes(x = rclr, y = Species, color = treat)) +
#   facet_grid( ~ passage) +
#   geom_point(position = position_jitterdodge())

p_effect <-
	all_taxa %>%
	ggplot(aes(x = effect, y = Species)) +
	geom_point() +

	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Aldex2 Effect Size", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2'))
p_effect

```

```{r stop}

knitr::opts_chunk$set(eval = FALSE)

```

```{r}

troph_genus_data <-
	aes_data %>%
	filter(Family %in% methanotrophs) %>%
	select(sample, count, Genus, passage, treat) %>%
	group_by(sample, Genus) %>%
	mutate(total = sum(count)) %>%
	select(-count) %>%
	ungroup() %>%
	group_by(sample) %>%
	mutate(rel_abund = total / sum(total)) %>%
	ungroup() %>%
	distinct() 
troph_genus_iqm <- 
	troph_genus_data %>%
	group_by(passage, treat, Genus) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")
troph_genus_data %>%
	ggplot(aes(x = rel_abund, y = Genus, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	facet_grid( ~ passage) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), data = troph_genus_iqm, position = position_dodge(width = 0.75), shape = 21, color = "black") +
	theme_bw()
ggsave("troph_genus.png")

aes_data %>%
	select(sample, count, Genus, passage, treat) %>%
	group_by(sample, Genus) %>%
	mutate(total = sum(count)) %>%
	select(-count) %>%
	ungroup() %>%
	group_by(sample) %>%
	mutate(rel_abund = total / sum(total)) %>%
	ungroup() %>%
	filter(Genus %in% c("Methylocella", "Microvirga", "Methylorosula")) %>%
	distinct() %>%
	ggplot(aes(x = rel_abund, y = Genus)) +
	geom_jitter() +
	facet_grid(treat ~ passage)

troph_family_data <-
	aes_data %>%
	select(sample, count, Family, passage, treat) %>%
	group_by(sample, Family) %>%
	mutate(total = sum(count)) %>%
	select(-count) %>%
	ungroup() %>%
	group_by(sample) %>%
	mutate(rel_abund = total / sum(total)) %>%
	ungroup() %>%
	filter(Family %in% methanotrophs) %>%
	distinct()
troph_family_iqm <- 
	troph_family_data %>%
	group_by(passage, treat, Family) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")
troph_family_data %>%
	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	facet_grid( ~ passage) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), data = troph_family_iqm, position = position_dodge(width = 0.75), shape = 21, color = "black") +
	theme_bw()
ggsave("troph_family.png")

troph_species_data <-
	aes_data %>%
	filter(Family %in% methanotrophs) %>%
	select(sample, count, Species, passage, treat) %>%
	group_by(sample, Species) %>%
	mutate(total = sum(count)) %>%
	select(-count) %>%
	ungroup() %>%
	group_by(sample) %>%
	mutate(rel_abund = total / sum(total)) %>%
	ungroup() %>%
	distinct()
troph_species_iqm <- 
	troph_species_data %>%
	group_by(passage, treat, Species) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")
troph_species_data %>%
	ggplot(aes(x = rel_abund, y = Species, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	facet_grid( ~ passage) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), data = troph_species_iqm, position = position_dodge(width = 0.75), shape = 21, color = "black") +
	theme_bw()
ggsave("troph_species.png")

```

```{r aldex-methanotrophs-family}

x.all <- aldex(troph_data_family, 
               conds, mc.samples=1000, test="t", effect=TRUE,
     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
    ylab="Difference")
aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
    ylab="Difference")

par(mfrow=c(1,2))
plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")
legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))

plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")

x.all

```

```{r aldex-methanotrophs-genus}

x.all <- aldex(troph_data_genus, 
               conds, mc.samples=1000, test="t", effect=TRUE,
     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
    ylab="Difference")
aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
    ylab="Difference")

par(mfrow=c(1,2))
plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")
legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))

plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")

x.all

```

```{r}
## ANCOM
```
```{r}

# Setup phyloseq objects

ps_family <- phyloseq(otu_table(troph_data_family,
                                taxa_are_rows = TRUE),
                      sample_data(sample_data))
ps_genus <- phyloseq(otu_table(troph_data_genus,
                                taxa_are_rows = TRUE),
                      sample_data(sample_data))

# Run Ancom

out_family = ancom(phyloseq = ps_family,  
                  p_adj_method = "BH",
                  prv_cut = 0.1, 
                  lib_cut = 0, 
                  main_var = "treat",
                  rand_formula = NULL, 
                  lme_control = NULL,
                  struc_zero = TRUE, 
                  neg_lb = FALSE, 
                  alpha = 0.05, 
                  n_cl = 1)

out = ancom(phyloseq = ps_genus,  
                  p_adj_method = "BH",
                  prv_cut = 0.1, 
                  lib_cut = 0, 
                  main_var = "treat",
                  rand_formula = NULL, 
                  lme_control = NULL,
                  struc_zero = TRUE, 
                  neg_lb = FALSE, 
                  alpha = 0.05, 
                  n_cl = 1)

```

```{r}

q_val = out$q_data
beta_val = out$beta_data
# Only consider the effect sizes with the corresponding q-value less than alpha
beta_val = beta_val * (q_val < 0.05) 
# Choose the maximum of beta's as the effect size
beta_pos = apply(abs(beta_val), 2, which.max) 
beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
                  FUN.VALUE = double(1))

# Number of taxa except structural zeros
n_taxa = ifelse(is.null(out$zero_ind), 
                nrow(feature_table), 
                sum(apply(out$zero_ind, 1, sum) == 0))

# Cutoff values for declaring differentially abundant taxa
cut_off = 0.7 * (n_taxa - 1)

df_fig_w = out$res %>%
  dplyr::mutate(beta = beta_max,
                direct = case_when(
                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
                  TRUE ~ "Not Significant"
                  )) %>%
  dplyr::arrange(W)
df_fig_w$taxon_id = factor(df_fig_w$taxon_id, levels = df_fig_w$taxon_id)
df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
df_fig_w$direct = factor(df_fig_w$direct, 
                     levels = c("Negative", "Positive", "Not Significant"))

p_w = df_fig_w %>%
  ggplot(aes(x = taxon_id, y = W, color = direct)) +
  geom_point(size = 2, alpha = 0.6) +
  labs(x = "Taxon", y = "W") +
  scale_color_discrete(name = NULL) + 
  geom_hline(yintercept = cut_off, linetype = "dotted", 
             color = "blue", size = 1.5) +
  geom_text(aes(x = 2, y = cut_off + 0.5, label = "W[0.7]"), 
            size = 5, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank())
p_w

```

```{r}

otu_table <-  aes_data %>% 
  filter(asv %in% troph_asvs) %>% 
  filter(passage == "G5") %>% 
  select(sample, count, Genus) %>% 
  group_by(Genus, sample) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  mutate(treat = substr(sample, 3, 3)) %>% 
  group_by(Genus, treat) %>% 
  mutate(prev = sum(total > 0)) %>% 
  ungroup() %>% 
  filter(prev > 2) %>% 
  select(-treat, -prev) %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

rownames(otu_table) <- otu_table[, 1]
otu_table <- otu_table[, -1]

conds <- 
  otu_table %>% 
  colnames() %>% 
  substr(3, 3)

x.all <- aldex(otu_table, 
               conds, mc.samples=1000, test="t", effect=TRUE,
     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
    ylab="Difference")
aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
    ylab="Difference")

par(mfrow=c(1,2))
plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")
legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))

plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")

x.all_effect_1 <-
  x.all %>% 
  filter(abs(effect) > 0.5) %>% 
  as_tibble(rownames = "asv")

x.all_effect_1 %>% 
  ggplot(aes(y = reorder(asv, effect), x = effect)) +
  geom_point()

```

```{r aldex_genus_everything}

otu_table <- 
  aes_data %>% 
  filter(passage == "G5") %>% 
  select(Genus, count, sample) %>% 
  group_by(Genus, sample) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  mutate(treat = substr(sample, 3, 3)) %>% 
  group_by(Genus, treat) %>% 
  mutate(prev = sum(total > 0)) %>% 
  ungroup() %>% 
  filter(prev > 2) %>% 
  select(-treat, -prev) %>% 
  pivot_wider(names_from = "sample", values_from = "total", values_fill = 0) %>% 
  as.data.frame()

rownames(otu_table) <- otu_table[, 1]

otu_table <- otu_table[, -1]

x.all <- aldex(otu_table, 
               conds, mc.samples=1000, test="t", effect=TRUE,
     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
    ylab="Difference")
aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
    ylab="Difference")

par(mfrow=c(1,2))
plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")
legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))

plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")

x.all_effect_1 <-
  x.all %>% 
  filter(abs(effect) > 1) %>% 
  as_tibble(rownames = "Genus")

x.all_effect_1 %>% 
  ggplot(aes(y = reorder(Genus, effect), x = effect)) +
  geom_point()

```

```{r ancova-clr}

gm <- function(x){

  exp(mean(log(x[x>0])))

}

rclr <- function(x) {
  log(x/gm(x))
}

aes_family <-
  aes_data %>%
  select(Family, count, sample, passage, treat) %>%
  group_by(Family, sample) %>%
  mutate(total = sum(count)) %>%
  select(-count) %>%
  ungroup() %>%
  distinct() %>%
  mutate(rep = as.numeric(substr(sample, 4, 5))) %>%
  select(-sample)

prevalent_families <- 
  aes_family %>% 
  mutate(passagetreat = paste0(passage, treat)) %>%
  group_by(Family, passagetreat) %>% 
  summarize(prev = sum(total > 0), .groups = "drop") %>% 
  pivot_wider(names_from = "passagetreat", values_from = "prev", values_fill = 0) %>% 
  filter(G2N/12 > 0.1 & G2P/12 > 0.1 & G5N/12 > 0.1 & G5P/12 > 0.1) %$%
  Family

aes_family_rclr <-
  aes_family %>% 
  filter(Family %in% prevalent_families) %>% 
  group_by(passage, treat, rep) %>%
  mutate(rclr = log(total/gm(total))) %>%
  select(-total) %>%
  ungroup() %>%
  pivot_wider(names_from = passage, values_from = rclr)

ancova <- function(df) {
  lm(G5 ~ G2 + treat, data = df, na.action = )
}

model_results <-
aes_family_rclr %>%
  nest(data = -Family) %>%
  mutate(model = map(data, ancova)) %>%
  mutate(tidy = map(model, tidy)) %>% 
  unnest(tidy) %>% 
  filter(term == "treatP")

model_results$p.adjust <- p.adjust(model_results$p.value, method = "BH")

model_results %>%
  filter(p.adjust < 0.05) %>% 
  arrange(estimate) %>% 
  print(n = 22)

model_results %>% 
  filter(p.adjust < 0.05) %>% 
  arrange(estimate) %>% 
  unnest(data) %>% 
  ggplot(aes(x = estimate, xmax = estimate + std.error, xmin = estimate - std.error, y = reorder(Family, estimate))) +
  geom_pointrange() +
  labs(x = "Estimate \n(log-ratio increase in P)", y = NULL) +
  theme_bw()
ggsave("ancova-plot.png", width = 5, height = 5)

model_results %>% 
  filter(p.adjust < 0.05) %>% 
  unnest(data) %>% 
  pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
  select(Family, treat, rep, passage, rclr) %>% 
  distinct() %>% 
  ggplot(aes(x = rclr, y = Family, color = treat)) +
  facet_grid( ~ passage) +
  geom_point(position = position_jitterdodge())

```

```{r}

aes_family_data <-
  aes_data %>% 
  select(asv, sample, passage, treat, count, Family) %>%
  group_by(sample, Family) %>% 
  mutate(total = sum(count)) %>% 
  ungroup() %>% 
  select(sample, treat, passage, Family, total) %>% 
  distinct()

prevalent_families <- 
  aes_family_data %>% 
  group_by(Family, treat, passage) %>% 
  summarize(prev = sum(total > 0), .groups = "drop") %>% 
  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
  filter(N/12 > 0.1 & P/12 > 0.1) %$%
  Family

aes_family_data <-
  aes_family_data %>% 
  filter(Family %in% prevalent_families) %>% 
  mutate(jar = substr(sample, 3, 5)) %>%
  select(-sample) %>%
  pivot_wider(names_from = "passage", values_from = "total", values_fill = 0)

ancova_model <- function(df) {
	lm(G5 ~ G2 + treat, data = df)
}

aes_family_nested <-
	aes_family_data %>%
	group_by(Family) %>%
	nest() %>%
	mutate(model = map(data, ancova_model)) %>%
	mutate(tidy = map(model, tidy)) %>%
	ungroup()

aes_ancova <- aes_family_nested %>%
	unnest(tidy) %>%
	filter(term == "treatP")

aes_ancova$p.adjust <- p.adjust(aes_ancova$p.value, method = "BH")

aes_ancova %>%
	arrange(estimate) %>%
	filter(p.adjust < 0.05) %>%
	print(n = 22)

```

```{r}

aes_family_data <-
  aes_data %>% 
  select(asv, sample, passage, treat, count, Family) %>%
  group_by(sample, Family) %>% 
  mutate(total = sum(count)) %>% 
  ungroup() %>% 
  select(sample, treat, passage, Family, total) %>% 
  distinct()

prevalent_families <- 
  aes_family_data %>% 
  group_by(Family, treat, passage) %>% 
  summarize(prev = sum(total > 0), .groups = "drop") %>% 
  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
  filter(N/12 > 0.1 & P/12 > 0.1) %$%
  Family

otu_table <- 
  aes_family_data %>% 
  filter(Family %in% prevalent_families) %>% 
  select(sample, Family, total) %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

rownames(otu_table) <- otu_table[,1]
otu_table <- otu_table[,-1]

sample_data <- aes_family_data %>% 
  select(sample, treat, passage) %>% 
  distinct() %>% 
  as.data.frame()

rownames(sample_data) <- sample_data[,1]
sample_data <- sample_data[, -1]
sample_data <-
	sample_data %>%
	mutate(passage12 = if_else(passage == "G2", 1, 2))

# Setup phyloseq object
aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE), sample_data(sample_data))

set.seed(123)
out2 = ancom(phyloseq = aes_ps,  p_adj_method = "holm",
             prv_cut = 0.10, lib_cut = 1000, main_var = "treat", 
             adj_formula = "passage12", 
             rand_formula = "~ passage12 | 1",
             lme_control = list(maxIter = 100, msMaxIter = 100, opt = "optim"), 
             struc_zero = TRUE, neg_lb = FALSE, alpha = 0.05, n_cl = 2)

res2 = out2$res

out2 = ancom(phyloseq = family_data2,  p_adj_method = "holm",
             prv_cut = 0.10, lib_cut = 1000, main_var = "group", 
             adj_formula = "nationality + timepoint + bmi_group", 
             rand_formula = "~ timepoint | subject",
             lme_control = list(maxIter = 100, msMaxIter = 100, opt = "optim"), 
             struc_zero = TRUE, neg_lb = TRUE, alpha = 0.05, n_cl = 2)

res2 = out2$res

```

```{r}

x.all_effect_1$Family

dt_data$taxa

```

```{r}
combined_da <-
  dt_data %>% 
  as_tibble %>% 
  full_join(sig_families, by = c("taxa" = "Family")) %>% 
  full_join(x.all_effect_1, by = c("taxa" = "Family")) %>% 
  full_join(df_fig_w[df_fig_w$detected_0.7,], by = c("taxa" = "taxon_id"))

shared_da <-
  dt_data %>% 
  as_tibble %>% 
  inner_join(sig_families, by = c("taxa" = "Family")) %>% 
  inner_join(x.all_effect_1, by = c("taxa" = "Family")) %>% 
  inner_join(df_fig_w[df_fig_w$detected_0.7,], by = c("taxa" = "taxon_id"))

shared_da %>% 
  ggplot(aes(y = reorder(taxa, effect), x = effect)) +
  geom_point()

aes_data %>% 
  filter(passage == "G5") %>% 
  select(sample, Family, count) %>% 
  group_by(sample, Family) %>% 
  summarize(total = sum(count)) %>% 
  filter(Family %in% shared_da$taxa) %>% 
  mutate(treat = substr(sample, 3, 3)) %>% 
  ggplot(aes(x = total, y = Family, color = treat)) +
  geom_jitter() +
  scale_x_log10()

rubini_seqs <-
  aes_data %>% 
  filter(Family == "Rubinisphaeraceae") %>% 
  select(asv, seq, Kingdom:Species) %>% 
  distinct()

write(rubini_seqs$seq, file = 'Output/rubini_seqs.txt')

```

```{r}

##############

First, I looked at which taxa are unique to treatments N and P. Several families
present in P and absent in N. Most of them had low prevalence (present in 3 or
fewer samples) and low abundance (median < 5 reads across samples). There were two families unique
to the Positive treatment with relatively high prevalence. This included an ASV
that was a member of the Bacteroidia Class with no lower taxonomic designation. 
This ASV had a prevalence of 10/12 and a median abundance of ~2 reads. The other
prevalent family was a member of the Silvanigrellaceae, a newly described family placed in
its own order. This family was present in
all 12 samples and was represented by three OTUs: one with no lower designation, one from the genus
Silvanigrella and one from the genus Candidatus Turbobacter. Silvanigrella was
isolated from a temperate fresh water lake (https://www.microbiologyresearch.org/content/journal/ijsem/10.1099/ijsem.0.001965).

Of the families unique to the Neutral treatment, only one had a prevalence greater than 2/12. 
This family, Armatimonadaceae, was present in half of the Neutral samples (6/12). This
family was represented by two ASVs in the genus Armatimonas. The type strain for
Armatimonas was isolated from the rhizoplane of Phragmites australis (https://www.microbiologyresearch.org/content/journal/ijsem/10.1099/ijs.0.025643-0)
```

```{r}
# families in to Passage 5, Treatment P
g5p<- 
  ps_fixed %>% 
  ps_filter(treat == "p", passage == 5) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()

# families in to Passage 5, Treatment N
g5n<- 
  ps_fixed %>% 
  ps_filter(treat == "n", passage == 5) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()

# Families unique to p in passage 5
g5p_unique <-
  g5p %>% 
  filter(Family %in% unique(g5p$Family)[!(unique(g5p$Family) %in% unique(g5n$Family))]) %>% 
  mutate(Prevalence = as.numeric(Abundance > 0))

# Abundance
Family_order<-
  g5p_unique %>% 
  group_by(OTU) %>% 
  summarize(median = median(Abundance), max = max(Abundance)) %>% 
  arrange(median, max)
g5p_unique$OTU <- factor(g5p_unique$OTU, levels = Family_order$OTU)

# Plot families unique to P5 in order of median abundance
g5p_unique %>% 
  ggplot(aes(y = Abundance, x = OTU)) +
  geom_tufteboxplot() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 24)) +
  labs(title = "Unique to P5 by abundance")

# Prevalence
Family_order<-
  g5p_unique %>% 
  group_by(OTU) %>% 
  summarize(sum = sum(Prevalence)) %>% 
  arrange(sum)
g5p_unique$OTU <- factor(g5p_unique$OTU, levels = Family_order$OTU)

# Plot families unique to P5 in order of prevalence
g5p_unique %>% 
  group_by(OTU) %>% 
  summarize(Prevalence = sum(Prevalence)) %>% 
  ggplot(aes(y = Prevalence, x = OTU)) +
  geom_point() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(1, 2, 3, 5, 10, 12)) +
  labs(title = "Unique to P5 by prevalence")

# Families unique to n in passage 5
g5n_unique <-
  g5n %>% 
  filter(Family %in% unique(g5n$Family)[!(unique(g5n$Family) %in% unique(g5p$Family))]) %>% 
  mutate(Prevalence = as.numeric(Abundance > 0))

# Abundance
Family_order<-
  g5n_unique %>% 
  group_by(OTU) %>% 
  summarize(median = median(Abundance), max = max(Abundance)) %>% 
  arrange(median, max)
g5n_unique$OTU <- factor(g5n_unique$OTU, levels = Family_order$OTU)

# Plot families unique to N5 in order of abundance
g5n_unique %>% 
  ggplot(aes(y = Abundance, x = OTU)) +
  geom_tufteboxplot() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(0, 10, 20, 30, 37)) +
  labs(title = "Unique to N5 by abundance")

# Prevalence
Family_order<-
  g5n_unique %>% 
  group_by(OTU) %>% 
  summarize(sum = sum(Prevalence)) %>% 
  arrange(sum)
g5n_unique$OTU <- factor(g5n_unique$OTU, levels = Family_order$OTU)

# Plot families unique to N5 in order of prevalence

g5n_unique %>% 
  group_by(OTU) %>% 
  summarize(Prevalence = sum(Prevalence)) %>% 
  ggplot(aes(y = Prevalence, x = OTU)) +
  geom_point() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(1, 2, 6)) +
  labs(title = "Unique to N5 by prevalence")

```

```{r}
# Differential Abundance Treatment in passage 5
```

```{r corncob_treatment, results=FALSE, fig.hight=40}

ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")

ps_fam_pas5 <- microViz::ps_filter(ps_fixed_family, passage == 5)

ps_fam_pas5_prev <- microViz::tax_filter(ps_fam_pas5, min_prevalence = 0.10)

sample_data(ps_fam_pas5_prev) =
  sample_data(ps_fam_pas5_prev)[,c('sample_id', 'treat')] %>% 
  sample_data()
ps = ps_fam_pas5_prev

# Differential abundance test
da_analysis <- differentialTest(formula = ~ treat,
                                 phi.formula = ~ 1,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ 1,
                                 test = "Wald", boot = FALSE,
                                 data = ps_fam_pas5_prev,
                                 fdr_cutoff = 0.05, B = 50)

#plot(da_analysis, level = c("Family"))
da_analysis$significant_models
dt_data <- plot(da_analysis, level = c("Family"), data_only = TRUE)

sig_taxa_treat <- tax_select(ps_fam_pas5_prev, da_analysis$significant_taxa)
ntaxa(sig_taxa_treat)
save(da_analysis, sig_taxa_treat, dt_data, file = 'Output/da_corncob.RData')

```

```{r corncob-plot, fig.width=6, fig.height=6}
dt_plot(dt_data)

```

```{r corncob_ch4}

ps_fam_pas5_trep <- ps_filter(ps_fam_pas5, treat == "p")

ps_fam_prev_trep_prev <- tax_filter(ps_fam_pas5_trep, min_prevalence = 0.10)

da_analysis <- differentialTest(formula = ~ ch4,
                                 phi.formula = ~ ch4,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ ch4,
                                 test = "Wald", boot = FALSE,
                                 data = ps_fam_prev_trep_prev,
                                 fdr_cutoff = 0.05)
otu_to_taxonomy(OTU = da_analysis$significant_taxa, data = ps_fam_prev_trep_prev)

plot(da_analysis, level = c("Family"))

# Check unfitted taxa
# otu_to_taxonomy(which(is.na(da_analysis$p)) %>% names, data = ps_gen_prev)
# otu_table(ps_gen_prev)[,which(is.na(da_analysis$p))]

# Differential variability test

dv_analysis <- differentialTest(formula = ~ passage + treat + ch4,
                                 phi.formula = ~ passage + treat + ch4,
                                 formula_null = ~ passage + treat + ch4,
                                 phi.formula_null = ~ passage + treat,
                                 data = ps_gen_prev,
                                 test = "Wald", boot = FALSE,
                                 fdr_cutoff = 0.05)
# otu_to_taxonomy(OTU = dv_analysis$significant_taxa, data = ps_gen_prev)
plot(dv_analysis, level = c("Family", "Genus"))

# Check unfitted taxa

# otu_to_taxonomy(which(is.na(dv_analysis$p)) %>% names, data = ps_gen_prev)
# otu_table(ps_gen_prev)[,which(is.na(dv_analysis$p))]

```