### 2022-12-06 manuscript/freewrite.Rmd

---
title: Artificial ecosystem selection reveals relationships between microbiome composition and ecosystem function
author: "Andrew H. Morris^1^* and Brendan J. M. Bohannan^1^"
output:
  word_document:
    reference_docx: reference.docx
bibliography: bibliography.bib
csl: vancouver.csl
---

^1^ Institute of Ecology and Evolution, University of Oregon, Eugene, OR, USA  
\* Correspondence: Andrew H. Morris, amorris3\@uoregon.edu  
Competing Interests: We declare we have no competing interests.

```{r options, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
#knitr::opts_knit$set(root.dir=normalizePath('../'))

```

# Questions

- What should be the order of the Intro?

    1. What are we interested in and why
    2. Past approaches and their limitations
    3. Our approach and why it's better
    4. Broader conceptual framework (Figure 1, Equation 1)

- Or:

    1. What are we interested in and why
    2. Past approaches and their limitations
    3. Broader conceptual framework (Figure 1, Equation 1)
    4. Our approach and why it's better

- Or:

    1. What are we interested in and why
    2. Broader conceptual framework (Figure 1, Equation 1)
    3. Past approaches and their limitations
    4. Our approach and why it's better

# Introduction

## What is the knowledge gap?

### What are we interested in, why are we interested in it, how has it been and studied and what are the shortfalls, and why precisely is our approach better?

Microbiomes mediate a variety of important ecosystem functions. This suggests
that understanding variation in microbiome composition might be important for
predicting the rates of ecosystem functions. Many studies have attempted to
document such a relationship. However, it is difficult to isolate the effect of
microbiome composition from other drivers of variation in ecosystem function.
For example, microbiome manipulations often confound "diversity" with broad
phylogenetic patterns or "composition" with abundance. Likewise, comparative
studies must control for abiotic variation statistically to identify the direct
effect of the microbiome, which relies on measuring the correct abiotic
variables and cannot account for unmeasured environmental variation. Here, we
overcome these limitations by using a selection approach to estimate the degree
to which an ecosystem function varies with microbiome composition. We want to
understand what is the relationship between _variation in the microbiome_ and
_variation in the rate of ecosystem function_?

There have been two general categories of approaches that investigators have
applied to address this question: comparative and manipulative. Comparative
studies sample natural variation in an ecosystem function across different
habitats and simultaneously measure variation in community composition.
Investigators can then correlate ecosystem function with community structure
while trying to control for environmental variation. These approaches have
documented important relationships between microbiomes and ecosystem
functions. For example, a meta-analysis of these studies observed a small but
significant unique contribution of the microbiome to variation in ecosystem
function after controlling for environmental variation (Graham 2016). In
addition, studies focusing on the correlation between the rate of an ecosystem
function and the abundance of an associated functional marker gene observe a
significant correlation, though this relationship is rare and contingent upon
both the function and the ecosystem sampled (Rocca 2015). Comparative studies
come with some unique challenges and limitations. One issues is the covariance
problem where microbiome attributes tend to covary with the abiotic conditions
within an environment and whether and how the investigator controls for these
variables affects the conclusions. In addition, it is difficult to know a
priori which environmental variables or community attributes to measure.
Finally, while these approaches can establish a potential magnitude and
direction for these relationships, they fail to establish a mechanism
underlying the relationship.

The other broad category of approaches to address this question are
manipulative approaches. Manipulative experiments try to alter microbial
community structure and observe the effect on function. For example,
reciprocal transplant and common garden experiments have shown that
microbiomes originating from different ecosystems inoculated onto the same
substrate or introduced into a common environment display distinct functional
rates (Cavigeli 2000, Balser 2005, Waldrop 2006, Strickland 2009). In
addition, manipulating diversity by filtering communities by cell size or
through dilution alters ecosystem function (Wagg 2014, 2020, dilution
citation?). However, manipulating the microbiome directly is challenging. Each
of the manipulative approaches applied confound community structure with other
factors such as abiotic conditions or phylogenetic relationships. For example,
reciprocal transplant and common garden experiments confound community
structure with the abiotic conditions carried on the inoculum (Reed and
Martiny 2007). In addition, manipulating microbial diversity through filtering
confounds diversity with broad phylogenetic categories such as nematodes,
fungi, and bacteria while manipulating diversity through dilution confounds
diversity with biomass and composition.

In this study, we sought to build on the observations of comparative and
manipulative studies by applying a different approach to the question of
whether variation in the microbiome contributes to variation in the rate of an
ecosystem function. We used artificial ecosystem selection to select for
microbiomes that performed a greater rate of ecosysem function. We then tested
whether variation in the microbiome contributed to variation in the rate of
ecosystem function and identified microbiome attributes that might explain
this relationship. There are several potential advantages to this approach for
documenting BEF relationships and for investigating potential mechanisms
underlying those relationships. In particular, by passaging microbiomes over
multiple rounds using a common substrate (or isogenic host), we can weaken the
covariance between microbes and the environment by repeatedly diluting the
abiotic effect. In addition, by inoculating our ecosystems with equally sized
subsamples, we eliminate the need to generate microbiome variation through
methods that are confounded with diversity, biomass, or cell size. Lastly, by
comparing our artificially selected community to a control community without
selection or with random selection, we can both control for changes in the
environment over time and identify significant genes or taxa associated with our
function of interest to establish a potential mechanism that explains this
variation.

One important conceptual distinction that we should make when investigating
biodiversity-ecosystem function relationships is to be clear that we are
interested in _variation_. In particular, we are interested in what drives
variation in the rate of an ecosystem function across ecosystems. For example,
we want to know whether variation in the rate of methane oxidation rate in soils
is driven by variation in microbial community structure. This question is
distinct from the question "Which microbes consume methane?" To illustrate this
distinction, we can draw an analogy with the genetic basis of traits in
organisms. If we are interested in whether variation in human height is
associated with variation in genotype, we can sample a population of humans,
measure their height, and look at the correlation between phenotype and genotype
using gene markers or proxies for genotypic similarity, such as kinship.
However, this does not address the question of "Which genes control human
height?" Likewise, we could go into the lab and perform an experiment on mice
where we knockout or over-express a gene thought to be associated with height
and observe height differences among the treatment groups. This would
demonstrate that a particular gene affects height, but it does _not_ demonstrate
that variation in genotype drives variation in height in a population.
___WHY DOES THIS MATTER?___

To understand variation in the rate of an ecosystem function, we can think about
different drivers of variation as illustrated in Figure 1 and Equation 1. In
this model, variation in the rate of an ecosystem function ($V_F$) is the sum of
the variation in the environmental conditions ($V_E$), variation in the
microbiome ($V_M$), and $\epsilon$, which represents variation unexplained by
the other two factors, technical error, and unmeasured environmental or
microbial variation. In addition, we can consider another variable,
$cov(V_E:V_M)$, which is the covariance between the microbiome and the
environment. We can conceptualize this as a situation in which a particular
environmental condition, such as pH or moisture content, selects for a particular
microorganism that is associated with variation in function. For simplicity, we
have omitted a final parameter, which is the interaction between the microbiome
and the environment, which would represent a change in the effect of the
microbiome under different environmental conditions. Here, we assume we're
talking about a single ecosystem and so assume this parameter to be zero.

![Figure 1. Conceptual diagram of the potential relationship between environmental variation, microbiome variation, and variation in the rate of an ecosystem function. We omitted the reverse arrows for illustration, but in theory those relationships also exist but we make the assumption that they are small. In addition, this diagram represents the relationships for a single environment, but one could imagine the same microbiome exerting a different influence on variation in function in a second environment, which would be illustrated by the interaction between microbiome and environment ($V_E:V_M$)`](Figure1.png)

$$
V_F = V_E + V_M + COV(V_E:V_M) + \epsilon
$$

How can we use this conceptual framework to address our original question: Does
variation in the microbiome contribute to variation in the rate of ecosystem
function? Under the null hypothesis that we can ignore the microbiome and
simply use abiotic environmental variables to predict function, we would assume
that $V_M$ is essentially zero and that all of the variation in function
($V_F$) is due to variation in the environment and the covariance between the
microbiome and the environment ($V_E + COV(V_E:V_M)$). In this case, it is
unnecessary to understand anything about the microbiome. However, if we consider
the results from Graham (2016), this hypothesis does not appear to match the
data. Graham observed that across the studies that they analyzed, environmental
variables explained an average of 56\% of the variation in ecosystem function
while microbial community structure explained an average of 31\% of the
variation. Models that included both explained an average of 65\% of the
variation. Putting these numbers together, we see that $V_M$ = 11\% (65 - 56),
$cov(V_E:V_M)$ = 20\% (31 - $V_M$ = 31 - 11), and $V_E$ = 36\% (56 -
$cov(V_E:V_M)$ = 56 - 20). Therefore, the environment explains 36% of the
variation directly
and 20% via covariation with the microbiome and the microbiome explains 11%
directly. This means at least 1/6 (11/65) of the explained variation can be
uniquely attributed to the microbiome according to the conclusions of Graham.

Of course, these numbers should be taken with a grain of salt. As discussed
earlier, comparative studies such as these are unable to break the covariance
between microbes and the environment. In this case, $V_M$ and $\epsilon$ may
represent the direct or indirect effect (via covariance) of unmeasured
environmental variables rather than a unique contribution of the microbiome. In
addition, the proportion of variation depends upon the magnitude of the
variation in $V_F$ and $V_E$. For example, if we were to include variation in
methane emissions of wetland and upland ecosystems, we might find that virtually
100\% of the variation in function was due to variation in the environment,
since methane production requires anaerobic conditions. However, if we looked
across wetlands, we might observe a smaller $V_F$ and $V_E$ and an equivalent
$V_M$. Under this scenario, even though $V_M$ was the same, $\frac{V_M}{V_F}$
would be much greater and therefore the environment would no longer explain
100\% of the variation. Therefore, any study that wishes to address this
question must consider the variation in the population that they are studying.

We performed artificial ecosystem selection on soil microbiomes by selecting on
the methane oxidation rate. To perform this experiment, we created 24 jars in an
identical manner using sterile potting mix inoculated with living soil. Each jar
was randomly assigned to either a positive selection treatment that underwent
selection on high soil methane oxidation rate or a random selection treatment.
At the end of each incubation, we selected jars with a high methane oxidation
rate (for the positive treatment), homogenized them, and used this homogenate to
inoculate a new set of sterile soil jars. We repeated this process over multiple
passges. The idea is that by shuffling microbiome composition through
subsampling and homogenizing the environment we are exploring the parameter
space for the bidoversity-ecosystem function relationship. In addition,
subsampling at each step would dilute away the effect of the abiotic
environment. The only way for there to be a response to selection would be for
microbes to actually grow in the medium.

Here are some hypothetical results we could observe. If soil methane oxidation
rate is entirely environmentally determined, i.e., in the parlance of our
conceptual framework that would mean that $V_E$ and $cov(V_E:V_M)$ are large and
that $V_M$ is relatively small, then we would expect to observe no response to
selection. This is because the effect of the abiotic environment would be
diluted at eaching passaging step. If variation in methane oxidation rate was
independent of microbes and the environment in our system then the varation
should be random and $\epsilon$ would be large relative to the other parameters
and again we would observe no response to selection. If variation in methane
oxidation rate is at least partially driven by variation in the microbiome then
we would expect manipulating microbiome composition through selection would
result in an increase in methane oxidation rate as a response to selection. This
would manifest as a correlation between donor and recipient jars in consecutive
passages.

(Aside: What we seem to be saying in the above paragraph is that passaging
ecosystem propagules is somehow intrinsically tied to the magnitude of Vm. This
seems to be based on the argument that Vm reproduces itself with high fidelity,
but Ve does not. Is this true? Adding a low pH 5g of soil to a neutral pH 45g of
soil doesn't shift the pH, but adding a 5g soil microbial inoculum to 45g of
autoclaved soil does shift the community? Is this any different than the
manipulative experiments we describe above? Or perhaps the argument is that
_repeated_ passaging dilutes away the environmental effect so by the time you
get to passage 5 it's 1/1e5 of the original soil inoculum, but if the microbes
have reproduced then they have rebuilt their population at each passage. I guess
this is the idea behind serial batch culturing??? IDK if a reviewer will be
convinced by this argument.)