### 2023-01-04 manuscript/manuscript.Rmd

---
title: Artificial ecosystem selection reveals relationships between microbiome composition and ecosystem function
author: "Andrew H. Morris^1^* and Brendan J. M. Bohannan^1^"
output:
  word_document:
    reference_docx: reference.docx
bibliography: bibliography.bib
csl: vancouver.csl
---

^1^ Institute of Ecology and Evolution, University of Oregon, Eugene, OR, USA  
\* Correspondence: Andrew H. Morris, amorris3\@uoregon.edu  
Competing Interests: We declare we have no competing interests.

```{r options, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir=normalizePath('../'))

```

```{r initialize-packages}
library(tidyverse)
library(broom)
library(broman)
library(morris)
# library(phyloseq)
library(ggpubr)
library(knitr)
library(corncob)
# library(ggthemes)
library(captioner)
library(knitr)
library(ggtext)
figures <- captioner()
tables <- captioner(prefix = "Table")

source('R/functions.R')
```

```{r load-data}
heritability <- read_tsv(paste0(der_dir, 'heritability.tsv'))
fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))

load(file = paste0(der_dir, 'decontam.RData'))

# CH4 models
load(file = paste0(der_dir, 'response_model.RData'))
load(file = paste0(der_dir, 'response_plot.RData'))
load(file = paste0(der_dir, 'herit_model.RData'))
load(file = paste0(der_dir, 'herit_plot.RData'))

# 16S models
load(file = paste0(der_dir, 'richness_models.RData'))
load(file = paste0(der_dir, 'phyloseq.RData'))
load(file = paste0(der_dir, 'beta_model.RData'))
load(file = paste0(der_dir, 'beta_fig.RData'))
load(file = paste0(der_dir, 'da_corncob.RData'))
load(file = paste0(der_dir, 'da_fig.RData'))
load(file = paste0(der_dir, 'dbrda.RData'))

```

# Abstract

Microbiomes mediate important ecosystem functions, yet it has proven difficult
to determine the relationship between microbiome composition and the rate of
ecosystem functions. This challenge remains because it is difficult to
manipulate microbiome composition directly, we often cannot know a priori which
members of a microbiome influence the rate of an ecosystem function, and
microbiomes can covary strongly with other drivers of ecosystem function, such
as the environment. We propose that artificial ecosystem selection can overcome
these challenges. Artificial ecosystem selection involves creating replicate
ecosystems, selecting replicates with a desired ecosystem-level trait, and
using the microbiomes from the selected ecosystems to inoculate a new set of
ecosystems. We performed artificial ecosystem selection on replicate soil
microbiomes to select for increased soil methane oxidation rate. We observed a
strong response to selection with a `r myround(per_slope, 1)`\% increase in
methane oxidation rate per passage. In addition, we
estimated that `r myround((coefficients(herit_model)[2] +
coefficients(herit_model)[4])*100, 1)`\% of the variation in methane oxidation
rate in these soils can be attributed to microbiome variation. We also found
that selection did not enrich for known methane oxidizers; rather microbes that
responded to selection included members of `r length(da_fams)` families not
known to oxidize methane, such as *Fimbriimonadaceae*, *Cytophagaceae*, and
*Diplorickettsiaceae*. This result is in contrast to the typical assumption that
the rate of an ecosystem function is limited by the final step in the associated
microbial pathway. Our study demonstrates that variation in microbiome
composition can contribute to variation in the rate of ecosystem function and
that manipulating microbiome composition could be a viable strategy for managing
ecosystem functions.

# Introduction

Microbiomes mediate a variety of important ecosystem functions. This suggests
that understanding variation in microbiome composition might be important for
predicting the rates of ecosystem functions 
[@conrad1996; @schimel1998; @crowther2019]. 
Many studies have attempted to document such a relationship
[@cavigelli2000; @balser2005; @waldrop2006; @strickland2009; @wagg2014; @wagg2019].
However, it is difficult to isolate the effect of variation in microbiome
composition from other drivers of variation in ecosystem function. For example,
microbiome manipulations often confound variation in diversity, composition, and
abundance. Likewise, comparative studies must control for abiotic variation
statistically to identify the direct effect of microbiome variation. This
approach relies on measuring the relevant abiotic variables and cannot account
for unmeasured environmental variation. Here, we overcome these limitations by
using a selection approach to estimate the degree to which an ecosystem function
varies with microbiome composition.

There have been two general categories of approaches that investigators have
used to estimate the degree to which an ecosystem function varies with
microbiome composition: comparative and manipulative. Comparative
studies sample natural variation in an ecosystem function across different
habitats and simultaneously measure variation in community composition.
Investigators can then correlate ecosystem function with aspects of community
composition while attempting to control for environmental variation. These
approaches have
documented important relationships between microbiomes and ecosystem
functions. For example, a meta-analysis of these studies observed a small but
significant contribution of the microbiome to variation in ecosystem
function after controlling for environmental variation [@graham2016]. In
addition, studies focusing on the correlation between the rate of an ecosystem
function and the abundance of an associated functional marker gene observe a
significant correlation, though this relationship is rare and contingent upon
both the function and the ecosystem sampled [@rocca2015]. However, comparative
studies come with some unique challenges and limitations. One issue is that
microbiome attributes tend to covary with the abiotic conditions within an
environment, and whether and how the investigator controls for these variables
can affect the conclusions of such a comparative study. In addition, it is
difficult to know a priori which environmental variables or community attributes
to measure. Finally, while these approaches can establish a potential magnitude
and direction for these relationships, it is often difficult to identify the
taxa or genes that explain the connection between composition and function.

The other broad category of approaches used to address this question are
manipulative approaches. Manipulative experiments try to alter microbial
community composition and observe the effect on function. For example,
reciprocal transplant and common garden experiments have shown that microbiomes
originating from different ecosystems inoculated into the same substrate or
introduced into a common environment display distinct functional rates 
[@cavigelli2000; @balser2005; @waldrop2006; @strickland2009]. In addition,
manipulating diversity by filtering communities by cell size or through dilution
has been shown to alter the rate of ecosystem functions 
[@wagg2014; @wagg2019; @griffiths2001]. However, manipulating the microbiome
directly is challenging. Each of the manipulative approaches applied confound
community composition with other factors. For example, reciprocal transplant and
common garden experiments confound community composition with the abiotic
conditions carried on the inoculum while manipulating composition through
dilution may confound composition with biomass [@reed2007].

(Side note: papers like the Strickland paper were actually asking a subtly
different question from the one that we are asking. Strickland was asking
whether communites can be functionally dissimilar as opposed to functionally
redundant. That is, that historical contingency plays a role in the functioning
of microbial communities. They demonstrated quite well that communities
originating from different environments could have unique functional rates.
However, is that due to microbiome-environment covariance where a different
environment (i.e., a different litter substrate) selects for a litter-degrading
community or is that a result of the unique effect of variation in the
community? They actually point this out in their Discussion, they suggest future
research should determine whether the differences between communities are driven
by taxonomic composition or is function the result of the selection environment
(I think of this is analogous to our "covariation" argument)? They also suggest
trying to adapt communities to a new environment through continuous inputs of
the same resource (just like in our experiment!))

In this study, we sought to build on the observations of comparative and
manipulative studies by applying a different approach to the question of whether
variation in the microbiome contributes to variation in the rate of an ecosystem
function. We used artificial ecosystem selection to select for microbiomes that
performed a greater rate of ecosystem function
[@swenson2000; @blouin2015; @panke-buisse2015]. We then tested whether variation
in the microbiome contributed to variation in the rate of ecosystem function and
identified microbiome attributes that might explain this relationship. There are
several potential advantages to this approach for documenting
microbiome-function relationships and for investigating potential mechanisms
underlying those relationships. By passaging microbiomes over multiple rounds
using a common environment, we can weaken the covariance between microbes and
the environment by repeatedly diluting abiotic effects. In addition, by
inoculating our ecosystems with equally sized subsamples, we eliminate the need
to generate microbiome variation through methods that are confounded with
biomass or cell size. Lastly, by comparing our artificially selected community
to a control community resulting from random selection, we can both control for
changes in the environment over time and identify significant genes or taxa
associated with our selection treatment to establish a potential mechanism that
explains this variation.

We applied artificial ecosystem selection to soil microbiomes by selecting on
soil methane oxidation rate. We chose this function because methane is a
globally important greenhouse gas and methane oxidation by soil bacteria is the
primary biological sink for atmospheric methane as well as possibly the main driver
of temporal methane dynamics [@kirschke2013]. In addition, there is evidence that
soil methane oxidation rate may vary with microbiome composition based on
comparative studies in a variety of arctic and tropical ecosystems 
[@meyer2020; @meyer2020a; @freitag2009; @freitag2010] as well as studies that
manipulate methanotroph richness [@schnyder2018]. Finally, methanotrophy is one
of the most deeply conserved microbial physiologies and is represented in a
narrow range of taxa, which suggests that the taxonomic composition of the
microbiome is more likely to be associated with the rate of methane oxidation
than other broader or more shallowly conserved functions
[@schimel1998; @martiny2013].

In this study, we evaluated whether variation in the microbiome contributes to
variation in the rate of ecosystem functions. To address this problem, we
applied artificial ecosystem selection to methane oxidation rate of the soil
microbiome and answered the following questions: Does variation in the relative
abundance of microbial taxa contribute to variation in soil methane oxidation
rate in the lab? How much of the variation in soil methane oxidation rate can
be attributed to the microbiome? What attributes of the microbiome regulate
variation in methane oxidation rate? And do these attributes match our
assumptions about what regulates methane oxidation rate in nature?

# Materials and Methods

## Experimental design

We performed an artificial ecosystem selection experiment 
[*sensu* @swenson2000] by passaging replicate soil microbiomes. The trait we
selected on was CH~4~ oxidation rate. This experiment had two selection lines
with twelve jars each for a total of twenty-four jars per passage. One line was
a positive selection line where the two or three jars with the highest CH~4~
oxidation rate were homogenized to inoculate the next set of positive jars. The
other line was a neutral selection line where an equal number of jars as the
positive line were chosen at random to inoculate the next set of neutral jars.
The number of jars chosen was based on the distribution of fluxes among the
positive jars, i.e., we chose the top three jars unless only two jars had
considerably greater CH~4~ oxidation rates based on visual inspection of
histograms. The experiment was carried out over five passages until a
significant divergence in functional rates was observed between the two
selection treatments based on a difference of slopes.

The initial soil microbiome was sampled from the top 10 cm of an upland mineral
soil under a deciduous forest ecosystem near the University of Oregon campus in
Eugene, OR, USA. Incubations were performed in 500 mL mason jars with rubber
septa installed in the lids. Each jar was sterilized with 70\% ethanol to which
was added 45 g of autoclaved potting mix, 5 g of living soil, and 3.5 mL of
sterile deionized water to bring them to 60\% of field capacity. The potting
mix was Lane Potting Mix from Lane Forest Products (Eugene, OR, USA). Each jar
was then capped and injected with 4.3 mL of 99\% CH~4~ to bring
the headspace concentration to approximately 1000 ppm CH~4~. To create the two
treatment groups, twenty-four jars were created in an identical manner and then
randomly assigned to either the positive or neutral selection line. Jars were
flushed and respiked with methane twice per week to maintain aerobic conditions and elevated
CH~4~ concentrations and were incubated at ambient temperature for
approximately four weeks. Methane oxidation rates were determined at the end
of the incubation period. For the positive treatment, the three jars with the
greatest CH~4~ oxidation rates were chosen to inoculate the next generation.
For the neutral treatment, three jars were randomly selected to inoculate the
next generation. For each treatment, the selected jars were homogenized and 5 g
of the homogenized soil was used to inoculate the next set of jars, which
represents a 10% subsampling. The next set of jars were created in an identical
manner to the first generation with fresh autoclaved potting mix and the same
moisture and CH~4~ content.

## Methane oxidation rate

Methane oxidation rates were determined after flushing and spiking jars to
approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from
each jar immediately after spiking and then at time points 3, 6, 24, and 48
hours for a 5-point curve. Samples were immediately injected into a SRI model
8610C gas chromatograph equipped with a flame ionization detector (SRI
Instruments, Torrance, CA, USA) to determine the headspace CH~4~ concentration.
We applied a first-order exponential decay function to determine the rate
constant (k, units = d^−1^; i.e., dCH~4~/dt = k[CH~4~]) of the exponential
decrease in methane. Oxidation rates are presented as the additive inverse of
$k$ (i.e., $-k$) so that a more positive value represents a greater
oxidation rate. The jars selected in passage two for the positive treatment had
the lowest CH~4~ oxidation rate of the twelve jars due to a calculation error
in the rate constant. Despite this error, we still observed a strong
response to selection demonstrating the power of this technique. In addition,
this likely reintroduced diversity lost during the selection process,
potentially increasing the variation available to be selected upon, as
is recommended for microbiome selection experiments [@sanchez2021].

## Soil DNA extraction and sequencing

A subsample of soil from the starting inoculum and from every jar in passages 2
and 5 was collected and stored at $-80^\circ$C. Soil DNA was extracted from
0.25 g soil. Negative controls were extracted from autoclaved potting mix and
DNase-free water. Extractions were performed using the DNeasy PowerSoil kit
(QIAGEN, Düsseldorf, Germany) and quantified using Qubit dsDNA HS Assay Kit
(Thermo Fisher Scientific, Inc., Waltham, MA, USA). To estimate the diversity
and relative abundance of the bacterial and archaeal taxa in our soil
ecosystems, we sequenced the V4 region of the 16S rRNA gene using the 515F -
806R primer combination [@caporaso2011]. PCR mixtures were: 10 $\mu$l NEBNext
Q5 Hot Start HiFi PCR master mix, 9.2 $\mu$l primer mixture (1.09 $\mu$M
concentration), and 0.8 $\mu$l of DNA template. Reaction conditions were:
$98^\circ$C for 30 s (initialization); 35 cycles of $98^\circ$C for 10 s
(denaturation), $61^\circ$C for 20 s (annealing), and $72^\circ$C for 20 s
(extension); and $72^\circ$C for 2 m (final extension). Reactions were
performed in triplicate and then combined.  Amplicons were purified twice using
0.8x ratio Mag-Bind RxnPure Plus isolation beads (Omega Bio-Tek, Norcross, GA,
USA). Sequencing libraries were prepared using a dual-indexing approach
[@fadrosh2014; @kozich2013]. Amplicon concentrations were quantified using
Qubit and multiplexed at equimolar concentration. Sequencing was performed at
the University of Oregon Genomics Core Facility on the Illumina NovaSeq 6000
with paired-end 150 bp reads (Illumina, Inc., San Diego, CA, USA).

## Bioinformatics

Bioinformatics processing was performed in 'R' [@rcoreteam2018].  Demultiplexed
sequencing reads were denoised using 'DADA2' to generate a table of amplicon
sequence variants (ASVs) [@callahan2016]. Taxonomic assignment was performed
using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The
presence of contaminants was evaluated using both the prevalence and frequency
methods from 'DECONTAM' [@davis2018]. Decontam identified 
`r length(which(contamdf$contaminant))` potential contaminants based on
prevalence and frequency. Visual inspection of abundance-concentration plots
indicated that `r length(contaminants)` of these were likely contaminants and
these ASVs were removed. Amplicon sequence variants that were assigned
chloroplast or mitochondria taxonomy were removed prior to analysis.

## Statistical Analysis

Statistical analyses were performed in 'R' [@rcoreteam2018]. To determine
whether there was a significant change in CH~4~ oxidation rate as a response to
selection, we tested a difference in slopes between the positive and neutral
selection lines. Residuals did not meet the assumptions of constant variance
and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
transformed prior to analysis. First, we tested if there was a difference of
slopes between the positive and neutral treatments based on the interaction
between passage and treatment. To test the interaction, we fit two nested models
with and without the interaction term and compared them using an F-test with the
'anova' function. We then present the slopes for each treatment, which
represented the change in CH~4~ oxidation rate per passage as a response to
selection.

We estimated the proportion of variation in methane oxidation rate due to
variation in the microbiome as the regression of divergence between the positive
and neutral treatment on the cumulative selection differential [@falconer1996].
Divergence was calculated as the mean methane oxidation rate of the positive
treatment minus the mean methane oxidation rate of the neutral treatment in each
passage. The selection differential was calculated as the difference between
the mean of the three selected jars and the mean of all twelve jars in a passage.
Cumulative selection differential was then the sum of the selection differential
from all preceding selection events. We test this as the regression of
divergence on cumulative selection differential using the 'lm' function.

Richness was estimated using the method from @hurlbert1971 with a subsample
size of `r min_seqs` using the 'rarefy' function in 'vegan' [@oksanen2019]. We
tested a difference in richness by both passage and treatment with a
Kruskal-Wallace test followed by a pairwise Wilcoxon test. Next, we estimated
beta-diversity as the Bray-Curtis dissimilarity by averaging 100 random subsets
with a subsample size of `r min_seqs` using the 'avgdist' function in 'vegan'
[@bray1957; @oksanen2019].  We tested a difference in centroid and dispersion
of beta diversity by passage and treatment using a permutational analysis of
variance (PERMANOVA) with 999 permutations using the 'adonis2' function from
'vegan' and tested a difference of group dispersions using 'betadisper' and
'anova' with 999 permutations [@mcardle2001; @oksanen2019].  Lastly, we tested
the correlation between CH~4~ oxidation rate and Bray-Curtis dissimilarity in
Passage 5 with a distance-based redundancy analysis (dbRDA) using the 'dbrda'
function in 'vegan' and estimated the p-value using a permutation F-test with
999 permutations [@mcardle2001; @oksanen2019]

To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
differential abundance between the two treatments in passage 5. We first
grouped ASVs at the family level. Any ASVs that lacked a family-level
taxonomic assignment were grouped at a higher taxonomic level. We then subset the
samples in Passage 5 and removed all families with a prevalence of less than
10\% in either treatment. We used three methods for testing differential
abundance: ANCOM-II, ALDEx2, and CORNCOB 
[@kaul2017; @lin2020; @fernandes2014; @martin2020]. We then identified the
consensus taxa that were significant with all three tests and plotted their
relative abundances. For ANCOM-II, we used the 'ancom' function in the
'ANCOM-BC' package with a cutoff of W = 0.7 [@kaul2017; @lin2020]. For ALDEx2,
we used the 'aldex' function in the 'ALDEx2' package with Welch's t-test and we
used an effect size of 1 as our significance threshold [@fernandes2014].
Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob'
package with the Wald test and without bootstrapping [@martin2020]. Lastly, to
test differentially abundant methanotrophs, we subset all ASVs within
methanotrophic families and tested their differential abundance aggregated at
the family and genus level using 'corncob'.

# Results

## Response to selection on methane oxidation rate

We observed a response to artificial selection on whole-ecosystem soil CH~4~ oxidation
rate (`r figures("response", display = "cite")`; difference of slopes:
F~`r compare_response_mods[2,4]`,`r compare_response_mods[2,2]`~ = 
`r myround(compare_response_mods[2,5], 2)`,
p = `r myround(compare_response_mods[2,6], 2)`). At the start of the experiment,
the Positive treatment had a mean CH~4~ oxidation rate that was
`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`\%
lower than the Neutral treatment
`r print_lm(response_model, 3, 'difference of y-intercepts')`. 
There was no change in CH~4~ oxidation rate of the Neutral treatment over the
five passages `r print_lm(response_model, 2)`. By contrast, the Positive
treatment had a `r myround(per_slope, 1)`\% increase in CH~4~ oxidation rate per
passage `r print_lm(response_model, 4)`.

We estimated the proportion of variation in methane oxidation rate due to
variation in the microbiome by regressing the divergence between the positive
and neutral treatments against the cumulative selection differential. We
observed a slope of 0.40, which would indicate that 40\% of the variation is due
to the microbiome, but this was not significant (statistics).

We estimated the ecosystem heritability defined as the proportion of variation
in methane oxidation due rate to variation in the microbiome as the regression
of mid-recipient on mid-donor (`r figures("herit", display = "cite")`).
Recipient CH~4~ oxidation rates were correlated with donor CH~4~ oxidation rates
in both the Positive treatment
`r paste0('(', 'slope', ' = ', myround(tidy(herit_model)[2, 2] + tidy(herit_model)[4, 2], 2), ', SE = ', myround(tidy(herit_model)[4, 3], 2), ', t = ', myround(tidy(herit_model)[4, 4], 2), ', p = ', myround(tidy(herit_model)[4, 5], 2), ')')` and 
the Neutral treatment `r print_lm(herit_model, 2)`.

## Taxonomic richness

Median ASV richness decreased from `r myround(rare_median[1, 2], 1)` (`r myround(rare_median[1, 3], 1)`) in Passage 2 to `r myround(rare_median[2, 2], 1)` (`r myround(rare_median[2, 3], 1)`) in Passage 5 (Kruskal-Wallace test: $\chi^2$ = 
`r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; pairwise Wilcoxon test: p < 0.001). However, there was no
difference in richness between the Positive and the Neutral treatment in
Passage 2 or 5 (pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
correlation between richness and CH~4~ oxidation rate across the two treatments in Passage 5 (Spearman's rho = `r myround(s5[1,1], 1)`, p = `r myround(s5[1,3], 1)`).

## Community dissimilarity

Bray-Curtis dissimilarity of the soil microbiome varied strongly by passage and
weakly by treatment with an interaction between passage and treatment (`r figures("beta", display = "cite")`). 
Passage explained `r myround(100*tidy_permanova[1, 4])`\% of the
variation in Bray-Curtis dissimilarity (F~`r tidy_permanova[1, 2]`,`r tidy_permanova[4, 2]`~ = `r myround(tidy_permanova[1, 5], 1)`, p = `r myround(tidy_permanova[1, 6], 3)`), treatment explained 
`r myround(100*tidy_permanova[2, 4])`\%
of the variation (F~`r tidy_permanova[2, 2]`,`r tidy_permanova[4, 2]`~ = `r myround(tidy_permanova[2, 5], 1)`, p = `r myround(tidy_permanova[2, 6], 3)`), and the interaction
between treatment and passage explained 
`r myround(100*tidy_permanova[3, 4])`\% of the variation (F~`r tidy_permanova[3, 2]`,`r tidy_permanova[4, 2]`~ = `r myround(tidy_permanova[3, 5], 1)`, p = `r myround(tidy_permanova[3, 6], 3)`). There was 
no difference in dispersion between treatments or passages (F~3,44~ = 0.91, p = 0.45). 
Finally, CH~4~ oxidation rate was correlated with Bray-Curtis dissimilarity across both treatments in Passage 5 and explained `r myround(dbrda_percent_explained[1], 1)`\% of the variation in Bray-Curtis dissimilarity (dbRDA: F~`r dbrda_aov[1, 1]`,`r dbrda_aov[2, 1]`~ = `r myround(dbrda_aov[1, 3], 2)`, p = `r myround(dbrda_aov[1, 4], 3)`)

## Taxa that responded to selection

To identify taxa that responded
to selection on soil CH~4~ oxidation rate, we tested the differential relative
abundance of families in the Positive jars relative to the Neutral jars within
Passage 5 using three methods and then plotted the taxa identified by all three methods. We identified `r length(da_fams)`
families that were enriched or depleted in the Positive treatment relative to
the Neutral treatment (`r figures("da", display = "cite")`).

Overall, none of the families enriched in the Positive selection treatment
contain known methanotrophs. Several taxa identified had a 
higher taxonomic designation that contains methanotrophs, for example, the
Gammaproteobacteria class had a large effect size.
The Gammaproteobacteria include the type I and type X methanotrophs in the families *Methylococcaceae* and *Methylothermaceae* [@stein2012].
However, the Gammaproteobacteria is among the most diverse 
groups in the Prokaryotes, so this is not strong evidence for a selection response
by methanotrophs [@garrity2005].
In addition, the *Puniceicoccaceae* is a member of the phylum Verrucomicrobia. The Verrucomicrobia is a diverse group that contain
known methanotrophs as well as ammonia-oxidizing bacteria [@freitag2003]. Other than these two groups, none of the other taxa enriched in the Positive treatment are known to be related to methanotrophs. Two groups in the Armatimonadales were enriched in the Positive treatment including the family *Fimbriimonadaceae* and an unclassified ASV from the order Armatimonadales [@im2012].
*Cytophagaceae* was also enriched in
the Positive treatment and contains a number of mainly aerobic heterotrophs
that can digest a variety of macromolecules [@mcbride2014]. 
The remaining families include the uncultured family 0319-6G20, *Diplorickettsiaceae*, *Rhodospirillaceae*, and an unclassified Kapabacteriales.

We did not identify any methanotrophic families in the overall differential abundance analysis. 
However, we wanted to look more closely at the known methanotrophs in our dataset to be sure they did not have an effect.
To do this, we subset all of the ASVs in our dataset that were in families that contained methanotrophs.
Only two families were represented: *Methylacidiphilaceae* and *Beijerinckiaceae*. 
Aggregating reads at the family level, neither family was differentially abundant between the two treatments.
However, aggregated at the genus level, a group of unclassified genera in the *Beijerinckiaceae* were depleted in the Positive treatment and the genus
*Rhodoblastus*, a member of the *Beijerinckiaceae*, was enriched in the
Positive treatment.  While many *Beijerinckiaceae* are methanotrophs, several
taxa in this family have lost the the ability to oxidize CH~4~ and it appears
that *Rhodoblastus* species are not able to grow on CH~4~, though they can grow
on methanol [@tamas2014, @dedysh2016]. Based on this analysis,
it appears that no methanotrophs were enriched in the Positive treatment.

# Discussion

In this study, we addressed whether variation in the microbiome contributes to
variation in the rate of ecosystem function. To test this, we applied artificial
ecosystem selection to soil microbiomes by selecting on the whole-ecosystem soil
methane oxidation rate. We observed an increase in methane oxidation rate in the
selection treatment relative to the control, which demonstrates that there was a
response to selection. Given that we observed a response to selection, we
conclude that variation in the microbiome contributes to variation in the
methane oxidation rate independent of the environment.

The ultimate goal of this study (and much of microbial ecology) could be
summarized as "Do microbes matter to ecosystem function (or health)?" and there
are at least three ways we could address that question. One would be to ask
whether microorganisms mediate function. This question is perhaps trivial. In
the case of methane emissions from soil, we know that microorganisms produce and
consume methane. We could also ask what would happen if we removed the
microorganisms. Again, this question is trivial. There are no known abiotic
reactions in soil that consume methane (outside of the troposphere, the
stratosphere, and the marine boundary layer @kirschke2013). Therefore, removing
the microbiome would likely eliminate methane production and consumption. Then,
the third question we could ask is whether *variation* in the microbiome is
associated with *variation* in the rate of ecosystem function. The answer to
this question is not so obvious.

If we look at variation in the rate of a microbially-mediated ecosystem function
across a set of habitats, that variation could be associated with variation in
the abiotic environmental conditions, variation in the composition of
microbiomes, or some interaction or covariance between the two (see Figure 5).
To address the question of whether variation in the microbiome matters for
ecosystem function, we need to identify the component of variation uniquely
associated with the microbiome. If we sample natural variation, as in the
comparative approach, it can be difficult to distinguish between the direct
effect of the microbiome and the effect of the microbiome due to covariance with
the environment. Likewise, with the manipulative approach even though we are
placing ecosystem propagules into a common environment or substrate, the
propagule itself carries with it environmental conditions that
cannot be removed. Generating variation in microbiome composition through
filtering or other similar means often confounds composition with other factors,
such as biomass or cell size. To overcome these limitations and to isolate
the direct effect of the microbiome on ecosystem function, we used
artificial ecosystem selection.

To perform artificial ecosystem selection, we generate a set of ecosystems in
bottles using a common, sterile substrate and inoculate them with a propagule
containing a diverse microbiome that is a subsample of an existing microbiome.
This generates variation in microbiome composition among the ecosystems. We can
then measure the rate of an ecosystem function for each bottle and select the
bottles with the highest (or lowest) rate of ecosystem function. The selected
ecosystems are then homogenized and used as the propagule to inoculate a new set of
ecosystems containing the same sterile substrate as the initial set. By
subsampling the microbiome, we can generate variation in the composition of the
microbiome that is independent of cell size or biomass. And by passaging the
microbiome through the same sterile substrate over multiple rounds of selection,
we can progressively minimize the effect of the abiotic conditions from the
donor environment on both the rate of function and the composition of the
microbiome. In this way, over multiple rounds of selection, we can observe
whether varying the microbiome can generate variation in the rate of our
ecosystem function. If we observe no change in ecosystem function from
passage to passage despite our imposed selection, we would conclude that this
ecosystem function in this set of ecosystems is driven primarily by
environmental variation or technical error among the replicates. However, if we
observe a response to selection, that is, a shift in the mean ecosystem function
between successive passages, we would conclude that variation in the microbiome can
generate variation in the rate of ecosystem function. And that is, in fact, what
we observed.

Given that variation in the microbiome is associated with variation in ecosystem
function, a reasonable follow-up question is "How much variation?" This is a
challenging question to answer since the value depends on the population of
ecosystems that is sampled and the estimate for one set of ecosystems may not
apply to another set. For example, we cannot estimate a percent of variation
explained in tropical rainforests and then assume that number applies in polar
wetlands. In addition, the magnitude of the estimate will depend on how much
variation is sampled. A large $\frac{V_M}{V_F}$ is not very interesting if $V_F$
is small.  Keeping these caveats in mind, one way to estimate this number is to
determine how much the recipient jars resemble the selected donor jars that were
used to inoculate them. We can calculate the response to selection as the
difference between two successive passages in their mean methane oxidation rate.
We will denote this as $R$. We can also calculate the strength of selection as
the difference in mean methane oxidation rate between the twelve jars in one
generation and the 3 jars chosen for selection in that generation, which we will
call the selection differential and denote as $S$. If we plot the cumulative
$R$ against the cumulative $S$, the slope of this relationship will equal the
proportion of variation explained by the microbiome.

Using this approach, we observed a relatively large percent of variation
explained by the microbiome, but this number was not significantly different
from zero.

The concept that we have just described is analogous to the concept of
heritability from quantitative genetics. Heritability for given a population is
defined as the proportion of variation in a phenotype associated with variation
in genotype [@lynch1998]. This concept has been applied above the population
level for groups within a species [@goodnight2000], as well as communities and
ecosystems [@whitham2003; @whitham2006]. One study previously estimated the
heritability of a community trait after artificial ecosystem selection using
parent-offspring regression and observed a relatively large heritability of 50\%
[@blouin2015]. However, parent-offspring regression assumes a randomly mating
population, which, by definition, does not apply to a selection experiment. In
the present study, we estimated the variation explained by the microbiome
somewhat differently using the breeder's equation ($R = S h^2$) and regressing
the observed response to selection against the imposed selection differential
(Figure 1; @falconer1996). We also observed a relatively large response to selection
(40\%), though this estimate had large error bars and was not significant so we
did not have the power to estimate an accurate percent of variation explained
for this experiment.

Asking how much of the variation in ecosystem function is due to variation in
the microbiome is a silly question. We cannot answer this questions broadly
for all communities or all ecosystems, because the answer depends in large part
on the magnitude of variation sampled.

What do I want to get across here:

This estimate is analogous to the concept of heritability from quantitative
genetics
Define heritability
Point out that it's been applied to communities and ecosystems
Mention the different ways of estimating heritability
Mention the assumptions inherent in some of these estimates (non-assortative
mating; particularly numbers of shared alleles based on kinship)
Point out that while heritability is a useful framework to borrow from in order
to describe the components of variation, there are limitations in its
application
Propose that we pursue new theory about how to understand what drives variation
in microbiome functions and microbiome-encoded traits

The challenge with applying the concept of heritability outside its original
scope is that it is unclear if the assumptions underlying these estimates apply
to microbiomes. For a genomic trait, we can rely on the assumption that
offspring share, on average, 50\% of the alleles of each of their
parents. It is less clear if offspring share 50\% of the microbes of each of
their parents or if it is sufficient for 50\% of the microbes to be _present_ or
if they must be in a similar relative abundance. Estimates of heritability, such
as parent-offspring regression and the ACE model rely on specific estimates of
the similarity of relatives based on their relatedness. It is unclear if these
assumptions apply to the microbiomes of host ecosystems. It is even less clear
whether they apply to non-host ecosystems that do not reproduce through sexual
recombination.

Given that variation in the microbiome is driving the observed variation in
methane oxidation rate, we next wanted to address what aspects of the
microbiome, in terms of alpha diversity, beta diversity, or the relative
abundance of taxa, explain this contribution of the microbiome to variation in
methane oxidation rate. There are three ways that the microbiome could respond
to selection in this experiment: gain or loss of species, changes in the
relative abundances of species, or changes within the genomes of the constituent
species. Using 16S rRNA gene sequences, we are unable to address the third
possibility of whether taxa in this experiment could have evolved changes in
their genomes as a result of selection.  Therefore, we will focus on the first
two possibilities.

Richness at the ASV level did not vary between the two treatments and there
were relatively few taxa gained or lost in the Positive selection treatment and
none of these were prevalent across the 12 jars in passage 5. Therefore, the
gain or loss of species is unlikely to explain the increase in methane
oxidation rate. However, we found that Bray-Curtis dissimilarity was greater
between the two treatments in Passage 5 than within each treatment and was
correlated with methane oxidation rate, which suggests that changes in the
relative abundance of taxa could explain the response to selection.

Even though we observed an increase in methane oxidation rate in the Positive
treatment and a difference in composition between the two treatments, we did
not observe an increase in the relative abundance of methanotrophs. This was
surprising given that methane consumption is not a common trait among microbes
and that it is often assumed that the rate of an ecosystem function is limited
by the final enzymatic step in a metabolic pathway [@rocca2015]. In certain
ecosystems, methane production and consumption are correlated with the
abundance of methanogens and methanotrophs as estimated from marker genes
[@meyer2020; @meyer2020a]. However, it does not appear generally true that the
rate of an ecosystem function is limited by the abundance of the gene that
encodes the final step in that pathway [@rocca2015]. Our results suggest that
ecosystem-scale methane oxidation rates can be altered by non-methanotrophs,
perhaps through ecological interactions with methanotrophic species.

Another important question is whether selection at the ecosystem scale resulted
in within-strain evolution. Using 16S rRNA gene sequencing we are unable to
resolve genomic changes within taxa that may have undergone evolution.
Therefore, it is possible that the response to selection was driven by changes
within the genomes of methanotrophs. But given the slow growth rate of
methanotrophs, often requiring weeks to years of serial batch culturing for
isolation, this seems like an unlikely explanation 
[@whittenbury1970; @hirayama2014; @kim2018]. Future work could analyze an
artificial ecosystem selection experiment using genome-resolved metagenomics to
try to detect strain-level changes in the constituent taxa. However,
reconstructing genomes from in situ soil samples has been difficult to achieve
and may require enrichment techniques, particularly for methanotrophs which are
a relatively rare component of the soil microbiome [@delmont2015]. If these
challenges could be overcome for soil metagenomics, we could begin to evaluate
whether ecosystem selection results in strain-level evolution or enriches for
certain metabolic pathways within the microbiome.

A fundamental question in microbial ecology is whether variation in the
microbiome contributes to variation in ecosystem function. We performed
artificial ecosystem selection on soil methane oxidation rate to test the
relationship between microbiome variation and soil methane oxidation rate. We
observed a response to selection on soil methane oxidation rate and found a
high heritability for methane oxidation, which suggests that variation in the
microbiome contributes to variation in soil methane oxidation rate
independent of the environment. This shows that manipulating the microbiome
could be a viable strategy for mitigating or enhancing certain
ecosystem functions.  Surprisingly, we did not observe an increase in the
relative abundance of methanotrophs as a response to selection. This
suggests that the rate of methane oxidation could be altered by
non-methanotrophs, possibly through ecological interactions with methanotrophs.
Future research should investigate other aspects of microbiome variation, such
as strain-level evolution or functional gene composition, to begin to describe
a more detailed mapping between the microbiome and ecosystem function. This
study demonstrates that soil methane oxidation rate can vary greatly with
variation in the microbiome and opens up future research opportunities to
establish a more detailed understanding of the relationship between the
microbiome and ecosystem function.

# Acknowledgments

This project was supported by the National Science Foundation Graduate Research
Fellowship Program (grant no. DGE 1255832) and the ARCS Foundation Florence and
Mike Nudelman Scholarship.

# Competing Interests

We declare we have no competing interests.

# Data Availability Statement

The 16S rRNA sequencing data generated during the current study are available
in the NCBI Sequence Read Archive (SRA) under BioProject accession number
PRJNA832314, https://www.ncbi.nlm.nih.gov/sra/PRJNA832314. The metadata
generated during the current study as well as the scripts to recreate the
analysis are available on Github,
https://github.com/amorris28/artificial_ecosystem_selection.

# Supplement

# References

::: {#refs}
:::

# Figures

```{r response, fig.width=4, fig.height=3}

p_response

```

`r figures("response", "Response to selection on soil CH~4~ oxidation rate. The y-axis is the log~10~ of the additive inverse of the first-order exponential decay constant $k$ (i.e., $-k$) with units day^-1^. The x-axis is the passage number. The orange line and points are the Positive treatment and the gray line and points are the Neutral treatment.")`

```{r herit, fig.width=5, fig.height=3}

p_herit

```

`r figures("herit", paste0("Ecosystem heritability as the regression of mid-recipient methane flux on mid-donor methane flux ($-k$ with units day^-1^). Donor is the mean oxidation rate of the three jars selected to inoculate the next passage. Recipient is the mean oxidation rate of the 12 jars in the next passage. The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment. Slope of the regression line estimates the proportion of variation in methane oxidation rate due to variation in the microbiome. For the Positive treatment, the variance explained by the microbiome was ", myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100, 1), "% (slope = ", myround(coefficients(herit_model)[2] + coefficients(herit_model)[4], 3), ", SE = ", myround(tidy(herit_model)$std.error[4], 2), ") and for the Neutral treatment the variance explained by the microbiome was ", myround(coefficients(herit_model)[2]*100, 1), "% (slope = ", myround(coefficients(herit_model)[2], 3), ", SE = ", myround(tidy(herit_model)$std.error[2], 2), ")."))`

```{r beta, results=FALSE, fig.height=2.5}

p_beta

```

`r figures("beta", "Principal component analysis plot of beta diversity for all jars. Dissimilarities are based on rarefied Bray-Curtis dissimilarity. Colors represent treatment and shapes represent passage number.")`

```{r da, results=FALSE}

p_log10

```

`r figures("da", "Differentially abundant family-level taxa identified by ANCOM-II, ALDEx2, and CORNCOB. Values on the x-axis are relative abundances on a log~10~ scale. Points are colored by treatment. Large points with error bars are the median relative abundance and interquartile range. Taxa in the top panel are enriched in the Positive treatment and taxa in the bottom panel are depleted in the Positive treatment. Taxa are sorted by their effect size with taxa at the top having the largest positive effect size and taxa at the bottom with the largest negative effect size.")`

![Figure 5. Variation in the rate of an ecosystem function is potentially the result of at least three components: variation in the abiotic environmental conditions, variation in microbiome composition, and the covariance between microbiomes and the environment. It is difficult with traditional study designs to isolate the direct effect of the microbiome from the effect of the environment via covariance with the microbiome. Here, we attempt to isolate the effect of the microbiome through artificial selection on microbiome composition. For simplicity, we omitted the reverse arrows as well as the interaction between microbiome and the environment, though these relationshps may also exist.](Figure1.png)