### 2023-02-20 manuscript/manuscript.Rmd

---
title: Artificial ecosystem selection reveals relationships between microbiome composition and ecosystem function
author: "Andrew H. Morris^1^* and Brendan J. M. Bohannan^1^"
output:
  word_document:
    reference_docx: reference.docx
bibliography: bibliography.bib
csl: vancouver.csl
---

{title max characters 150; current = 106 w/spaces}

^1^ Institute of Ecology and Evolution, University of Oregon, Eugene, OR, USA  
\* Correspondence: Andrew H. Morris, amorris3\@uoregon.edu  
Competing Interests: We declare we have no competing interests.

```{r options, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir=normalizePath('../'))

```

```{r initialize-packages}
library(tidyverse)
library(broom)
library(broman)
library(morris)
# library(phyloseq)
library(ggpubr)
library(knitr)
library(corncob)
# library(ggthemes)
library(captioner)
library(knitr)
library(ggtext)
library(patchwork)
figures <- captioner()
tables <- captioner(prefix = "Table")

source('R/functions.R')
```

```{r load-data}
heritability <- read_tsv(paste0(der_dir, 'heritability.tsv'))
fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))

load(file = paste0(der_dir, 'decontam.RData'))

load(file = paste0(der_dir, 'plots.RData'))
load(file = paste0(der_dir, 'models.RData'))

# # CH4 models
# load(file = paste0(der_dir, 'response_model.RData'))
# load(file = paste0(der_dir, 'response_plot.RData'))
# load(file = paste0(der_dir, 'herit_model.RData'))
# load(file = paste0(der_dir, 'herit_plot.RData'))
# 
# # 16S models
# load(file = paste0(der_dir, 'richness_models.RData'))
# load(file = paste0(der_dir, 'phyloseq.RData'))
# load(file = paste0(der_dir, 'beta_model.RData'))
# load(file = paste0(der_dir, 'beta_fig.RData'))
# load(file = paste0(der_dir, 'da_corncob.RData'))
# load(file = paste0(der_dir, 'da_fig.RData'))
# load(file = paste0(der_dir, 'dbrda.RData'))

```

# Abstract

{max 250 words; current = 249}

Microbiomes mediate important ecosystem functions, yet it has proven difficult
to determine the relationship between microbiome composition and the rate of
ecosystem functions. This challenge remains because it is difficult to
manipulate microbiome composition directly, we often cannot know _a priori_ which
microbiome members influence the rate of an ecosystem function, and
microbiomes can covary strongly with other drivers of ecosystem function, such
as the environment. We propose that artificial ecosystem selection can overcome
these challenges. Artificial ecosystem selection involves creating replicate
ecosystems, selecting replicates with a desired ecosystem-level trait, and
using the microbiomes from the selected ecosystems to inoculate a new set of
ecosystems. We performed artificial ecosystem selection on replicate soil
microbiomes to select for increased soil methane oxidation rate. We observed a
strong response to selection with a `r myround(per_slope, 1)`\% increase in
methane oxidation rate per passage. We
estimated that 40\% of the variation in methane oxidation
rate in these soils can be attributed to microbiome variation. We also found
that selection did not enrich for known methane oxidizers; instead, `r length(da_fams)` families not
known to oxidize methane, including *Fimbriimonadaceae*, *Cytophagaceae*, and
*Diplorickettsiaceae*, were enriched by selection. This result is in contrast to the typical assumption that
the rate of an ecosystem function is limited by the final step in the associated
microbial pathway. Our study demonstrates that variation in microbiome
composition can contribute to variation in the rate of ecosystem function and
that manipulating microbiome composition could be a viable strategy for managing
ecosystem functions.

{max text 5000 words; current text = 4549}

# Introduction

Microbiomes mediate a variety of important ecosystem functions relevant to human
health, agriculture, and global change. As a result, there is great interest in
understanding how to manipulate the microbiome to achieve desirable outcomes
within these domains [@conrad1996; @schimel1998; @crowther2019]. However, for
microbiome manipulations to be successful, variation in the microbiome must
contribute directly to variation in the magnitude of the function of interest
independent of other factors. Many studies have attempted to document such a
relationship
[@cavigelli2000; @balser2005; @waldrop2006; @strickland2009; @wagg2014; @wagg2019].
However, it is difficult to isolate the direct effect of variation in microbiome
composition from other drivers of variation in ecosystem function, such as the
indirect effect of the environment on function through microbiome assembly.
Here, we overcome these limitations by using a selection approach to estimate
the degree to which an ecosystem function varies with microbiome composition.

There have been two general categories of approaches that investigators have
used to estimate the degree to which an ecosystem function varies with
microbiome composition: comparative and manipulative. Comparative
studies sample natural variation in an ecosystem function across different
habitats and simultaneously measure variation in community composition.
Investigators can then correlate ecosystem function with aspects of community
composition while attempting to control for environmental variation. These
approaches have
documented important relationships between microbiomes and ecosystem
functions. For example, a meta-analysis of these studies observed a small but
significant contribution of the microbiome to variation in ecosystem
function after controlling for environmental variation [@graham2016]. In
addition, studies focusing on the correlation between the rate of an ecosystem
function and the abundance of an associated marker gene (i.e., a gene that codes
for a protein assumed to be involved in the ecosystem function) sometimes observe a
significant correlation, though this relationship is rare and contingent upon
both the function and the ecosystem sampled [@rocca2015]. However, comparative
studies come with some unique challenges and limitations. One issue is that
microbiome attributes tend to covary with the abiotic conditions within an
environment, and whether and how the investigator controls for these variables
can affect the conclusions of such a comparative study. In addition, it is
difficult to know _a priori_ which environmental variables or community attributes
to measure. Finally, while these approaches can establish a potential magnitude
and direction for these relationships, it is often difficult to identify the
taxa or genes that explain the connection between composition and function.

The other broad category of approaches used to address this question are
manipulative approaches. Manipulative experiments try to alter microbial
community composition and observe the effect on function. For example,
reciprocal transplant and common garden experiments have shown that microbiomes
originating from different ecosystems inoculated into the same substrate or
introduced into a common environment display distinct functional rates 
[@cavigelli2000; @balser2005; @waldrop2006; @strickland2009]. In addition,
manipulating diversity by filtering communities by cell size or through dilution
has been shown to alter the rate of ecosystem functions 
[@wagg2014; @wagg2019; @griffiths2001]. However, manipulating the microbiome
directly is challenging. Each of the manipulative approaches applied confound
community composition with other factors. For example, reciprocal transplant and
common garden experiments can confound community composition with the abiotic
conditions introduced with the inoculum, while manipulating composition through
dilution may confound composition with biomass [@reed2007].

In this study, we sought to build on the observations of comparative and
manipulative studies by applying a different approach to the question of whether
microbiome variation contributes to variation in the rate of an ecosystem
function. We used artificial ecosystem selection to select for microbiomes that
performed a greater rate of ecosystem function
[@swenson2000; @blouin2015; @panke-buisse2015]. We then tested whether variation
in the microbiome contributed to variation in the rate of ecosystem function and
identified microbiome attributes that might explain this relationship. There are
several potential advantages to this approach for documenting
microbiome-function relationships and for investigating the mechanisms
underlying those relationships. By passaging microbiomes over multiple rounds
using a common environment, we can weaken the covariance between microbes and
the environment by repeatedly diluting abiotic effects. In addition, by
inoculating our ecosystems with equally-sized subsamples, we eliminate the need
to generate microbiome variation through methods that are confounded with
biomass or cell size. Lastly, by comparing our artificially selected community
to a control community resulting from random selection, we can both control for
changes in the environment over time and identify genes or taxa that are
associated with the ecosystem function under selection.

We applied artificial ecosystem selection to soil microbiomes by selecting on
soil methane oxidation rate. We chose this function because methane is a
globally important greenhouse gas and methane oxidation by soil bacteria is the
primary biological sink for atmospheric methane [@kirschke2013]. In addition,
there is evidence that soil methane oxidation rate may vary with microbiome
composition based on comparative studies in a variety of arctic and tropical
ecosystems [@meyer2020; @meyer2020a; @freitag2009; @freitag2010] as well as
studies that manipulate methanotroph richness [@schnyder2018]. Finally,
methanotrophy is one of the most deeply conserved microbial physiologies and is
represented in a narrow range of taxa, which suggests that the taxonomic
composition of the microbiome is more likely to be associated with the rate of
methane oxidation than other broader or more shallowly conserved functions
[@schimel1998; @martiny2013].

In this study, we used artificial ecosystem selection on methane oxidation rate
to address the following questions: Does variation in the relative abundance of
microbial taxa contribute to variation in soil methane oxidation rate in our
system? Which attributes of the microbiome are associated with variation in
methane oxidation rate, and do these attributes match our assumptions about the
factors that regulate methane oxidation rate in nature?

# Materials and Methods

## Experimental design

We performed an artificial ecosystem selection experiment 
[*sensu* @swenson2000] by passaging replicate soil microbiomes. The trait we
selected on was methane (CH~4~)-oxidation rate. This experiment had two selection lines
with twelve jars each for a total of twenty-four jars per passage. One line was
a positive selection line where the two or three jars with the highest CH~4~
oxidation rate were homogenized to inoculate the next set of positive jars. The
other line was a neutral selection line where an equal number of jars as the
positive line were chosen at random to inoculate the next set of neutral jars.
The number of jars chosen was based on the distribution of fluxes among the
positive jars, i.e., we chose the top three jars unless only two jars had
considerably greater CH~4~ oxidation rates based on visual inspection of
histograms. The experiment was carried out over five passages until a
significant divergence in functional rates was observed between the two
selection treatments based on a difference of slopes.

The initial soil microbiome was sampled from the top 10 cm of an upland mineral
soil under a deciduous forest ecosystem near the University of Oregon campus in
Eugene, OR, USA. Incubations were performed in 500 mL mason jars with rubber
septa installed in the lids. Each jar was sterilized with 70\% ethanol to which
was added 45 g of autoclaved potting mix, 5 g of living soil, and 3.5 mL of
sterile deionized water to bring them to 60\% of field capacity. The potting
mix was Lane Potting Mix from Lane Forest Products (Eugene, OR, USA). Each jar
was then capped and injected with 4.3 mL of 99\% CH~4~ to bring
the headspace concentration to approximately 1000 ppm CH~4~. To create the two
treatment groups, twenty-four jars were created in an identical manner and then
randomly assigned to either the positive or neutral selection line. Jars were
flushed and respiked with CH~4~ twice per week to maintain aerobic conditions
and elevated CH~4~ concentrations and were incubated at ambient temperature for
approximately four weeks. Methane oxidation rates were determined at the end of
the incubation period. For the positive treatment, the three jars with the
greatest CH~4~ oxidation rates were chosen to inoculate the next generation.
For the neutral treatment, three jars were randomly selected to inoculate the
next generation. For each treatment, the selected jars were homogenized and 5 g
of the homogenized soil was used to inoculate the next set of jars, which
represents a 10% subsampling. The next set of jars were created in an identical
manner to the first generation with fresh autoclaved potting mix and the same
moisture and CH~4~ content.

## Methane oxidation rate

Methane oxidation rates were determined after flushing and spiking jars to
approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from
each jar immediately after spiking and then at time points 3, 6, 24, and 48
hours for a 5-point curve. Samples were immediately injected into a SRI model
8610C gas chromatograph equipped with a flame ionization detector (SRI
Instruments, Torrance, CA, USA) to determine the headspace CH~4~ concentration.
We applied a first-order exponential decay function to determine the rate
constant (k, units = d^−1^; i.e., dCH~4~/dt = k[CH~4~]) of the exponential
decrease in methane. Oxidation rates are presented as the additive inverse of
$k$ (i.e., $-k$) so that a more positive value represents a greater
oxidation rate. The jars selected in passage two for the positive treatment had
the lowest CH~4~ oxidation rate of the twelve jars due to a calculation error
in the rate constant. Despite this error, we still observed a strong
response to selection demonstrating the power of this technique. In addition,
this likely reintroduced diversity lost during the selection process,
potentially increasing the variation available to be selected upon, as
is recommended for microbiome selection experiments [@sanchez2021].

## Soil DNA extraction and sequencing

A subsample of soil from the starting inoculum and from every jar in passages 2
and 5 was collected and stored at $-80^\circ$C. Soil DNA was extracted from
0.25 g soil. Negative controls were extracted from autoclaved potting mix and
DNase-free water. Extractions were performed using the DNeasy PowerSoil kit
(QIAGEN, Düsseldorf, Germany) and quantified using Qubit dsDNA HS Assay Kit
(Thermo Fisher Scientific, Inc., Waltham, MA, USA). To estimate the diversity
and relative abundance of the bacterial and archaeal taxa in our soil
ecosystems, we sequenced the V4 region of the 16S rRNA gene using the 515F -
806R primer combination [@caporaso2011]. PCR mixtures were: 10 $\mu$l NEBNext
Q5 Hot Start HiFi PCR master mix, 9.2 $\mu$l primer mixture (1.09 $\mu$M
concentration), and 0.8 $\mu$l of DNA template. Reaction conditions were:
$98^\circ$C for 30 s (initialization); 35 cycles of $98^\circ$C for 10 s
(denaturation), $61^\circ$C for 20 s (annealing), and $72^\circ$C for 20 s
(extension); and $72^\circ$C for 2 m (final extension). Reactions were
performed in triplicate and then combined.  Amplicons were purified twice using
0.8x ratio Mag-Bind RxnPure Plus isolation beads (Omega Bio-Tek, Norcross, GA,
USA). Sequencing libraries were prepared using a dual-indexing approach
[@fadrosh2014; @kozich2013]. Amplicon concentrations were quantified using
Qubit and multiplexed at equimolar concentration. Sequencing was performed at
the University of Oregon Genomics Core Facility on the Illumina NovaSeq 6000
with paired-end 150 bp reads (Illumina, Inc., San Diego, CA, USA).

## Bioinformatics

Bioinformatics processing was performed in 'R' [@rcoreteam2018].  Demultiplexed
sequencing reads were denoised using 'DADA2' to generate a table of amplicon
sequence variants (ASVs) [@callahan2016]. Taxonomic assignment was performed
using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The
presence of contaminants was evaluated using both the prevalence and frequency
methods from 'DECONTAM' [@davis2018]. Decontam identified 
`r length(which(contamdf$contaminant))` potential contaminants based on
prevalence and frequency. Visual inspection of abundance-concentration plots
indicated that `r length(contaminants)` of these were likely contaminants and
these ASVs were removed. Amplicon sequence variants that were assigned
chloroplast or mitochondria taxonomy were removed prior to analysis.

## Statistical Analysis

Statistical analyses were performed in 'R' [@rcoreteam2018]. To determine
whether there was a significant change in CH~4~ oxidation rate as a response to
selection, we tested a difference in slopes between the positive and neutral
selection lines. Residuals did not meet the assumptions of constant variance
and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
transformed prior to analysis. First, we tested if there was a difference of
slopes between the positive and neutral treatments based on the interaction
between passage and treatment. To test the interaction, we fit two nested models
with and without the interaction term and compared them using an F-test with the
'anova' function. We then present the slopes for each treatment, which
represented the change in CH~4~ oxidation rate per passage as a response to
selection.

We estimated the proportion of variation in methane oxidation rate due to
variation in the microbiome as the regression of divergence between the positive
and neutral treatment on the cumulative selection differential [@falconer1996].
Divergence was calculated as the mean methane oxidation rate of the positive
treatment minus the mean methane oxidation rate of the neutral treatment in each
passage. The selection differential was calculated as the difference between
the mean of the three selected jars and the mean of all twelve jars in a passage.
Cumulative selection differential was calculated as the sum of the selection
differential from all preceding selection events. We then regressed cumulative
divergence on cumulative selection differential using the 'lm' function.

Richness was estimated using the method from @hurlbert1971 with a subsample
size of `r min_seqs` calculated via the 'rarefy' function in 'vegan' [@oksanen2019]. We
tested a difference in richness by both passage and treatment with a
Kruskal-Wallace test followed by a pairwise Wilcoxon test. Next, we estimated
beta-diversity as the Bray-Curtis dissimilarity by averaging 100 random subsets
with a subsample size of `r min_seqs` using the 'avgdist' function in 'vegan'
[@bray1957; @oksanen2019].  We tested a difference in centroid and dispersion
of beta diversity by passage and treatment using a permutational analysis of
variance (PERMANOVA) with 999 permutations using the 'adonis2' function from
'vegan' and tested a difference of group dispersions using 'betadisper' and
'anova' with 999 permutations [@mcardle2001; @oksanen2019].  Lastly, we tested
the correlation between CH~4~ oxidation rate and Bray-Curtis dissimilarity in
Passage 5 with a distance-based redundancy analysis (dbRDA) using the 'dbrda'
function in 'vegan' and estimated the p-value using a permutation F-test with
999 permutations [@mcardle2001; @oksanen2019]

To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
differential abundance between the two treatments in passage 5. We first
grouped ASVs at the family level. Any ASVs that lacked a family-level
taxonomic assignment were grouped at a higher taxonomic level. We then subset the
samples in Passage 5 and removed all families with a prevalence of less than
10\% in either treatment. We used three methods for testing differential
abundance: ANCOM-II, ALDEx2, and CORNCOB 
[@kaul2017; @lin2020; @fernandes2014; @martin2020]. We then identified the
consensus taxa that were significant with all three tests and plotted their
relative abundances. For ANCOM-II, we used the 'ancom' function in the
'ANCOM-BC' package with a cutoff of W = 0.7 [@kaul2017; @lin2020]. For ALDEx2,
we used the 'aldex' function in the 'ALDEx2' package with Welch's t-test and we
used an effect size of 1 as our significance threshold [@fernandes2014].
Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob'
package with the Wald test and without bootstrapping [@martin2020]. Lastly, to
test differentially abundant methanotrophs, we subset all ASVs within
methanotrophic families and tested their differential abundance aggregated at
the family and genus level using 'corncob'.

# Results

## Response to selection on methane oxidation rate

We observed a response to artificial selection on whole-ecosystem soil CH~4~ oxidation
rate (`r figures("response", display = "cite")`; difference of slopes:
F~`r compare_response_mods[2,4]`,`r compare_response_mods[2,2]`~ = 
`r myround(compare_response_mods[2,5], 2)`,
p = `r myround(compare_response_mods[2,6], 2)`). At the start of the experiment,
the Positive treatment had a mean CH~4~ oxidation rate that was
`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`\%
lower than the Neutral treatment
`r print_lm(response_model, 3, 'difference of y-intercepts')`. 
There was no change in CH~4~ oxidation rate of the Neutral treatment over the
five passages `r print_lm(response_model, 2)`. By contrast, the Positive
treatment had a `r myround(per_slope, 1)`\% increase in CH~4~ oxidation rate per
passage `r print_lm(response_model, 4)`.

To estimate the proportion of variation in CH~4~ oxidation rate due to
variation in microbiome composition, we regressed divergence between the
positive selection treatment and the control against the cumulative selection
differential (`r figures("herit", display = "cite")`). This estimate is analogous
to estimates of "microbiability" from the animal breeding literature, which
quantifies the
variation in a host trait that is due to microbiome variation [@difford2016]. 
The slope of the regression of divergence on cumulative selection differential
provides an estimate of realized microbiability (h^2^ &pm; SE), which in our
experiment was `r myround(tidy(rh_model)[[2, 2]], 2)` &pm;
`r myround(tidy(rh_model)[[2, 3]], 2)`, though this was not significant (F~`r glance(rh_model)$df`,`r glance(rh_model)$df.residual`~ = 
`r myround(glance(rh_model)$statistic, 2)`, 
p = `r myround(glance(rh_model)$p.value, 2)`).

## Taxonomic richness

Median ASV richness decreased from `r myround(rare_median[1, 2], 1)` (`r myround(rare_median[1, 3], 1)`) in Passage 2 to `r myround(rare_median[2, 2], 1)` (`r myround(rare_median[2, 3], 1)`) in Passage 5 (Kruskal-Wallace test: $\chi^2$ = 
`r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; pairwise Wilcoxon test: p < 0.001). However, there was no
difference in richness between the Positive and the Neutral treatment in
Passage 2 or 5 (pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
correlation between richness and CH~4~ oxidation rate across the two treatments in Passage 5 (Spearman's rho = `r myround(s5[1,1], 1)`, p = `r myround(s5[1,3], 1)`).

## Community dissimilarity

Bray-Curtis dissimilarity of the soil microbiome varied strongly by passage and
weakly by treatment with an interaction between passage and treatment (`r figures("beta", display = "cite")`). 
Passage explained `r myround(100*tidy_permanova[1, 4])`\% of the
variation in Bray-Curtis dissimilarity (F~`r tidy_permanova[1, 2]`,`r tidy_permanova[4, 2]`~ = `r myround(tidy_permanova[1, 5], 1)`, p = `r myround(tidy_permanova[1, 6], 3)`), treatment explained 
`r myround(100*tidy_permanova[2, 4])`\%
of the variation (F~`r tidy_permanova[2, 2]`,`r tidy_permanova[4, 2]`~ = `r myround(tidy_permanova[2, 5], 1)`, p = `r myround(tidy_permanova[2, 6], 3)`), and the interaction
between treatment and passage explained 
`r myround(100*tidy_permanova[3, 4])`\% of the variation (F~`r tidy_permanova[3, 2]`,`r tidy_permanova[4, 2]`~ = `r myround(tidy_permanova[3, 5], 1)`, p = `r myround(tidy_permanova[3, 6], 3)`). There was 
no difference in dispersion between treatments or passages (F~3,44~ = 0.91, p = 0.45). 
Finally, CH~4~ oxidation rate was correlated with Bray-Curtis dissimilarity across both treatments in Passage 5 and explained `r myround(dbrda_percent_explained[1], 1)`\% of the variation in Bray-Curtis dissimilarity (dbRDA: F~`r dbrda_aov[1, 1]`,`r dbrda_aov[2, 1]`~ = `r myround(dbrda_aov[1, 3], 2)`, p = `r myround(dbrda_aov[1, 4], 3)`)

## Taxa that responded to selection

To identify taxa that responded
to selection on soil CH~4~ oxidation rate, we tested the differential relative
abundance of families in the Positive jars relative to the Neutral jars within
Passage 5 using three methods and then plotted the taxa identified by all three methods. We identified `r length(da_fams)`
families that were enriched or depleted in the Positive treatment relative to
the Neutral treatment (`r figures("da", display = "cite")`).

Overall, none of the families enriched in the Positive selection treatment
contain known methanotrophs. Several taxa identified had a 
higher taxonomic designation that contains methanotrophs, for example, the
Gammaproteobacteria class had a large effect size.
The Gammaproteobacteria include the type I and type X methanotrophs in the families *Methylococcaceae* and *Methylothermaceae* [@stein2012].
However, the Gammaproteobacteria is among the most diverse 
groups in the Prokaryotes, so this is not strong evidence for a selection response
by methanotrophs [@garrity2005].
In addition, the *Puniceicoccaceae* is a member of the phylum Verrucomicrobia. The Verrucomicrobia is a diverse group that contain
known methanotrophs as well as ammonia-oxidizing bacteria [@freitag2003]. Other than these two groups, none of the other taxa enriched in the Positive treatment are known to be related to methanotrophs. Two groups in the Armatimonadales were enriched in the Positive treatment including the family *Fimbriimonadaceae* and an unclassified ASV from the order Armatimonadales [@im2012].
*Cytophagaceae* was also enriched in
the Positive treatment and contains a number of mainly aerobic heterotrophs
that can digest a variety of macromolecules [@mcbride2014]. 
The remaining families include the uncultured family 0319-6G20, *Diplorickettsiaceae*, *Rhodospirillaceae*, and an unclassified Kapabacteriales.

We did not identify any methanotrophic families in the overall differential abundance analysis. 
However, we wanted to look more closely at the known methanotrophs in our dataset to be sure that they did not have an effect.
To do this, we subset the ASVs in our dataset that were in families that contained methanotrophs.
Only two families were represented: *Methylacidiphilaceae* and *Beijerinckiaceae*. 
Aggregating reads at the family level, neither family was differentially abundant between the two treatments.
However, aggregated at the genus level, a group of unclassified genera in the *Beijerinckiaceae* were depleted in the Positive treatment and the genus
*Rhodoblastus*, a member of the *Beijerinckiaceae*, was enriched in the
Positive treatment.  While many *Beijerinckiaceae* are methanotrophs, several
taxa in this family have lost the ability to oxidize CH~4~ and it appears
that *Rhodoblastus* species are not able to grow on CH~4~, though they can grow
on methanol [@tamas2014; @dedysh2016]. Based on this analysis,
it appears that no methanotrophs were enriched in the Positive treatment.

# Discussion

We used artificial ecosystem selection to estimate the contribution of variation
in microbiome composition to variation in the rate of an ecosystem function,
methane oxidation in soil, independent of environmental variation. Understanding
how and to what degree microbiome variation contributes to variation in
ecosystem function is important for many reasons. For example, successful
microbiome manipulations require that the manipulated microbiome contribute to
variation in ecosystem function independent of other drivers of ecosystem
variation (such as variation in environmental conditions). This is because the
drivers of variation in ecosystem functions can interact in complicated ways
(Figure 5);
for example, environmental variation can directly contribute by providing
conditions that select for microbial groups that in turn alter the rates of
ecosystem functions. Determining the independent contribution of microbiome
variation to ecosystem function is crucial because if microbiome composition is
driven primarily by environmental conditions, then introducing a desirable taxon
through microbiome manipulation without altering the environment will likely be
unsuccessful at shifting the targeted ecosystem function. 
The artificial selection approach is different from the comparative and
manipulative approaches used in past attempts at answering this question,
because it can control for both the direct effect of environment on function as
well as the indirect effect on environment via its impact on microbiome
assembly.

In our study, we observed an increase in methane oxidation rate in the
selection treatment relative to the control, which demonstrates that there was a
response to selection. Given that we observed a response to selection, we
conclude that variation in the microbiome contributes to variation in the
methane oxidation rate independent of the environment. This suggests that
microbiome manipulations could be an effective approach for altering the rate of
methane oxidation in this soil, and that the artificial selection approach may
be useful in determining the potential for microbiome manipulations for other
functions in other ecosystems.

Given that variation in the microbiome is associated with variation in the rate
of an ecosystem function in our system, a reasonable follow-up question is "how
much variation in ecosystem function is associated with microbiome variation in
this system?" One way to estimate this is to determine how much the recipient
jars resemble the selected donor jars that were used to inoculate them
[@falconer1996]. We can
calculate the response to selection as the difference between two successive
passages in their mean methane oxidation rate. We will denote this as $R$. We
can also calculate the strength of selection as the difference in mean methane
oxidation rate between the twelve jars in one generation and the three jars chosen
for selection in that generation, which we will call the selection differential
and denote as $S$. If we plot the cumulative $R$ against the cumulative $S$, the
slope of this relationship will equal the proportion of variation explained by
the microbiome. If the change in mean function from passage one to passage two
($R$) is equal to the difference in mean function between the twelve jars in
passage one and the three jars selected to inoculate passage two ($S$), then we
would conclude that 100% of the variation is due to variation in the microbiome.
Likewise, if recipients do not resemble the donors in their mean methane
oxidation rate and simply wander randomly, then we would conclude that all of
the variation is due to the environment or technical variation.

This relationship between microbiome variation and ecosystem function variation
is analogous to the concept of "heritability" [@lynch1998] used by quantitative
geneticists, or more precisely the concept of "microbiability" [@difford2016] proposed
by microbiome scientists who study host-associated microbiomes. Although rarely
used in the study of environmental microbiomes, this concept could be very
useful for understanding and manipulating microbially-mediated functions in a
variety of ecosystems. In our experiment, variation in microbiome taxonomic
composition statistically explained (i.e., was associated with) 40% of the
variation we observed in the rate of methane oxidation.  This suggests that
there is substantial potential for altering this ecosystem function through
microbiome manipulation in this soil. It is very likely that the
“environmental microbiability” will be different for other ecosystem functions
in this soil and for methane oxidation in other soils. However, our experiment
demonstrates that this relationship is measurable and provides an example of how
this can be accomplished.

Given that variation in the microbiome is driving substantial variation in
methane oxidation rate in our system, we next wanted to determine which aspects of the
microbiome best explain this contribution of microbiome variation to
methane oxidation rate. There are three inter-related ways that microbiomes could have responded
to selection in this experiment: gain or loss of taxa, changes in the
relative abundances of taxa, or changes within the genomes of the constituent
taxa. We surveyed microbiome variation via 16s rRNA ribotyping in our
experiment, which allowed us to deeply sample taxonomic diversity but did not
allow us to directly address whether taxa in this experiment evolved genomic changes
as a result of selection. However, if such genomic changes resulted in increased
persistence or abundance of the population with these changes, this would be
detectable. Therefore, we will focus on the first
two possibilities.

Richness at the ASV level did not vary between the two treatments and there
were relatively few taxa gained or lost in the Positive selection treatment and
none of these were prevalent across the 12 jars in passage 5. Therefore, the
gain or loss of species is unlikely to explain the increase in methane
oxidation rate. However, we found that Bray-Curtis dissimilarity was greater
between the two treatments in Passage 5 than within each treatment and was
correlated with methane oxidation rate, which suggests that changes in the
relative abundance of taxa could explain the response to selection.

Even though we observed an increase in methane oxidation rate in the Positive
treatment and a difference in composition between the two treatments, we did
not observe an increase in the relative abundance of methanotrophs. This was
surprising given that methane consumption is not a common trait among microbes
and that it is often assumed that the rate of an ecosystem function is limited
by the final enzymatic step in the underlying metabolic pathway [@rocca2015]. In certain
ecosystems, methane production and consumption are correlated with the
abundance of methanogens and methanotrophs as estimated from marker genes
[@meyer2020; @meyer2020a]. However, our results suggest that
ecosystem-scale methane oxidation rates can be altered by non-methanotrophs,
perhaps through ecological interactions with methanotrophic species. This
suggests that simple assumptions about how microbes contribute to rate variation
in ecosystem function may not apply universally, and it demonstrates the
importance of using biologically "agnostic" approaches (that make few starting
assumptions) to linking microbial taxa to ecosystem functions [@morris2020].
Artificial ecosystem selection is an important example of such an approach.

There is increasing interest in using artificial selection for understanding and
manipulating the microbiomes associated with plants and animals (aka “microbiome
breeding”; @mueller2022). Our study demonstrates that artificial ecosystem
selection can also be an important tool for exploring the relationship between
microbiome composition and ecosystem function in non-host systems. This
approach can provide unique information about the independent contribution of
microbiomes to ecosystem functions. Such information is crucial if we are to
successfully manipulate environmental microbiomes to alter ecosystem functions,
whether to improve crop productivity [@chaparro2012]  or
ameliorate the impacts of environmental change [@jansson2020].

# Acknowledgments

This project was supported by the National Science Foundation Graduate Research
Fellowship Program (grant no. DGE 1255832) and the ARCS Foundation Florence and
Mike Nudelman Scholarship.

# Competing Interests

We declare we have no competing interests.

# Data Availability Statement

The 16S rRNA sequencing data generated during the current study are available
in the NCBI Sequence Read Archive (SRA) under BioProject accession number
PRJNA832314, https://www.ncbi.nlm.nih.gov/sra/PRJNA832314. The metadata
generated during the current study as well as the scripts to recreate the
analysis are available on Github,
https://github.com/amorris28/artificial_ecosystem_selection.

# Supplement

# References

::: {#refs}
:::

# Figures

```{r response, fig.width=4, fig.height=3}

p_response

```

`r figures("response", "Response to selection on soil CH~4~ oxidation rate. The y-axis is the log~10~ of the additive inverse of the first-order exponential decay constant $k$ (i.e., $-k$) with units day^-1^ so that a more positive value represents a higher CH~4~ oxidation rate.")`

```{r herit, fig.width=7}
# 
herit_plot
#`r figures("herit", paste0("Ecosystem heritability as the regression of mid-recipient methane flux on mid-donor methane flux ($-k$ with units day^-1^). Donor is the mean oxidation rate of the three jars selected to inoculate the next passage. Recipient is the mean oxidation rate of the 12 jars in the next passage. The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment. Slope of the regression line estimates the proportion of variation in methane oxidation rate due to variation in the microbiome. For the Positive treatment, the variance explained by the microbiome was ", myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100, 1), "% (slope = ", myround(coefficients(herit_model)[2] + coefficients(herit_model)[4], 3), ", SE = ", myround(tidy(herit_model)$std.error[4], 2), ") and for the Neutral treatment the variance explained by the microbiome was ", myround(coefficients(herit_model)[2]*100, 1), "% (slope = ", myround(coefficients(herit_model)[2], 3), ", SE = ", myround(tidy(herit_model)$std.error[2], 2), ")."))`
```

`r figures("herit", paste0("Divergence between the positive selection line and the control line over five passages. Values are CH~4~ oxidation rate as the additive inverse of the first-order exponential decay constant k (day^-1^). Regression of divergence on cumulative selection differential provides an estimate of realized heritability (h^2^ &pm; SE) and the slope of the regression was ", myround(tidy(rh_model)[[2, 2]], 2), " &pm; ", myround(tidy(rh_model)[[2, 3]], 2), "."))`

```{r beta, results=FALSE, fig.height=2.5}

p_beta

```

`r figures("beta", "Non-metric multidimensional scaling plot of beta diversity for all jars. Dissimilarities are based on rarefied Bray-Curtis dissimilarity. Colors represent treatment and shapes represent passage number.")`

```{r da, results=FALSE}

p_log10

```

`r figures("da", "Differentially abundant family-level taxa identified by ANCOM-II, ALDEx2, and CORNCOB. Values on the x-axis are relative abundances on a log~10~ scale. Large points with error bars are the median relative abundance and interquartile range. Taxa in the top panel are enriched in the Positive treatment and taxa in the bottom panel are depleted in the Positive treatment. Taxa are sorted by their effect size with taxa at the top having the largest positive effect size and taxa at the bottom with the largest negative effect size.")`

![Figure 5. Variation in the rate of an ecosystem function is potentially the result of at least three components: variation in the abiotic environmental conditions, variation in microbiome composition, and the covariance between microbiomes and the environment. It is difficult with traditional study designs to isolate the direct effect of the microbiome from the effect of the environment via covariance with the microbiome. Here, we attempt to isolate the effect of the microbiome through artificial selection on microbiome composition. For simplicity, we omitted the reverse arrows as well as the interaction between microbiome and the environment, though these relationshps may also exist.](Figure1.png)