### 2023-02-27 analysis/analysis.Rmd
update figures and analysis scripts

---
title: "Analysis"
author: "Andrew Morris"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
here::i_am("analysis/analysis.Rmd")

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = TRUE)

```

```{r initialize_packages, message=FALSE}
library(morris)
library(broom)
library(knitr)
library(broman)
library(phyloseq)
library(knitr)
library(magrittr)
library(vegan)
library(ggpubr)
library(ggtext)
library(ALDEx2)
library(corncob)
library(ANCOMBC)
library(tidyverse)
library(here)
library(patchwork)
library(extrafont)

source(here('R/', 'functions.R'))

set.seed(1337)

theme_set(theme_classic(base_size = 7) +
            theme(panel.border = element_rect(fill = NA, linewidth = 1.0625),
                  axis.line = element_blank(),
                  text = element_text(family = 'Arial')))
```

```{r import_data, message=FALSE}
fluxes <- read_tsv(here(der_dir, 'fluxes.tsv'))
heritability <- read_tsv(here(der_dir, 'heritability.tsv'))
aes_data <- read_tsv(here(der_dir, 'community_data.tsv'))
```

# Clean data

```{r}
# Shift the passage numbers so the first passage becomes "Passage 0"
fluxes <-
  fluxes %>% 
  mutate(passage = passage - 1)
```

```{r fix-treatment}

aes_data <-
  aes_data %>% 
  mutate(passage_treat = treat) %>% 
  separate(passage_treat, c('passage', "treat"), sep = 2, remove = FALSE) %>% 
  select(sample, passage_treat, passage, treat, selected, ch4, asv, count, Kingdom:Species, seq)

```

```{r}
aes_data %>% 
    select(sample, count, Phylum, Family) %>% 
    group_by(sample) %>% 
    mutate(sum_count = sum(count)) %>% 
    mutate(ra = count / sum_count) %>% 
    filter(is.na(Family)) %>%
    # filter(Family == "") %>%
    summarize(sum = sum(ra))

```

```{r fix-taxonomy}

# There are "Unknown Family" taxa with Genus and Class, but Order is Incertae Sedis.
# I'm just going to remove everything below Class and call it "Unclassified"

fill_taxonomy <- function(tab) {
    if(is.na(tab[[1]])) tab[[1]] = "Unclassified"
    for(i in 2:length(tab)) {
      if(is.na(tab[[i]])){
        if(grepl("unclassified", tab[[i-1]])) tab[[i]] = tab[[i-1]]
        else tab[[i]] = paste(tab[[i-1]], "unclassified", sep = "_")
      }
    }
    tab
}

aes_data <-
  aes_data %>% 
  nest(data = -c(asv, Kingdom:Species)) %>% 
  nest(taxa = Kingdom:Species) %>% 
  mutate(taxa = map(taxa, ~fill_taxonomy(.x))) %>% 
  unnest(taxa) %>% 
  unnest(data)

```

```{r remove_chloroplasts_mitochondria}

chloroplast_seqs <- 
  aes_data %>% 
  filter(Order == 'Chloroplast') %>% 
  select(asv, seq) %>% 
  distinct() %$%
  seq
# 10 ASVs were identified as chloroplast in their taxonomy.
# seq matches complete genomes of chloroplasts on NCBI so I will remove them.

aes_data <- 
  aes_data %>% 
  filter(Order != 'Chloroplast')

mitochondria_seqs <-
  aes_data %>% 
  filter(Family == 'Mitochondria') %>%
  select(asv, seq) %>% 
  distinct() %$%
  seq

# 99 ASVs were identified as Mitochondria

aes_data <- 
  aes_data %>% 
  filter(Family != 'Mitochondria')

```

# Overview

The final dataset includes `r length(unique(fluxes$treat))` treatments across `r length(unique(fluxes$passage))` passages (generations). This includes a total of n = `r nrow(fluxes)` samples. The samples are broken down across passages and treatments as in the following table:

```{r}
fluxes %>% 
  group_by(flux_date, treat) %>% 
  summarize(n = n_distinct(rep), .groups='drop') %>% 
  kable()
```

# Reponse to Selection

This script takes in the flux data. It then plots the raw data
and the log-transformed data. Finally, it fits a linear model to the flux data
over time per treatment and tests a difference of slope between the positive
and the neutral treatment.

## Plot the raw data

```{r plot_raw}

ggplot(fluxes, aes(x = passage, y = ch4, color = treat)) +
    stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
    scale_color_manual(
        name = "Treatment", 
        labels = c('Control', 'Selection'), 
        values = c('gray40', 'darkorange2')) +
    geom_jitter(width = 0.1) + 
    labs(x = "Passage Number", 
         y = "Methane Oxidation Rate (-k)") +
  scale_x_continuous(labels = 1:5)
```

## Plot the log-transformed data

```{r plot_log10}

ratio = calc_plot_ratios(fluxes$passage, fluxes$log_ch4)

# (p_response_cl <- ggplot(fluxes, aes(x = passage, y = log_ch4, color = treat)) +
#     stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
#     scale_color_manual(
#       name = "Treatment", 
#       labels = c('Control', 'Selection'), 
#       values = c('gray40', 'darkorange2')) +
#     geom_jitter(#aes(alpha = selected), 
#       width = 0.1) + 
#         labs(x = "Passage Number", 
#                   y = expression(log[10] ~ CH[4] ~ "(-k)")) +
#   scale_x_continuous(labels = 1:5) +
#     coord_fixed(ratio))

# save(p_response, file = here(der_dir, 'response_plot.RData'))

# ggsave(here(fig_dir, 'fig2.jpg'), p_response, width = 4, height = 3)
# ggsave(here(fig_dir, 'fig2.svg'), p_response, width = 4, height = 3)
# ggsave(here(fig_dir, 'fig2_cl.png'), p_response_cl, width = 4, height = 3)

(p_response_bw <- ggplot(fluxes, aes(x = passage, y = log_ch4, fill = treat, linetype = treat)) +
    stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x', color = 'black',
                linewidth = 0.5) +
    scale_fill_manual(
      name = "Treatment", 
      labels = c('Control', 'Selection'), 
      values = c('white', 'black')) +
    scale_linetype_manual(
      name = "Treatment", 
      labels = c('Control', 'Selection'), 
      values = c(3, 1)) +
        geom_jitter(shape = 21, 
      width = 0.1,
      size = 1) + 
        labs(x = "Passage Number", 
                  y = expression(log[10] ~ CH[4] ~ "(-k)")) +
  scale_x_continuous(labels = 1:5) +
    coord_fixed(ratio)
    )

ggsave(here(fig_dir, 'fig2.eps'), p_response_bw, width = 88, height = 65, units = 'mm', device = cairo_ps)
# embed_fonts(here(fig_dir, 'fig2_bw.eps'))

# (p_response_gs <- ggplot(fluxes, aes(x = passage, y = log_ch4, color = treat)) +
#     stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x', linewidth = 0.5) +
#     scale_color_manual(
#       name = "Treatment", 
#       labels = c('Control', 'Selection'), 
#       values = c('gray50', 'black')) +
#     geom_jitter(#aes(alpha = selected), 
#       width = 0.1,
#       size = 0.5) + 
#         labs(x = "Passage Number", 
#                   y = expression(log[10] ~ CH[4] ~ "(-k)")) +
#   scale_x_continuous(labels = 1:5) +
#     coord_fixed(ratio)
# )
# ggsave(here(fig_dir, 'fig2_gs.png'), p_response_gs, width = 88, height = 65, units = 'mm')

```

## Test a difference of slopes

```{r raw_model}
raw_model <- lm(ch4 ~ passage * treat, data = fluxes)

ggplot(raw_model, aes(x = .fitted, y = .resid)) + geom_point()

```

Since the residuals are not normally distributed, we will use the log-transformed data.

```{r response_model}

response_model <- lm(log_ch4 ~ passage * treat, data = fluxes)

ggplot(response_model, aes(x = .fitted, y = .resid)) + geom_point()

# Percent increase in CH4 per passage
# Back calculate the slope of the positive treatment

pos_slope <- 
    response_model$coef[["passage:treatp"]]

per_slope <-
    (10^pos_slope - 1) * 100

```

```{r no_treat_model}
# No treatment model
no_treat_model <- lm(log_ch4 ~ passage, data = fluxes)

```

Here I fit two models. The first is the full model which includes passage number, treatment, and the interaction between the two. The second model includes only passage with no effect of treatment. Comparing the models using `anova`

```{r compare_models}
# Compare interaction to no interaction
compare_response_mods <- anova(no_treat_model, response_model)
compare_response_mods <- tidy(compare_response_mods)
kable(compare_response_mods)
```

Here is the full interaction model:

```{r print_model}
kable(tidy(response_model))
```

```{r save_model}
# save(response_model, compare_response_mods, per_slope, file = here(der_dir, 'response_model.RData'))
```

The intercept for the Neutral treatment was significantly different from zero
at `r myround(tidy(response_model)[[1, 2]], 2)` 
(SE = `r myround(tidy(response_model)[[1, 3]], 2)`, 
t = `r myround(tidy(response_model)[[1, 4]], 2)`, 
p = `r myround(tidy(response_model)[[1, 5]], 2)`). 
The Positive
treatment started at a significantly lower methane oxidation rate than the
Neutral treatment with a y-intercept of 
`r myround(tidy(response_model)[[3, 2]], 2)` relative to Neutral 
(SE = `r myround(tidy(response_model)[[3, 3]], 2)`, 
t = `r myround(tidy(response_model)[[3, 4]], 2)`, 
p = `r myround(tidy(response_model)[[3, 5]], 2)`). 
There was no change in methane oxidation rate of the Neutral
treatment over the five passages 
(slope = `r myround(tidy(response_model)[[2, 2]], 2)`, 
SE = `r myround(tidy(response_model)[[2, 3]], 2)`, 
t = `r myround(tidy(response_model)[[2, 4]], 2)`, 
p = `r myround(tidy(response_model)[[2, 5]], 2)`). 
By contrast, there was a significant increase in the Positive
treatment of `r myround(tidy(response_model)[[4, 2]], 2)` per passage which is
approximately a `r round((pos_slope - 1) * 100, 0)`% increase in methane
oxidation rate per passage 
(SE = `r myround(tidy(response_model)[[4, 3]], 2)`, 
t = `r myround(tidy(response_model)[[4, 4]], 2)`, 
p = `r myround(tidy(response_model)[[4, 5]], 2)`).

# Heritability

This script imports the heritability data. This dataset includes the mean flux
within treatment by generation for 1) all parents, 2) selected parents, and 3)
offspring. Then, it plots offspring on the selected parents with the two
treatments combined and then separately for the two treatments.

## Test heritability as divergence on cumulative selection differential

```{r heritability}

divergence <-
    fluxes %>% 
    select(passage, treat, log_ch4) %>% 
    group_by(passage, treat) %>% 
    summarize(mean = mean(log_ch4), .groups = "drop") %>% 
    pivot_wider(names_from = "treat", values_from = "mean") %>% 
    mutate(divergence = p - n, .keep = 'unused')

# Selection Differential
sd <- 
    heritability %>% 
    mutate(sd = selected - parental) %>% 
    select(passage, treat, sd) %>% 
    pivot_wider(names_from = 'treat', values_from = 'sd') %>% 
    mutate(sd_sum = p + n, .keep = 'unused')

sd %>% 
    arrange(passage) %>% 
    mutate(sd_csum = cumsum(sd_sum)) ->
    sd_csum

# Realized Heritability
rh <-
    divergence %>% 
    left_join(sd_csum, by = "passage") %>% 
    filter(passage != 0 )
```

```{r}

ratio_div <- calc_plot_ratios(divergence$passage, divergence$divergence)
(
p_div <-
ggplot(divergence, 
       aes(x = passage, 
           y = divergence)) +
    # stat_smooth(method = 'lm', 
    #             se = FALSE, 
    #             formula = 'y ~ x',
    #             color = "black",
    #             linewidth = 0.5) +
        geom_line() +
    geom_point() + 
    labs(x = expression(atop("Passage Number", " ")), 
         y = expression(atop("Divergence", log[10] ~ CH[4] ~ "(-k)"))) +
    coord_fixed(ratio_div) +
    scale_x_continuous(labels = 1:5) + 
    scale_y_continuous(n.breaks = 4) +
    theme(plot.margin = margin(0, 0.25, 0, 0.5, "cm"))
)
ratio_rh <- calc_plot_ratios(rh$sd_csum, rh$divergence)

(
p_rh <-
rh %>% 
ggplot(aes(x = sd_csum, 
           y = divergence)) +
    stat_smooth(method = 'lm',
                se = F,
                formula = 'y ~ x',
                color = "black",
                linewidth = 0.5) +
    geom_point(size = 1) + 
    labs(x = expression(atop("Cumulative Selection Differential", log[10] ~ CH[4] ~ "(-k)")), 
         y = expression(atop("Divergence", log[10] ~ CH[4] ~ "(-k)"))) +
    scale_y_continuous(n.breaks = 4) +
    # theme(plot.margin = margin(0, 0.5, 0, 0.25, "cm")) +
    coord_fixed()
)

remove_y <- theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
)
p <- list(p_div, p_rh)
herit_plot <- wrap_plots(p, nrow = 1) + plot_layout(guides = "collect")

rh_model<-lm(divergence~sd_csum,data = rh)
summary(rh_model)

# save(herit_plot, file = here(der_dir, "herit_plot.RData"))

# ggsave(here(fig_dir, "fig3.png"), width = 8*.9, height = 4*.9)
# ggsave(here(fig_dir, "fig3.svg"), width = 8*.9, height = 4*.9)
ggsave(here(fig_dir, "fig3.eps"),
       p_rh,
       width = 88,
       height = 45, 
       units = 'mm',
       device = cairo_ps)
# ggsave(here(fig_dir, "fig3.eps"), 
#        p_rh,
#        width = 5,
#        height = 2.5)
# save(rh_model, file = here(der_dir, 'herit_model.RData'))

```

# Community Analysis

## Exploratory

```{r n-seqs}

aes_data %>% 
  group_by(sample) %>% 
  summarize(total = sum(count)) %>% 
  ggplot(aes(x = total)) +
  geom_histogram() +
  theme_classic() +
  labs(title = "Distribution sequencing depths",
       x = "Library size",
       y = "Frequency")

seq_range <-
  aes_data %>% 
  group_by(sample) %>% 
  summarize(total = sum(count)) %>% 
  summarize(seq_range = range(total)) %$% 
  seq_range

min_seqs <- seq_range[1]

a_analysis <- 
  aes_data %>%
  count(sample, count) %>%
  nest(data = -sample) %>%
  mutate(n_seqs = map_dbl(data, ~sum(.x$count * .x$n)),
         sobs = map_dbl(data, ~sum(.x$n)),
         rare = map_dbl(data, ~rarefy(rep(.x$count, .x$n), sample=seq_range[1]))) %>%
  select(-data)

a_analysis %>%
  select(sample, n_seqs, sobs, rare) %>%
  mutate(treat = substr(sample, 1, 3)) %>% 
  select(-sample) %>% 
  group_by(treat) %>% 
  pivot_longer(-c(treat)) %>%
  ggplot(aes(x=treat, y=value)) +
  facet_wrap(~ name, nrow = 3, scales = "free") +
  geom_boxplot() +
  geom_jitter() +
  theme_classic() +
  labs(title = "AES", x = "Treatment", y = "Value")

```

This is the number of reads (n_seqs), the rarefied richness (rare), and 
the observed richness (sobs for "species observed") for the four treatments:
generations 2 and 5 and selection treatments neutral and positive. As you can
see, there were more sequences from G2 than G5 and also greater richness
in G2 than G5. However, that richness difference is robust to rarefying
to the level of the shallowest sample. There doesn't seem to be a systemic
difference in sequencing depth or richness between the neutral and positive 
treatments in either generation.

### Richness

#### Plot

```{r alpha-diversity-plot}

# aes_data %>% 
#   group_by(sample) %>% 
#   summarize(n_seqs = sum(count)) %>% 
#   summarize(min = min(n_seqs))

aes_df <-
  aes_data %>% 
  select(sample, asv, count) %>%
  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
  as.data.frame()

rownames(aes_df) <- aes_df$sample
aes_df <- aes_df[, -1]

rare <- 
  rarefy(aes_df, seq_range[1]) %>% 
  as_tibble(rownames = "sample") %>% 
  rename(rare = value)

aes_data %>% 
  inner_join(rare, by = 'sample') %>% 
  select(sample, passage_treat, rare) %>% 
  distinct %>% 
  ggplot(aes(x = passage_treat, y = rare)) +
  geom_boxplot() +
  geom_jitter() +
  labs(x = "Passage x Treatment", y = "Rarefied Richness") +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())

```

```{r compare_richness_methods}

aes_df_long <-
  aes_df %>% 
  as_tibble(rownames = "sample") %>% 
  pivot_longer(starts_with("ASV")) %>%     
  group_by(sample) %>% 
  uncount(value)

subsample <- function(data, sample_size) {
  data %>% 
    sample_n(sample_size) %>% 
    summarize(s = n_distinct(name))
}

empirical <- map_dfr(1:10, ~subsample(aes_df_long, seq_range[1]), .id = "iters")
empirical <-
empirical %>% 
  group_by(sample) %>% 
  summarize(s = mean(s))

inner_join(rare, empirical, by = "sample") %>% 
  ggplot(aes(x = rare, y = s)) +
  geom_abline(intercept = 0, slope = 1, color = "gray") +
  geom_point()
```

This plot shows the correlation between Hurlbert's (1971) approach to
richness estimation (rare) and empirically rarefying 10 times and taking
the mean richness (s).

#### Hypothesis Test

```{r alpha-diversity-test}

rare_data <-
  aes_data %>% 
  select(sample, passage_treat, passage, treat, ch4) %>% 
  distinct %>% 
  inner_join(rare, by = "sample")

kt <- kruskal.test(rare ~ passage_treat, data = rare_data)
kt <- tidy(kt)

wt <- pairwise.wilcox.test(rare_data$rare, 
                           rare_data$passage_treat, 
                           p.adjust.method = "BH")
wt <- tidy(wt)

rare_median <-
  rare_data %>% 
  group_by(passage) %>% 
  summarize(median = median(rare),
            IQR = quart(rare)[2] - quart(rare)[1])

```

There is an effect of passage x treatment on richness (Kruskal-Wallace test: chi-squared = 
`r kt[1,1]`, df = `r kt[1,3]`, p = `r kt[1, 2]`). The pairwise test shows that
G2 is different from G5, but N and P are not different in either generation
(Wilcoxon/Mann-Whitney Test).

```{r}

kable(wt)

```

```{r alpha-diversity-ch4-test}

s <- cor.test(~ ch4 + rare, data = rare_data, method = "spearman", alternative = "two.sided") %>% 
  tidy

s5 <-
  rare_data %>% 
  filter(passage == "G5") %>% 
  cor.test(~ ch4 + rare, data = ., method = "spearman", alternative = "two.sided") %>%
  tidy

min_seqs <- paste(substr(as.character(min_seqs), 1, 3), substr(as.character(min_seqs), 4, 6), sep = ",")
# save(kt, wt, rare_median, s5, min_seqs, file = here(der_dir, "richness_models.RData"))

```

A Spearman's rank correlation test shows no correlation between
methane oxidation rate and richness (Spearman's rho = `r s[1,1]`, p = `r s[1,3]`).

## Beta Diversity

```{r bray-curtis-nmds, messages=FALSE}

dist <- vegdist(aes_df, method="bray")

if(file.exists(here(der_dir, "rare_dist_bray.rds"))) {
  rare_dist_bray <- 
    readRDS(here(der_dir, "rare_dist_bray.rds"))
} else {
  rare_dist_bray <- avgdist(aes_df, dmethod="bray", sample=seq_range[1])
  rare_dist_bray %>% 
    saveRDS(here(der_dir, "rare_dist_bray.rds"))
}

nmds <- metaMDS(rare_dist_bray)

nmds_scores <-
  scores(nmds, display = "sites") %>% 
  as_tibble(rownames = "sample")

nmds_data <-
  aes_data %>% 
  select(sample, passage, treat) %>% 
  distinct %>% 
  inner_join(nmds_scores, by = "sample")

centroids <-
  nmds_data %>% 
  group_by(passage, treat) %>% 
  summarize(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2),
            .groups = "drop")
# p_beta_cl <-
#   nmds_data %>%   
#   ggplot(aes(x = NMDS1, y = NMDS2, shape = passage, color = treat)) +
#   geom_point() +
#   stat_ellipse(show.legend = FALSE) +
#   coord_fixed() +
#   scale_shape_discrete(name = "Passage", labels = c(2, 5)) +
#   scale_color_manual(name = "Treatment",
#                        labels = c('Control', 'Selection'),
#                        values = c('gray40', 'darkorange2')) #+
  # theme_bw() +
  # theme(
  #   plot.background = element_blank(),
  #   panel.grid.major = element_blank(),
  #   panel.grid.minor = element_blank())
  #   
# pcoa <- cmdscale(rare_dist_bray, eig = TRUE, add = TRUE)
# round(pcoa$eig*100/sum(pcoa$eig),1)
# pcoa_per_exp <- 100 * eigenvals(pcoa) / sum(eigenvals(pcoa))
# 
# ratio = pcoa_per_exp[2]/pcoa_per_exp[1]
# 
# pcoa_scores <-
#   scores(pcoa, display = "sites") %>%
#   as_tibble(rownames = "sample")
# 
# pcoa_data <-
#   aes_data %>%
#   select(sample, passage, treat) %>%
#   distinct %>%
#   inner_join(pcoa_scores, by = "sample")
# 
# # p_beta <-
#   pcoa_data %>%
#   ggplot(aes(x = Dim1, y = Dim2, shape = passage, color = treat)) +
#   geom_point() +
#   stat_ellipse(show.legend = FALSE) +
#   coord_fixed(ratio) +
#   scale_shape_discrete(name = "Passage", labels = c(2, 5)) +
#   scale_color_manual(name = "Treatment",
#                        labels = c('Neutral', 'Positive'),
#                        values = c('gray40', 'darkorange2')) #+
  # theme_bw() +
  # theme(
  #   plot.background = element_blank(),
  #   panel.grid.major = element_blank(),
  #   panel.grid.minor = element_blank())

# save(p_beta, file = here(der_dir, "beta_fig.RData"))

# ggsave(here(fig_dir, 'fig4.png'), p_beta, height = 2.5, width = 4) 
# ggsave(here(fig_dir, 'fig4.svg'), p_beta, height = 2.5, width = 4) 
# ggsave(here(fig_dir, 'fig4_cl.png'), p_beta_cl, height = 2.5, width = 4)

(p_beta_bw <-
  nmds_data %>%   
  ggplot(aes(x = NMDS1, y = NMDS2, shape = passage, fill = treat, 
             # linetype = treat
             )) +
  geom_point() +
  # stat_ellipse(
  #     # show.legend = FALSE
  #     ) +
  #   scale_linetype_manual(
  #     name = "Treatment", 
  #     labels = c('Control', 'Selection'), 
  #     values = c(3, 1)) +
  coord_fixed() +
  scale_shape_manual(name = "Passage", 
                     labels = c(2, 5),
                     values = c(24, 21)) +
  scale_fill_manual(name = "Treatment",
                       labels = c('Control', 'Selection'),
                       values = c('white', 'black')) +
  guides(fill=guide_legend(override.aes = list(shape=21)))
  # scale_color_manual(name = "Treatment",
  #                      labels = c('Neutral', 'Positive'),
  #                      values = c('white', 'black')) #+
  # theme_bw() +
  # theme(
  #   plot.background = element_blank(),
  #   panel.grid.major = element_blank(),
  #   panel.grid.minor = element_blank())
  #   
)
ggsave(here(fig_dir, 'fig4.eps'), 
       p_beta_bw, 
       height = 60,
       width = 180,
       # height = 60,
       # width = 88,
       units = 'mm', 
       device = cairo_ps
       )

# (p_beta_gs <-
#   nmds_data %>%   
#   ggplot(aes(x = NMDS1, y = NMDS2, shape = passage, color = treat)) +
#   geom_point() +
#   # stat_ellipse(
#   #     # show.legend = FALSE
#   #     ) +
#   coord_fixed() +
#   scale_shape_manual(name = "Passage",
#                      labels = c(2, 5),
#                      values = c(17, 19)) +
#   scale_color_manual(name = "Treatment",
#                        labels = c('Control', 'Selection'),
#                        values = c('gray50', 'black')) +
#   guides(fill=guide_legend(override.aes = list(shape=21)))
# )
# ggsave(here(fig_dir, 'fig4_gs.png'), p_beta_gs, height = 2.5*1.5, width = 4*1.5)

```

```{r}
dist_df <-
  rare_dist_bray %>% 
  as.matrix %>% 
  as_tibble(rownames = "sample")

meta_dist <-
  aes_data %>% 
  select(sample, passage_treat, passage, treat, ch4) %>% 
  distinct %>% 
  inner_join(dist_df, by = "sample")

permanova <- adonis2(rare_dist_bray ~ passage * treat, data = meta_dist, permutations = 999)

betadisper(rare_dist_bray, meta_dist$passage_treat) %>% 
  anova()

tidy_permanova <- 
  tidy(permanova)

# save(tidy_permanova, file = here(der_dir, "beta_model.RData"))

```

There is a significant difference of centroid between passage, treatment, and
their interaction. Testing a difference of dispersion using betadisper
shows no difference in variance between the four groups.

### dbrda methane in passage 5

```{r dbrda, warnings=FALSE}
passage5_df <-
  aes_data %>%  
  filter(passage == "G5") %>% 
  group_by(asv) %>% 
  mutate(total = sum(count)) %>% 
  filter(total != 0) %>% 
  ungroup() %>% 
  select(-total) %>% 
  select(sample, asv, count) %>%
  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
  as.data.frame()

rownames(passage5_df) <- passage5_df$sample
passage5_df <- passage5_df[, -1]

if(file.exists(here(der_dir, "rare_dist_bray_p5.rds"))) {
  rare_dist_bray_p5 <- 
    readRDS(here(der_dir, "rare_dist_bray_p5.rds"))
} else {
rare_dist_bray_p5 <- avgdist(passage5_df, dmethod="bray", sample=seq_range[1])
  rare_dist_bray_p5 %>% 
    saveRDS(here(der_dir, "rare_dist_bray_p5.rds"))
}

rare_dist_df <-
  rare_dist_bray_p5 %>% 
  as.matrix %>% 
  as_tibble(rownames = "sample")

dbrda_meta_dist <-
  aes_data %>% 
  select(sample, passage_treat, passage, treat, ch4) %>% 
  distinct %>% 
  inner_join(rare_dist_df, by = "sample")

dbrda_ch4 <- dbrda(rare_dist_bray_p5 ~ rank(ch4), 
                      data = dbrda_meta_dist, 
                      na.action = "na.omit",
                      add = TRUE)

dbrda_aov <-
  dbrda_ch4 %>% 
  anova()

dbrda_percent_explained <- 100 * eigenvals(dbrda_ch4) / sum(eigenvals(dbrda_ch4))

dbrda_scores <-
  scores(dbrda_ch4, display = "sites") %>% 
  as_tibble(rownames = "sample")

dbrda_data <-
  aes_data %>% 
  select(sample, passage, treat, ch4) %>% 
  distinct %>% 
  inner_join(dbrda_scores, by = "sample")

dbrda_data %>%   
  ggplot(aes(x = dbRDA1, y = MDS1, shape = passage, color = treat)) +
  geom_point() +
  stat_ellipse(show.legend = FALSE) +
  coord_fixed() +
  scale_shape_discrete(name = "Passage") +
  scale_color_manual(name = "Treatment",
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2')) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())

# save(dbrda_aov, dbrda_percent_explained, file = here(der_dir, "dbrda.RData"))

```

## Differential Abundance

### Unique to each treatment

#### Family

```{r}
g5n <- 
  aes_data %>% 
  filter(passage_treat == "G5N") %>% 
  select(Family) %>% 
  distinct() %$%
  Family

g5p <- 
  aes_data %>% 
  filter(passage_treat == "G5P") %>% 
  select(Family) %>% 
  distinct() %$%
  Family

p_not_n <- g5p[!g5p %in% g5n]

n_not_p <- g5n[!g5n %in% g5p]

n_not_p_prevalence <-
  aes_data %>% 
  filter(passage_treat == "G5N") %>% 
  filter(Family %in% n_not_p) %>% 
  select(sample, Family, count) %>% 
  mutate(prevalence = as.numeric(count > 0)) %>% 
  select(-count) %>% 
  group_by(Family) %>% 
  summarize(prevalence = sum(prevalence))

n_not_p_prevalence %>% 
  ggplot(aes(x = prevalence)) +
  geom_histogram() +
  scale_x_continuous(breaks = 1:12) +
  theme_classic() +
  labs(title = "Unique to Positive")

# n_not_p_prevalence %>%
#   filter(prevalence == 1) %>% 
#   nrow

# n_not_p_prevalence %>%
#   filter(prevalence == 2) %>% 
#   nrow
# n_not_p_prevalence %>%
#   filter(prevalence > 2) %$%
#   Family

p_not_n_prevalence <-
  aes_data %>% 
  filter(passage_treat == "G5P") %>% 
  filter(Family %in% p_not_n) %>% 
  select(sample, Family, count) %>% 
  mutate(prevalence = as.numeric(count > 0)) %>% 
  select(-count) %>% 
  group_by(Family) %>% 
  summarize(prevalence = sum(prevalence))

p_not_n_prevalence %>% 
  ggplot(aes(x = prevalence)) +
  geom_histogram() +
  scale_x_continuous(breaks = 1:12) +
  theme_classic() +
  labs(title = "Unique to Neutral")
# 
# p_not_n_prevalence %>%
#   filter(prevalence == 1) %>% 
#   nrow
# 
# p_not_n_prevalence %>%
#   filter(prevalence == 2) %>% 
#   nrow
p_not_n_over_6 <- p_not_n_prevalence %>%
  filter(prevalence > 6)

```

First, I looked at which Families are unique to each treatment in passage 5.
There were `r length(n_not_p)` families unique to N and `r length(p_not_n)`
unique to P. However, of the  `r length(n_not_p)` unique to P, 
`r n_not_p_prevalence %>% filter(prevalence == 1) %>% nrow` had a prevalence
of 1 and `r n_not_p_prevalence %>% filter(prevalence == 2) %>% nrow` had
a prevalence of 2 leaving only a single family, which had a prevalence of
`r n_not_p_prevalence %>% filter(prevalence > 2) %$% prevalence`, which was `r n_not_p_prevalence %>% filter(prevalence > 2) %$% Family`.
Of the `r length(p_not_n)` Families unique to P, 2 had prevalences greater than
6, which were `r p_not_n_over_6$Family[1]` with a prevalence of
`r p_not_n_over_6$prevalence[1]` and `r p_not_n_over_6$Family[2]` with
a prevalence of `r p_not_n_over_6$prevalence[2]`.

#### Genus

```{r}
g5n <- 
  aes_data %>% 
  filter(passage_treat == "G5N") %>% 
  select(Genus) %>% 
  distinct() %$%
  Genus

g5p <- 
  aes_data %>% 
  filter(passage_treat == "G5P") %>% 
  select(Genus) %>% 
  distinct() %$%
  Genus

p_not_n <- g5p[!g5p %in% g5n]

n_not_p <- g5n[!g5n %in% g5p]

n_not_p_prevalence <-
  aes_data %>% 
  filter(passage_treat == "G5N") %>% 
  filter(Genus %in% n_not_p) %>% 
  select(sample, Genus, count) %>% 
  mutate(prevalence = as.numeric(count > 0)) %>% 
  select(-count) %>% 
  group_by(Genus) %>% 
  summarize(prevalence = sum(prevalence))

n_not_p_prevalence %>% 
  ggplot(aes(x = prevalence)) +
  geom_histogram() +
  scale_x_continuous(breaks = 1:12) +
  theme_classic() +
  labs(title = "Unique to Positive")

p_not_n_prevalence <-
  aes_data %>% 
  filter(passage_treat == "G5P") %>% 
  filter(Genus %in% p_not_n) %>% 
  select(sample, Genus, count) %>% 
  mutate(prevalence = as.numeric(count > 0)) %>% 
  select(-count) %>% 
  group_by(Genus) %>% 
  summarize(prevalence = sum(prevalence))

p_not_n_prevalence %>% 
  ggplot(aes(x = prevalence)) +
  geom_histogram() +
  scale_x_continuous(breaks = 1:12) +
  theme_classic() +
  labs(title = "Unique to Neutral")

p_not_n_over_6 <- p_not_n_prevalence %>%
  filter(prevalence > 6)

n_not_p_over_6 <- n_not_p_prevalence %>%
  filter(prevalence > 6)

```

#### Species

```{r}
g5n <- 
  aes_data %>% 
  filter(passage_treat == "G5N") %>% 
  select(Species) %>% 
  distinct() %$%
  Species

g5p <- 
  aes_data %>% 
  filter(passage_treat == "G5P") %>% 
  select(Species) %>% 
  distinct() %$%
  Species

p_not_n <- g5p[!g5p %in% g5n]

n_not_p <- g5n[!g5n %in% g5p]

n_not_p_prevalence <-
  aes_data %>% 
  filter(passage_treat == "G5N") %>% 
  filter(Species %in% n_not_p) %>% 
  select(sample, Species, count) %>% 
  mutate(prevalence = as.numeric(count > 0)) %>% 
  select(-count) %>% 
  group_by(Species) %>% 
  summarize(prevalence = sum(prevalence))

n_not_p_prevalence %>% 
  ggplot(aes(x = prevalence)) +
  geom_histogram() +
  scale_x_continuous(breaks = 1:12) +
  theme_classic() +
  labs(title = "Unique to Positive")

p_not_n_prevalence <-
  aes_data %>% 
  filter(passage_treat == "G5P") %>% 
  filter(Species %in% p_not_n) %>% 
  select(sample, Species, count) %>% 
  mutate(prevalence = as.numeric(count > 0)) %>% 
  select(-count) %>% 
  group_by(Species) %>% 
  summarize(prevalence = sum(prevalence))

p_not_n_prevalence %>% 
  ggplot(aes(x = prevalence)) +
  geom_histogram() +
  scale_x_continuous(breaks = 1:12) +
  theme_classic() +
  labs(title = "Unique to Neutral")

p_not_n_over_6 <- p_not_n_prevalence %>%
  filter(prevalence > 3)

n_not_p_over_6 <- n_not_p_prevalence %>%
  filter(prevalence > 3)

```

## Setup Differential Abundance Data

### Family

```{r setup-differential-abundance-data-family}

aes_G5 <-
  aes_data %>% 
  filter(passage == "G5")

aes_family_G5 <-
  aes_G5 %>% 
  select(asv, sample, treat, count, Family) %>%
  group_by(sample, Family) %>% 
  mutate(total = sum(count)) %>% 
  ungroup() %>% 
  select(sample, treat, Family, total) %>% 
  distinct()

prevalent_families <- 
  aes_family_G5 %>% 
  group_by(Family, treat) %>% 
  summarize(prev = sum(total > 0), .groups = "drop") %>% 
  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
  filter(N/12 > 0.1 & P/12 > 0.1) %$%
  Family

otu_table <- 
  aes_family_G5 %>% 
  filter(Family %in% prevalent_families) %>% 
  select(sample, Family, total) %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

rownames(otu_table) <- otu_table[,1]
otu_table <- otu_table[,-1]

sample_data <- aes_family_G5 %>% 
  select(sample, treat) %>% 
  distinct() %>% 
  as.data.frame()

rownames(sample_data) <- sample_data[,1]

```

#### Corncob

```{r}

da_analysis <- differentialTest(formula = ~ treat,
                                phi.formula = ~ 1,
                                formula_null = ~ 1,
                                phi.formula_null = ~ 1,
                                test = "Wald", 
                                boot = FALSE,
                                data = otu_table,
                                sample_data = sample_data,
                                fdr_cutoff = 0.05, B = 100,
                                taxa_are_rows = TRUE)

dt_data <- plot(da_analysis, 
                # level = c("Family"), 
                data_only = TRUE)

dt_plot <- function(dt_data) {
  ggplot(dt_data, aes(x = x, xmin = xmin, xmax = xmax, y = reorder(taxa, x))) +
    geom_pointrange() +
    geom_vline(xintercept = 0) +
    labs(x = "Estimate", y = NULL) +
    coord_cartesian(clip="off") +
    # theme(panel.border = element_blank()) +
    theme_classic() +
      theme(axis.text.y = element_markdown())

}

dt_data %>%
  mutate(taxa = str_replace(taxa, "(.*)", "*\\1*"),
         taxa = str_replace(taxa, "\\*(.*)_unclassified\\*",
                                "Unclassified *\\1*")) %>% 
  dt_plot()

```

#### ANCOM

```{r}
# 
# # Install ANCOMBC
# if (!requireNamespace("BiocManager", quietly=TRUE))
#     install.packages("BiocManager")
# BiocManager::install("ANCOMBC")
# 
# # Load ANCOM
# library(ANCOMBC)

# Setup phyloseq object
aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE),
                   sample_data(sample_data))

# Save/Load output so you don't have to rerun ANCOM
if(file.exists(here(der_dir, "ancom_out.rds"))) {
  out <- 
    readRDS(here(der_dir, "ancom_out.rds"))
} else {
  # Run Ancom
  out = ancom(phyloseq = aes_ps,  
              p_adj_method = "BH",
              prv_cut = 0.1, 
              lib_cut = 0, 
              main_var = "treat",
              rand_formula = NULL, 
              lme_control = NULL,
              struc_zero = TRUE, 
              neg_lb = FALSE, 
              alpha = 0.05, 
              n_cl = 1)  
  out %>% 
    saveRDS(here(der_dir, "ancom_out.rds"))
}

res = out$res
q_val = out$q_data
beta_val = out$beta_data
# Only consider the effect sizes with the corresponding q-value less than alpha
beta_val = beta_val * (q_val < 0.05) 
# Choose the maximum of beta's as the effect size
beta_pos = apply(abs(beta_val), 2, which.max) 
beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
                  FUN.VALUE = double(1))
# Number of taxa except structural zeros
n_taxa = ifelse(is.null(out$zero_ind), 
                nrow(feature_table), 
                sum(apply(out$zero_ind, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = 0.7 * (n_taxa - 1)

df_fig_w = res %>%
  dplyr::mutate(beta = beta_max,
                direct = case_when(
                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
                  TRUE ~ "Not Significant"
                  )) %>%
  dplyr::arrange(W)
df_fig_w$taxon_id = factor(df_fig_w$taxon_id, levels = df_fig_w$taxon_id)
df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
df_fig_w$direct = factor(df_fig_w$direct, 
                     levels = c("Negative", "Positive", "Not Significant"))

p_w = df_fig_w %>%
  ggplot(aes(x = taxon_id, y = W, color = direct)) +
  geom_point(size = 2, alpha = 0.6) +
  labs(x = "Taxon", y = "W") +
  scale_color_discrete(name = NULL) + 
  geom_hline(yintercept = cut_off, linetype = "dotted", 
             color = "blue", size = 1.5) +
  geom_text(aes(x = 2, y = cut_off + 0.5, label = "W[0.7]"), 
            size = 5, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank())
p_w

```

#### Aldex2

```{r}

conds <- colnames(otu_table) %>% 
  substr(3, 3)

if(file.exists(here(der_dir, "aldex_x_all.rds"))) {
  x.all <- 
    readRDS(here(der_dir, "aldex_x_all.rds"))
} else {
  x.all <- aldex(otu_table, 
                 conds, mc.samples=1000, test="t", effect=TRUE,
                 include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
  x.all %>% 
    saveRDS(here(der_dir, "aldex_x_all.rds"))
}

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
    ylab="Difference")
aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
    ylab="Difference")

par(mfrow=c(1,2))
plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")
legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))

plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")

x.all_effect_1 <-
  x.all %>% 
  filter(abs(effect) > 1) %>% 
  as_tibble(rownames = "Family")

# x.all_effect_1 %>% 
#   # inner_join(taxa_table, by = "asv")
#   ggplot(aes(y = reorder(Family, effect), x = effect)) +
#   geom_point()

# x.all_effect_1 %>% 
#   select(Family, effect)

```

#### Comparison

##### All Significant

```{r}

# CORNCOB
corncob_taxa <- dt_data$taxa

# ANCOM
ancom_taxa <- as.character(df_fig_w[df_fig_w$detected_0.7, ]$taxon_id)

# ALDEX2
aldex_taxa <- x.all_effect_1$Family

at_least_one_taxa <-
	aes_family_G5 %>%
	group_by(sample) %>%
	mutate(rel_abund = (total / sum(total)) * 100) %>%
	filter(Family %in% ancom_taxa | Family %in% aldex_taxa | Family %in% corncob_taxa) %>%
	mutate(Family = str_replace(Family, "(.*)", "*\\1*"),
	       Family = str_replace(Family, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))

at_least_one_iqm <- 
	at_least_one_taxa %>%
	group_by(treat, Family) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")

sort_order <-
	at_least_one_taxa %>%
	group_by(Family) %>%
	summarize(median = median(rel_abund)) %>%
	arrange(median)

at_least_one_taxa$Family <- factor(at_least_one_taxa$Family, levels = sort_order$Family)

at_least_one_taxa %>%
	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
			data = at_least_one_iqm, 
			position = position_dodge(width = 0.75), 
			shape = 21, 
			color = "black") +
	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Relative Abundance (%)", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2'))

```

##### Consensus Taxa

```{r}
aldex_effect <-
  x.all_effect_1 %>% 
  select(Family, effect)

all_taxa <-
	aes_family_G5 %>%
	group_by(sample) %>%
	mutate(rel_abund = (total / sum(total)) * 100) %>%
  left_join(aldex_effect, by = "Family") %>% 
	filter(Family %in% ancom_taxa & Family %in% aldex_taxa & Family %in% corncob_taxa) %>% 
  ungroup()

da_fams <-
  all_taxa %>% 
  select(Family) %>% 
  distinct() %$%
  Family

all_taxa <-
  all_taxa %>%
	mutate(Family = str_replace(Family, "(.*)", "*\\1*"),
	       Family = str_replace(Family, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))
all_taxa$Family[all_taxa$Family == "Unclassified *0319-6G20*"] <- "Uncultured family 0319-6G20"
# all_taxa$Family[all_taxa$Family == "Unclassified *Chloroplast*"] <- "Chloroplast"
all_taxa$Family[all_taxa$Family == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
all_taxa$Family[all_taxa$Family == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
all_taxa$Family[all_taxa$Family == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"

all_iqm <- 
	all_taxa %>%
	group_by(treat, Family) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")

sort_order <-
	all_taxa %>%
  select(Family, effect) %>% 
  distinct() %>% 
  arrange(effect)
	# group_by(Family) %>%
	# summarize(median = median(rel_abund)) %>%
	# arrange(median)

all_taxa$Family <- factor(all_taxa$Family, levels = sort_order$Family)
all_iqm$Family <- factor(all_iqm$Family, levels = sort_order$Family)
all_taxa <-
  all_taxa %>% 
  mutate(direction = if_else(effect > 0, 0, 1))
all_iqm <-
all_taxa %>% 
  select(Family, direction) %>% 
  distinct() %>% 
  full_join(all_iqm, by = "Family")
# (p_log10_cl <-
# 	all_taxa %>%
# 	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
# 	geom_point(position=position_jitterdodge()) +
# 	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
# 			data = all_iqm, 
# 			position = position_dodge(width = 0.75), 
# 			shape = 21, 
# 			color = "black") +
# 	theme_bw() +
# 	theme(axis.text.y = element_markdown()) +
# 	labs(x = "Relative Abundance (%)", y = NULL) +
# 	scale_color_manual(name = "Treatment",
# 			   labels = c('Control', 'Selection'),
# 			   values = c('gray40', 'darkorange2')) +
# 	scale_fill_manual(name = "Treatment",
# 			   labels = c('Control', 'Selection'),
# 			   values = c('gray40', 'darkorange2')) +
#     facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
# theme(
#   strip.background = element_blank(),
#   strip.text.y = element_blank()
# ) +
# 	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
# 		      labels = c(0.001, 0.01, 0.1, 1)))

# save(p_log10, da_fams, file = here(der_dir, "da_fig.RData"))

# ggsave(here(fig_dir, 'fig5.jpg'), p_log10, width = 5, height = 5)
# ggsave(here(fig_dir, 'fig5.svg'), p_log10, width = 5, height = 5)
# ggsave(here(fig_dir, 'fig5_cl.png'), p_log10_cl, width = 5, height = 5)
(p_log10_bw <-
	all_taxa %>%
	ggplot(aes(x = rel_abund, y = Family, fill = treat)) +
	geom_point(position=position_jitterdodge(jitter.width = 0.1,
	                                         dodge.width = 0.4), shape = 21, size = 1) +
        theme_bw(base_size = 7) +
            theme(panel.border = element_rect(fill = NA, linewidth = 1.0625),
                  axis.line = element_blank(),
                  text = element_text(family = 'Arial')) +
	# geom_jitter(shape = 21, height = 0.1) +
	# geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, color = treat), 
	# 		data = all_iqm, 
	# 		position = position_dodge(width = 0.75), 
	# 		shape = 21, 
	# 		# color = "black"
	# 		) +
	# theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Relative Abundance (%)", y = NULL) +
	# scale_color_manual(name = "Treatment",
	# 		   labels = c('Control', 'Selection'),
	# 		   values = c('black', 'gray')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Control', 'Selection'),
			   values = c('white', 'black')) +
    facet_grid(direction ~ ., 
               scale = "free_y",
               space = "free")  + 
theme(
  strip.background = element_blank(),
  strip.text.y = element_blank()) +
	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
		      labels = c(0.001, 0.01, 0.1, 1))
    )
ggsave(here(fig_dir, 'fig5.eps'), 
       p_log10_bw, 
       width = 100, 
       height = 88,
       units = 'mm',
       device = cairo_ps
       )

# (p_log10_gs <-
# 	all_taxa %>%
# 	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
# 	# geom_point(position=position_jitterdodge(), shape = 21) +
# 	geom_jitter(#shape = 21, 
# 	            height = 0.1) +
# 	# geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, color = treat), 
# 	# 		data = all_iqm, 
# 	# 		position = position_dodge(width = 0.75), 
# 	# 		shape = 21, 
# 	# 		# color = "black"
# 	# 		) +
# 	theme_bw() +
# 	theme(axis.text.y = element_markdown()) +
# 	labs(x = "Relative Abundance (%)", y = NULL) +
# 	# scale_color_manual(name = "Treatment",
# 	# 		   labels = c('Control', 'Selection'),
# 	# 		   values = c('black', 'gray')) +
# 	scale_color_manual(name = "Treatment",
# 			   labels = c('Control', 'Selection'),
# 			   values = c('gray50', 'black')) +
#     facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
# theme(
#   strip.background = element_blank(),
#   strip.text.y = element_blank()
# ) +
# 	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
# 		      labels = c(0.001, 0.01, 0.1, 1)))
# ggsave(here(fig_dir, 'fig5_gs.eps'), p_log10_gs, width = 5, height = 5)

## All
# ggsave(here(fig_dir, "shared_taxa.png"), p)
# ggsave(here(fig_dir, "shared_taxa_log.png"), p_log10, width = 5, height = 5)

# aes_family_rclr_plot <-
#   aes_family_rclr %>% 
# 	filter(Family %in% ancom_taxa & Family %in% aldex_taxa & Family %in% corncob_taxa) %>% 
#   pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
#   select(Family, treat, rep, passage, rclr) %>% 
#   distinct() %>% 	
#   mutate(Family = str_replace(Family, "(.*)", "*\\1*"),
# 	       Family = str_replace(Family, "\\*(.*)_unclassified\\*",
# 				    "Unclassified *\\1*"))
# 
# aes_family_rclr_plot$Family[aes_family_rclr_plot$Family == "Unclassified *0319-6G20*"] <- "Uncultured family 0319-6G20"
# aes_family_rclr_plot$Family[aes_family_rclr_plot$Family == "Unclassified *Chloroplast*"] <- "Chloroplast"
# aes_family_rclr_plot$Family <- factor(aes_family_rclr_plot$Family, levels = sort_order$Family)
# ggplot(aes_family_rclr_plot, aes(x = rclr, y = Family, color = treat)) +
#   facet_grid( ~ passage) +
#   geom_point(position = position_jitterdodge())

(p_effect <-
	all_taxa %>%
	ggplot(aes(x = effect, y = Family)) +
	geom_point() +

	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Aldex2 Effect Size", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Control', 'Selection'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Control', 'Selection'),
			   values = c('gray40', 'darkorange2'))
    )

```

## Methanogens and Methanotrophs

Methanotrophy is (mostly) conserved at the family level. The Type I and Type X methanotrophs are in the Gammaproteobacteria in the families _Methylococcaceae_ and _Methylothermaceae_. There are also methanotrophs in the Verrumicrobiota in the family _Methylacidiphilaceae_. The Type II methanotrophs are in the _Alphaproteobacteria_ and are found in the families _Methylocystaceae_ and _Beijerinckiaceae_. However, the _Beijerinckiaceae_ also contain non-methanotrophs including N2-fixers and photoautotrophs.

```{r methanotroph-lists}
# List of methanotrophs from Table 1 of Stein 2012 
troph_genera <- c('Methylococcus',
                  'Methylocaldum',
                  'Methylobacter',
                  'Methylomicrobium',
                  'Methylomonas',
                  'Methylosarcina',
                  'Methylosoma',
                  'Crenothrixa',
                  'Clonothrixa',
                  'Methylosphaera',
                  'Methylohalobiusa',
                  'Methylothermusa',
                  'Methylocystis',
                  'Methylosinus',
                  'Methylocella',
                  'Methylocapsa',
                  'Methyloferula',
                  'Methylomirabilis',
                  'Methylacidiphilum')
troph_genera <- c(troph_genera, paste(troph_genera, "unclassified", sep = "_"))

methanotrophs <- c('Methylococcaceae', 
                   'Methylothermaceae', 
                   'Methylocystaceae', 
                   'Beijerinckiaceae', 
                   'NC10',
                   'Methylacidiphilaceae')
methanotrophs <- c(methanotrophs, paste(methanotrophs, "unclassified", sep = "_"))

troph_orders <- c('Methylococcales')

methanogen_classes <- c("Methanococci", "Methanomicrobia", "Methanobacteria")

methanogen_classes <- c(methanogen_classes, paste(methanogen_classes, "unclassified", sep = "_"))

```

```{r}

aes_data %>% 
  select(Family, Genus, count, sample) %>% 
  group_by(sample, Genus, Family) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  group_by(sample) %>% 
  mutate(rel_abund = total / sum(total)) %>% 
  mutate(rclr = log(total/gm(total))) %>% 
  filter(Family %in% methanotrophs) %>% 
  select(-Family) %>% 
  mutate(treat = substr(sample, 3, 3), passage = substr(sample, 1, 2)) %>% 
  pivot_longer(rel_abund:rclr, names_to = "method", values_to = "value") %>% 
  ggplot(aes(x = value, y = Genus, color = treat)) +
  geom_jitter() +
  facet_grid(passage ~ method, scales = "free") +
  theme_bw()

aes_data %>% 
  select(Family, count, sample) %>% 
  group_by(sample, Family) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  group_by(sample) %>% 
  mutate(rel_abund = total / sum(total)) %>% 
  mutate(rclr = log(total/gm(total))) %>% 
  filter(Family %in% methanotrophs) %>% 
  mutate(treat = substr(sample, 3, 3), passage = substr(sample, 1, 2)) %>% 
  pivot_longer(rel_abund:rclr, names_to = "method", values_to = "value") %>% 
  ggplot(aes(x = value, y = Family, color = treat)) +
  geom_jitter() +
  facet_grid(passage ~ method, scales = "free") +
  theme_bw()

troph_asvs <-
  aes_data %>% 
  select(Kingdom:Species, asv) %>% 
  distinct() %>% 
  filter(Genus %in% troph_genera) %$%
  asv

troph_asvs <-
  aes_data %>% 
  select(Kingdom:Species, asv) %>% 
  distinct() %>% 
  filter(Family %in% methanotrophs) %$%
  asv

```

Of the five methanotroph families, two are represented in this dataset:
_Methylacidiphilaceae_ and _Beijerinckiaceae_. From these two families,
there are eleven genera represented. These are spread across
`r length(troph_asvs)` ASVs

```{r}

aes_data %>% 
  select(Class, count, sample) %>% 
  group_by(sample, Class) %>% 
  mutate(total = sum(count)) %>% 
  select(-count) %>% 
  distinct() %>% 
  filter(Class %in% methanogen_classes)

```

A grand total of 10 reads map to methanogen classes across the entire dataset,
including 2 reads in Methanobacteria (sample G2N02) and 8 reads of
Methanomicrobia (sample G2N07).

### Methanotroph differential abundance

```{r}

troph_data <-
  aes_data %>% 
  select(sample, asv, treat, passage, count, Genus, Family) %>% 
  filter(passage == "G5") %>% 
  select(-passage) %>% 
  filter(asv %in% troph_asvs)

troph_data_family <-
  troph_data %>% 
  select(sample, Family, count) %>% 
  group_by(sample, Family) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

rownames(troph_data_family) <- troph_data_family[, 1]
troph_data_family <- troph_data_family[, -1]

troph_data_genus <-
  troph_data %>% 
  select(sample, Genus, count) %>% 
  group_by(sample, Genus) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

rownames(troph_data_genus) <- troph_data_genus[, 1]
troph_data_genus <- troph_data_genus[, -1]

sample_data <- 
  troph_data %>% 
  select(sample, treat) %>% 
  distinct() %>% 
  as.data.frame()
conds <- sample_data[, 2]

```

```{r corncob-methanotrophs}
if(file.exists(here(der_dir, "cc_family.rds"))) {
  cc_family <- readRDS(here(der_dir, "cc_family.rds"))
} else {
cc_family <- differentialTest(formula = ~ treat,
                                phi.formula = ~ 1,
                                formula_null = ~ 1,
                                phi.formula_null = ~ 1,
                                test = "Wald", 
                                boot = TRUE,
                                data = troph_data_family,
                                sample_data = sample_data,
                                fdr_cutoff = 0.05, B = 1000)  
cc_family %>% saveRDS(here(der_dir, "cc_family.rds"))
}

plot(cc_family)

if(file.exists(here(der_dir, "cc_genus.rds"))) {
  cc_genus <- 
    readRDS(here(der_dir, "cc_genus.rds"))
} else {
cc_genus <- differentialTest(formula = ~ treat,
                                phi.formula = ~ 1,
                                formula_null = ~ 1,
                                phi.formula_null = ~ 1,
                                test = "Wald", 
                                boot = TRUE,
                                data = troph_data_genus,
                                sample_data = sample_data,
                                fdr_cutoff = 0.05, B = 1000) 
cc_genus %>% 
    saveRDS(here(der_dir, "cc_genus.rds"))
}

plot(cc_genus)

```

```{r save_outputs}

# save(p_response, herit_plot, p_beta, p_log10, file = here(der_dir, 'plots.RData'))
save(response_model, compare_response_mods, per_slope, rh_model, kt, wt, rare_median, s5, min_seqs, tidy_permanova, dbrda_aov, dbrda_percent_explained, da_fams, file = here(der_dir, 'models.RData'))

# save(p_response, file = here(der_dir, 'response_plot.RData'))
# save(response_model, compare_response_mods, per_slope, file = here(der_dir, 'response_model.RData'))
# 
# save(herit_plot, file = here(der_dir, "herit_plot.RData"))

# save(rh_model, file = here(der_dir, 'herit_model.RData'))
# save(kt, wt, rare_median, s5, min_seqs, file = here(der_dir, "richness_models.RData"))
# save(p_beta, file = here(der_dir, "beta_fig.RData"))
# save(tidy_permanova, file = here(der_dir, "beta_model.RData"))
# save(dbrda_aov, dbrda_percent_explained, file = here(der_dir, "dbrda.RData"))
# save(p_log10, da_fams, file = here(der_dir, "da_fig.RData"))

```